### 2019-09-19 Documentation/Pollution_AEJ_Final.tex

% Article Format: American Economic Review
% Xin Tang @ Stony Brook University
% Last Updated: August 2014
\documentclass[twoside,12pt,leqno]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                    Text Layout                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Page Layout
\usepackage[hmargin={1.2in,1.2in},vmargin={1.5in,1.5in}]{geometry}
\topmargin -1cm        % read Lamport p.163
\oddsidemargin 0.04cm   % read Lamport p.163
\evensidemargin 0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth 16.59cm
\textheight 21.94cm
\renewcommand\baselinestretch{1.15}
\parskip 0.25em
\parindent 1em
\linespread{1}

% Set header and footer
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[LE,RO]{\thepage}
%\fancyhead[CE]{\textit{JI QI, XIN TANG AND XICAN XI}}
%\fancyhead[CO]{\textit{THE SIZE DISTRIBUTION OF FIRMS AND INDUSTRIAL POLLUTION}}
\cfoot{}
\renewcommand{\headrulewidth}{0pt}
%\renewcommand*\footnoterule{}
%\setcounter{page}{1}

% Font
\renewcommand{\rmdefault}{ptm}
\renewcommand{\sfdefault}{phv}
%\usepackage[lite]{mtpro2}
% use Palatinho-Roman as default font family
%\renewcommand{\rmdefault}{ppl}
\usepackage[scaled=0.88]{helvet}
\makeatletter   % Roman Numbers
\newcommand*{\rom}[1]{\expandafter\@slowromancap\romannumeral #1@}
\makeatother
\usepackage{CJK}

% Section Titles
\renewcommand\thesection{\textnormal{\textbf{\Roman{section}.}}}
\renewcommand\thesubsection{\textnormal{\Alph{subsection}.}}
\usepackage{titlesec}
\titleformat*{\section}{\bf \center}
\titleformat*{\subsection}{\it \center}
\renewcommand{\refname}{\textnormal{REFERENCES}}

% Appendix
\usepackage[title]{appendix}
\renewcommand{\appendixname}{APPENDIX}

% Citations
\usepackage[authoryear,comma]{natbib}
\renewcommand{\bibfont}{\small}
\setlength{\bibsep}{0em}
\usepackage[%dvipdfmx,%
            bookmarks=true,%
            pdfstartview=FitH,%
            breaklinks=true,%
            colorlinks=true,%
            %allcolors=black,%
            citecolor=blue,
            linkcolor=red,
            pagebackref=true]{hyperref}

% Functional Package
\usepackage{enumerate}
\usepackage{url}      % This package helps to typeset urls

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                    Mathematics                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath,amssymb,amsfonts,amsthm,mathrsfs,upgreek}
% Operators
\newcommand{\E}{\mathbb{E}}
\newcommand{\e}{\mathrm{e}}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\plim}{plim}
\renewcommand{\vec}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\gvec}[1]{{\boldsymbol{#1}}}

% New Environments
\newtheorem{result}{Result}
\newtheorem{assumption}{Assumption}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\setlength{\unitlength}{1mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                 Tables and Figures                   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{threeparttable,booktabs,multirow,array} % This allows notes in tables
\usepackage{floatrow} % For Figure Notes
\floatsetup[table]{capposition=top}
\usepackage[font={sc,footnotesize}]{caption}
\DeclareCaptionLabelSeparator{aer}{---}
\captionsetup[table]{labelsep=aer}
\captionsetup[figure]{labelsep=period}
\usepackage{graphicx,pstricks,epstopdf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Begin Document Content %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\vspace{-1cm}\Large{{\textsf{The Size Distribution of Firms and Industrial Water Pollution\\A Quantitative Analysis of China}}}\thanks{Disclaimer: The results and conclusions in this paper are those of the authors, and do not necessarily reflect the views of the China Ministry of Environmental Protection, or those of the International Monetary Fund. An online appendix containing additional materials is available on the author's website: {\color{blue}https://sites.google.com/site/zjutangxin/}.}}
\author{\normalsize{\textit{By}\ \textsc{Ji Qi},\ \textsc{Xin Tang},\textsc{\ and\ }\textsc{Xican Xi}}\thanks{Qi: Chinese Academy for Environmental Planning, Room 709-B BeiKeChuangYe Tower, No.10 Dayangfang, Beiyuan Road, Beijing, China, 100012 (e-mail: qiji@caep.org.cn); Tang: Wuhan University and International Monetary Fund, 700 19th St NW, Washington D.C., United States, 20431 (e-mail:zjutangxin@gmail.com); Xi: School of Economics, Fudan University, 600 Guoquan Road, Shanghai, China, 200433 (e-mail:xicanxi@fudan.edu.cn). We are grateful for the comments from Alexis Anagnostoupolos, Marina Azzimonti, Juan Carlos Conesa, Berthold Herrendorf, Nicolai Kuminoff, Pietro Peretto, Edward Prescott, Diego Restuccia, Todd Schoellman, Kerry Smith, Daniel Yi Xu, Gustavo Ventura, the editor and three anonymous referees. We also benefited from the comments received at Arizona State University, the Chinese Academy of Environmental Planning, Duke University, the IMF, U.S International Trade Commission, Peking University, Shanghai University of Finance and Economics, Singapore Management University, Stony Brook University, Washington State University, Wuhan University, and Zhejiang University. We sincerely thank the China Ministry of Ecology and Environment for granting us access to the National General Survey of Pollution Sources data. This paper has been screened to ensure that no confidential data are revealed.}} %The authors declare that they have no relevant or material financial interests that relate to the research described in this paper.}}
\date{\textnormal{\today}}

\begin{document}
\maketitle
%\renewcommand{\abstractname}{}
\vspace{-1.5em}
%\begin{abstract}
\begin{center}
\begin{minipage}[c]{0.8\textwidth}
{%\normalsize
%\noindent \textit{We study the role of firm-size distribution in industrial pollution using a quantitative model featuring a distribution of firms. Micro-level data support a negative correlation between firm size and polluting intensity. Fixed costs associated with advanced treatment technologies are shown to be important in explaining the negative size-intensity correlation. Evidence suggests that product market frictions limit the expansion of productive firms in China, which leads to survivals of unprofitable small polluting firms. Quantitative analysis using a model consistent with these empirical regularities shows that while having moderate effect on output, size distribution exerts considerable effect over industrial pollution. (\textit{JEL} E01, E23, O44, Q52, Q53, Q56)}
\noindent \textit{We argue that misallocation across firms amplifies industrial water pollution by distorting the firm size distribution in China. Firm-level data indicate that larger firms are more likely to use clean technology but face higher distortions. In a heterogeneous firms model with an endogenous choice of pollution treatment technologies, we show that distortions that increase with firm-level TFP lower the adoption of clean technology, amplify aggregate pollution intensity, and lower aggregate output. Quantitatively, eliminating these correlated distortions would increase output by 30\% and decrease pollution by 20\%. Meanwhile, environmental regulations have sizable impact on pollution but only limited effect on output. (JEL E23, L11, O44, Q52, Q53, Q56)}}
\end{minipage}
\end{center}

\textnormal{Severe} pollution has accompanied China's remarkable economic growth in the past few decades, causing environmental degradation, public health damage, and millions of premature deaths.\footnote{\citet{WorldBank_Cost_2007} estimates that the cost of air and water pollution in China is about 6\% of GDP. The health cost of pollution is estimated to be 4\% of GDP, the majority of which is associated with premature deaths caused by pollution. %\citet{Ebenstein:2012} finds that industrial water pollution has a large effect on deaths from digestive cancers in China, while \citet{Ebensteinetal:2015} show that air pollution is negatively associated with life expectancy.
See \citet{Vennemoetal:2009} and \citet{ZhengKahn:2013} for review of empirical studies on environmental pollution in China.} %Therefore it has become a major concern for both the public and policymakers. %\footnote{The Chinese government has initiated a campaign against pollution. According to the \textit{Report on the Work of the Government 2014}, the State Council has allocated \$600 billion of special funds to reduce air and water pollution during the Thirteenth Five-Year Plan (2016-2020).}
Understanding the driving forces behind this severe pollution is crucial for designing and evaluating anti-pollution policies. An outstanding question is whether the severe pollution is an inevitable consequence of rapid economic growth, or largely exacerbated by policy distortions and/or market inefficiencies.

In this paper, we argue that misallocation across firms amplifies industrial water pollution by distorting the firm size distribution in China. Using nationally representative data on firm production, pollution emissions and pollution treatment technologies, we document two facts on firm size and pollution intensity (pollutants per output) in China.\footnote{We use the \textit{National General Survey of Pollution Sources} (NGSPS) and \textit{National Economic Census} (CNEC) in China. Section \rom{1} provides more details.} First, large firms have lower pollution intensity than small firms, with an elasticity of $-$0.37 between firm size and pollution intensity for the top five polluting industries in China.\footnote{The industries are ranked by their emissions of \textit{Chemical Oxygen Demand} (COD). The five industries are (with two-digit GB/T4547-2002 code in parentheses): Paper and Paper Products (C22), Processing of Food from Agricultural Products (C13), Textiles (C17), Raw Chemical Materials and Chemical Products (C26), and Beverages (C15).} % We find 7- to 32-fold differences in pollution intensity between firms in the largest and smallest quartiles of the firm size distribution for the top five polluting industries in China.
Furthermore, we find that this is not only due to fewer pollutants per output are generated by large firms during production, but also that large firms are also more likely to use advanced end-of-pipe treatment technologies that require sizable fixed installation costs. %\footnote{The negative correlation between firm size/productivity and pollution intensity has also been found in other countries. For example, for developing countries \citet{Dasguptaetal:1998} find that small firms have higher air pollution intensity in Brazil and Mexico, and \citet{Martin:2013} and \citet{BarrowsOllivier:2016} find that smaller firms have higher energy intensity in India. For developed countries, \citet{Bloometal:2010} find that better managed firms have lower energy intensity in the United Kingdom, and \citet{ShapiroWalker:2015} find that more productive firms have lower pollution intensity in the United States. However, the richness of our data allows us to establish empirically a unique link between firm size and the adoption of clean technology, and thus to provide a mechanism for this negative correlation that is observable.}\textsuperscript{, }
%\footnote{Our data reveal that large firms are more likely to not only use advanced \textit{end-of-pipe} treatment technologies to remove a larger proportion of pollutants, but also generate less pollutants in production. This could be because that the production technologies used by large firms are also more environmentally friendly, or also because they are using less inputs. While we cannot directly observe firms' production technologies, we do have data on firms' end-of-pipe pollution treatment technologies.}
Second, large firms account for a smaller proportion of the total employment in China than in the United States. For the top five polluting industries in China, firms with more than 400 employees account for 20\% of the total employment compared with close to 70\% for the U.S. counterparts. %\footnote{See \citet{Axtell:2001}, \citet{Luttmer:2007}, and \citet{RossihansbergWright:2007} for theories and evidence regarding the heavy right tail of the U.S. firm and establishment distributions.}
If we take the U.S. economy as a less distorted benchmark, these facts suggest that distortions to firm size distribution could have important environmental consequences in China. %This fact is suggestive of the size-dependent distortions that reallocate factors away from large productive firms in China if we take the U.S. economy as a relatively distortion-free benchmark. Indeed, we find that in China, the average product of labor and capital increase strongly with firm size and productivity [\citet{HsiehKlenow:2009, HsiehKlenow:2014}].\footnote{This measure of distortion captures the effects of many of the factors that distort factor allocation across firms, including taxes and subsidies, regulations, local protectionism, transportation costs, and financial frictions. A prominent example of such size-dependent distortions is internal trade barriers that impede the flow of goods across regions, which presumably affect large firms more since these are more likely to sell goods across regions [\citet{Young:2000}, \citet{TombeZhu:2015}, and \citet{Eberhardtetal:2016}]. In addition, as is the case in many other developing countries, small firms are more likely to evade taxes because of imperfect tax enforcement.}
%Formally, we follow the approach developed by \citet{HsiehKlenow:2009, HsiehKlenow:2014}, and estimate distortions that affect factor allocation across firms using firm-level variations in average products of capital and labor. We find that firm-level distortions increase with firm size and productivity, consistent with the fact that large firms account for a smaller fraction of total employment in China.\footnote{It is well understood in the literature that this measure of distortions captures the effects of many factors that distort factor allocation across firms, including taxes and subsidies, regulations, transportation costs and financial frictions, among others. A prominent example of such size-dependent distortions is internal trade barriers that impede the flow of goods across regions, which presumably affect large firms more, since they are more likely to sell goods across regions. \citet{TombeZhu:2015} find that the internal trade barriers have a large impact on China's aggregate productivity. In addition, as is the case in many other developing countries, small firms are more likely to evade taxes due to imperfect tax enforcement.}

Motivated by these facts, we build a two-sector (polluting and non-polluting) model of heterogeneous firms based on \citet{Lucas:1978b} to evaluate how firm-level distortions affect aggregate output and pollution. The model has the following key features. First, firms in each sector differ in their productivity levels. There is a stand-in household with a continuum of members, and each member draws a sector specific managerial ability from a given distribution. When members choose to be manager, they enter the sector their managerial ability belongs to and the ability determines the TFP of the firm they run. Second, firms differ in pollution intensity as well. As in \citet{CopelandTaylor:1994}, pollution emissions are a by-product of production for firms in the polluting sector. Following Fact 1 above, we assume that generated emissions per output during production is decreasing in firm output.\footnote{While our data reveal that large firms generate fewer pollutants per output, we do not know precisely the reasons behind that. %It may be because large firms use fewer or cleaner inputs to produce the same amount of output. For example, \citet{Bloometal:2010} find that better managed firms use less energy to produce the same amount of output in the U.K.. As another example, the paper pulp industry in China uses two different types of inputs: bagasse and wood. While bagasse, which generates 140--180 kg COD per ton, is mostly used by firms with an annual production of {\color{black}less} than 100 k-tons, wood, which generates 30--55 kg COD per ton, is mostly used by firms with an annual production of {\color{black}more} than 100 k-tons.
Without further data, we model the emissions generated during production in a reduced-form way, which is then calibrated to the data. Section \rom{1}.B discusses this point in more details.} Third, firms have the option to adopt an advanced end-of-pipe treatment technology, which we refer to as \textit{clean} technology later, after paying a fixed adoption cost. Firms in the polluting sector are subject to environmental regulations such that firms that do not use the clean technology face a penalty that takes away a fraction of their profits. %\footnote{In reality, a powerful tool used by the Chinese government to enforce environmental regulations, is to randomly audit and then shut down firms that do not comply with environmental regulations for a certain period of time, which would cost these firms a fraction of their profits. See for example, Trevor Nace, ``China shuts down tens of thousands of factories in widespread pollution crackdown,'' Forbes, October 2017.} % (\url{https://www.forbes.com/sites/trevornace/2017/10/24/china-shuts-down-tens-of-thousands-of-factories-in-widespread-pollution-crackdown})
As a result, our model predicts that only firms above a certain size threshold would choose the clean technology, which is qualitatively consistent with the fact that large firms are more likely to use clean technology.

Finally, we assume that firms face distortions that increase with their TFP, or \emph{correlated} distortions following the terminology of \citet{RestucciaRogerson:2008}, which are modeled as progressive output taxes.\footnote{\citet{Guneretal:2008} study a closely related type of distortions, namely \emph{size-dependent} distortions, which increase with firm or establishment size (the amount of capital, labor or output). An example is the many labor regulations in Italy, Spain and France that start to bind on firms with more than a certain number of employees. Both correlated and size-dependent distortions lead to a reallocation of resources from high TFP to low TFP firms, and they are essentially equivalent when distortions do not lead to rank reversal defined as more productive firms being smaller [\citet{Hopenhayn:2014}].} We find in our data that firms with higher TFP exhibit substantially higher average products of capital and labor in both the polluting and non-polluting sectors, indicating that more productive firms do face higher distortions [\citet{HsiehKlenow:2009, HsiehKlenow:2014}]. %\footnote{Pioneered by \citet{HsiehKlenow:2009, HsiehKlenow:2014}, this approach to measure distortions captures the effects of a number of factors that distort factor allocation across firms, including taxes and subsidies, regulations, local protectionism, transportation costs, and financial frictions. A prominent example of correlated distortions is internal trade barriers that impede the flow of goods across regions, which presumably affect large productive firms more since these are more likely to sell goods across regions [\citet{Young:2000}, \citet{TombeZhu:2015}, and \citet{Eberhardtetal:2016}]. In addition, as is the case in many other developing countries, small firms are more likely to evade taxes because of imperfect tax enforcement.}
As emphasized by previous studies, such as \citet{RestucciaRogerson:2008} and \citet{HsiehKlenow:2014}, correlated distortions lower both mean firm size and aggregate output, since they limit the operations of productive firms and allow too many unproductive firms to survive and expand. %\footnote{See \citet{RestucciaRogerson:2008}, \citet{Bartelsmanetal:2013}, \citet{HsiehKlenow:2014}, \citet{AdamopoulosRestuccia:2014}, \citet{BentoRestuccia:2016} among many others.}
A new insight from our model is that correlated distortions also increase aggregate pollution intensity for two reasons. First, they lower the profits of productive firms which reduces their incentives to adopt the clean technology. Therefore fewer firms adopt clean technology in the equilibrium with correlated distortions. Second, they take production factors away from large productive firms that also have lower pollution intensity, which drives up the average pollution intensity further.

%When connecting our model with firm-level data, we are able to quantify the aggregate effects of correlated distortions in a comprehensive way, taking into account the general equilibrium responses. Specifically, we can assess structurally the three channels through which distortions affect aggregate pollution, following the convention in environmental economics: scale (change in aggregate output), composition (reallocation across sectors), and technique (change in sectoral pollution intensity).\footnote{See, for example, \citet{GrossmanKrueger:1993}, \citet{CopelandTaylor:2004}, and \citet{Levinson:2009}, among others.} Moreover, our model and data allow us to further decompose the technique effect into: within sector reallocation across firms, and the change in the adoption of clean technology.

Conventionally, environmental economics decomposes changes in aggregate pollution into three effects: scale (change in aggregate output), composition (reallocation across sectors), and technique (change in sectoral pollution intensity) effects.\footnote{See for example, \citet{GrossmanKrueger:1993}, \citet{CopelandTaylor:2004} and \citet{Levinson:2009} among others.} Our model allows us to assess structurally how distortions affect aggregate pollution through the three channels. Moreover, our model and data allow us to further decompose the technique effect into within sector reallocation across firms and the change in the adoption of clean technology.

To discipline our quantitative analysis, we calibrate our model to match the observed pollution intensity, the adoption of clean technology, the correlation between distortions and firm TFP, and the firm size distribution in the Chinese data. We map the polluting sector to the top five polluting industries mentioned above, which account for 80\% of the total emissions in the data, and map the non-polluting sector to the rest of the manufacturing sector. We then use the calibrated model to evaluate the effects of two policies: removing correlated distortions from both sectors and strengthening environmental regulations.\footnote{Here we follow the so-called \emph{indirect} approach in the study of misallocation [\citet{RestucciaRogerson:2013}] by quantifying the overall effects of a variety of policies and institutions that relocate production factors from high to low TFP firms. Several examples of such policies include size-dependent taxes and subsidies, selective enforcement of regulations, financial constraints, poor enforcement of contracts, and internal trade costs, among others. This approach has been widely used in the misallocation literature, for instance, by \citet{HsiehKlenow:2009, HsiehKlenow:2014}, \citet{Bartelsmanetal:2013}, \citet{AdamopoulosRestuccia:2014}, and \citet{BentoRestuccia:2016}, among others.}%\footnote{Here we take the so-called \emph{indirect} approach in the study of misallocation (\citet{RestucciaRogerson:2013}), and quantify the overall effects of misallocation without identifying the specific underlying sources of the misallocation, such as size-dependent taxes and subsidies, selective enforcement of regulations, financial constraints, internal trade costs, and poor enforcement of contracts, among others. This approach has been widely used in the misallocation literature, for example, by \citet{HsiehKlenow:2009, HsiehKlenow:2014}, \citet{Bartelsmanetal:2013}, and \citet{AdamopoulosRestuccia:2014}, \citet{BentoRestuccia:2016} among others.}

%Moreover, we find that in China, the average product of labor and capital increase strongly with firm size and productivity [\citet{HsiehKlenow:2009, HsiehKlenow:2014}], suggesting the existence of size-dependent distortions.
% Specifically, if we take the U.S. economy as a relatively distortion-free benchmark, Fact 2 suggests that there could exist size-dependent distortions that reallocate factors away from large productive firms China, which we indeed find from the positive correlation between average factor product and firm level productivity [\citet{HsiehKlenow:2009, HsiehKlenow:2014}].\footnote{This measure of distortion captures the effects of many of the factors that distort factor allocation across firms, including taxes and subsidies, regulations, local protectionism, transportation costs, and financial frictions. A prominent example of such size-dependent distortions is internal trade barriers that impede the flow of goods across regions, which presumably affect large firms more since these are more likely to sell goods across regions [\citet{Young:2000}, \citet{TombeZhu:2015}, and \citet{Eberhardtetal:2016}]. In addition, as is the case in many other developing countries, small firms are more likely to evade taxes because of imperfect tax enforcement.} The literature has found that size-dependent distortions limit the operations of large productive firms and allow too many small unproductive firms to survive, which lowers both mean firm size and aggregate output.\footnote{See for instance \citet{Guneretal:2008}, \citet{Bartelsmanetal:2013}, \citet{HsiehKlenow:2014}, and \citet{BentoRestuccia:2016} among others.}
% An additional insight that we provide in this study is that size-dependent distortions also increase aggregate pollution intensity for two reasons. First, they lower the profits of firms which makes clean technology investment unaffordable to many firms. Second, they reduce the factors allocated to large productive firms which are more likely to use advanced pollution treatment technologies, and therefore reduce the share of total output that is produced using clean technologies. However, since size-dependent distortions both increase aggregate pollution intensity and reduce aggregate output, it is a quantitative question whether they also amplify aggregate pollution.

% To guide our analysis, we use a quantitative model to organize our empirical findings as well as quantifying the effects of size-dependent distortions on aggregate output and pollution. In particular, we extend the classic \citet{Lucas:1978b} span-of-control model to include a distinction between polluting and non-polluting sector, size-dependent distortions, an endogenous choice of pollution treatment technologies, and imperfect environmental regulations. In our model, there is a stand-in household with a continuum of members. Household members are endowed with different sectoral specific managerial talents and make occupational choices based on their talents. If a member chooses to be an entrepreneur, s/he uses her managerial talent (which becomes the firm-level productivity), capital, and labor to produce the output in the sector where the managerial talent belongs. S/he also has to pay implicit taxes that depend on her talent, which aims to capture size-dependent distortions in an intuitive way. Firm size is therefore determined by the managerial talent and distortions jointly.

% A novel feature of our model is that firms in the polluting sector also make decisions on pollutants treatment technologies. The installation of clean technology incurs a fixed cost and we assume that firms without clean technology will lose a fraction of their profits, capturing various regulatory costs that firms using dirty technology face parsimoniously. The fixed costs associated with the installation of clean technology lead to increasing returns to scale, which implies that only firms above a certain size threshold choose to install the clean technology in our model. For a given distribution of managerial talents and distortions, the model generates endogenously a distribution of firm sizes and a negative association between firm size and pollution intensity. To discipline our quantitative analysis, we require our benchmark model to match the observed pollution intensity, clean technology adoption rate, size-dependent distortions, and firm size distribution in China. %Distortions in the model is measured using variations in average products of capital and labor across firms following \citet{HsiehKlenow:2009, HsiehKlenow:2014}.

Our quantitative results show that removing correlated distortions from both sectors would increase mean firm size by 130\% and aggregate output by 30\%, while lowering the pollution intensity of the polluting sector by 42\%. Since both sectors share similar distributions of firm TFP and distortions in the data, eliminating correlated distortions induces little reallocation across sectors. Therefore, aggregate pollution drops by 23\% because the technique effect dominates the scale effect. About half of the technique effect comes from the reallocation of factors across firms, while the rest is from a 27 percentage points increase in the adoption rate of clean technology. %Meanwhile, the proportion of firms using clean technology would increase by 27\%.% The drop in pollution comes from both the reduction in pollutants generated during the production stage and the increase in the adoption rate of clean technologies in the treatment stage. Each stage contributes to about 50\% of the total reduction.%\footnote{As noted earlier in this section, firms can take actions to cut their emissions levels in two stages. In the model, we focus on firms' choice of end-of-pipe treatment technologies, which we can directly observe and measure in the data. We capture the decrease in pollution intensity during the production stage in a reduced-form way, which is calibrated to the data.}
%The expansion of productive firms is crucial to both channels.
To isolate the importance of the \textit{progressiveness} of distortions, we solve a version of the model in which the progressive taxes are replaced with flat taxes while holding the total tax revenue constant. Our results show that while the progressiveness of distortions does not imply large output losses (7\%), it plays a central role in amplifying aggregate pollution (30\%).
% where all firms in the economy face the same level of distortions. In our model, the size dependency of the distortions does not imply large output loss.  However, it plays a central role in determining the pollution level. Removing the distortions altogether provide a useful upper bound to gauge the maximum gains that can be obtained by moving the economy to its first best.

In the second experiment, we increase the regulations to the extent that the fraction of firms adopting clean technology is the same as in the first experiment. We find that the policy reduces aggregate pollution by 14\%, but has very limited effect on aggregate output in the long run. %\footnote{A complete welfare analysis of tightening environmental regulations is beyond the scope of this paper, since we do not have enough data to estimate the increase in government expenditures when environmental regulations are tightened. We leave this important question to future research.} %
Comparing this with the first experiment, we find that tightening environmental regulations {improves} the resource allocation on the \textit{extensive} margin of resource allocation by driving small unproductive firms out of the economy. However, the allocation {worsens} on the \textit{intensive} margin in the sense that among the remaining active firms, the production of medium-sized firms expands more at the expense of that of large firms. %In addition, environmental regulation also amplifies the effects of size-dependent distortions by increasing the average size of active firms and consequently the average level of distortions in the economy.
In contrast, the removal of correlated distortions improves the allocation on both margins.

Overall, the key message from our findings is that policy distortions and market inefficiencies force developing countries such as China to operate within the production possibility frontier between economic output and environmental quality. Therefore, contrary to a popular belief among some policymakers in developing countries, there is not necessarily a trade-off between economic growth and pollution. Both an increase in output and a reduction in pollution can be attained by reducing policy distortions and market inefficiencies.

%Our findings have novel implications for the relationship between economic growth and pollution, both for the case of China and for developing countries in general. The environmental Kuznets curve (EKC henceforth) predicts an inverse U-shaped relationship between pollution and income per person, which implies a trade-off between economic growth and pollution when the income per person is low. There is a popular belief that China and other developing countries on the increasing part of the EKC and any policies that promote economic growth would necessarily cause more pollution. Our results suggest that a double-dividend where increase in the output and reduction in pollution are achieved simultaneously [\citet{Goulder:1994}] can be attained with policies that target at reducing the size-dependent economic frictions. Therefore, our paper sheds light on how environmental pollution can be controlled without damaging economic growth in developing countries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                       STARTING LITERATURE REVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textit{Related Literature.}---This paper is closely related to studies on the aggregate consequences of misallocation across heterogeneous firms. Pioneers in this field such as \citet{Guneretal:2008}, \citet{RestucciaRogerson:2008}, and \citet{HsiehKlenow:2009} establish that firm-level distortions can potentially cause large losses to aggregate output, especially those that correlate with firm size or productivity.\footnote{For more recent works, see \citet{Bartelsmanetal:2013}, \citet{Bhattacharyaetal:2013}, \citet{HsiehKlenow:2014}, \citet{BentoRestuccia:2016}, and \citet{Guneretal:2017}, among others. \citet{RestucciaRogerson:2013}, \citet{Hopenhayn:2014b}, and \citet{RestucciaRogerson:2017} summarize important contributions to the literature.} Our study shows that the impact of firm-level distortions goes beyond aggregate economic output: they cause not only a large decrease in aggregate output, but also a large increase in aggregate pollution. By considering both aggregate output and pollution, our study thus provides a more complete understanding of the consequences of policy distortions given the large costs of pollution in developing countries such as China.%\footnote{\citet{GreenstoneJack:2015} summarize recent studies on the pollution problems of developing countries, and document poor environmental quality and high health and productivity costs of pollution in these countries.}
%We view our approach as complementary to that work. In this paper, we build a macroeconomic model with rich quantitative implications which facilitates the investigation of policy related questions. Our focus on the firm size distribution and treatment technology adoption also distinguishes our paper from that literature, which studies the roles of product mix, consumer preference, etc. In connection with literature focusing on the cross-country comparison of firm size distributions [for example \citet{Poschke:2014}], our paper is also a candidate of structural interpretations of the EKC.

% We contribute to a large strand of the literature on the relationship between economic growth and environment. At the heart of the literature is the environmental Kuznets curve. \citet{GrossmanKrueger:1993, GrossmanKrueger:1995}, followed by many others, establish empirically a hump-shaped relationship between a country's per-capita income and its environmental quality.\footnote{\citet{CopelandTaylor:2004} provides a review on this literature.} %\footnote{\citet{CopelandTaylor:2004} identify three channels through which economic growth can affect environmental outcomes: scale (total production), technology, and industry composition. However, our study contributes to the literature by emphasizing the importance of a new channel, namely factor allocation across firms, and its interaction with the scale and technology channels. Further, in contrast to most studies of this topic, pollution treatment technology is endogenous in our study and we use the model and data to quantify the role of the adoption of such endogenous technology.} Many scholars have interpreted this fact as evidence of an unavoidable trade-off between economic growth and environmental improvement when per-capita income is low, and there is a popular belief that growth-promoting policies necessarily lower environmental quality in China and other developing countries. We challenge this view by showing that removing distortions leads to both higher aggregate output and lower aggregate pollution in China. %\footnote{Several recent studies also examine the extent to which factor allocation across firms affects aggregate pollution. By using Indian manufacturing data, \citet{BarrowsOllivier:2016} find that the reallocation of resources across firms explains more than 50\% of the decline in aggregate $\text{CO}_2$ emissions intensity in India between 1990 and 2010. Similarly, by using U.S. manufacturing data, \citet{ShapiroWalker:2015} find that emissions reductions in the United States are primarily driven by the decline in within-product emissions intensity. They further use a quantitative model to show that the increasing stringency of environmental regulations explains most of the emissions reduction. \citet{Martin:2013} estimates empirically the relative contribution of changes in different policies for India. Our work complements these studies in that we emphasize the importance of size-dependent distortions, and our data allow us to speak directly of quantifying the importance of adopting endogenous treatment technology.} The broad message from our findings is that policy distortions and market inefficiencies typically force developing economies to operate within the production possibility frontier between economic output and environmental quality. Therefore, both an increase in output and a reduction in pollution can be attained by reducing policy distortions and market inefficiencies.%\footnote{There is a large literature in environmental economics that discusses the possibility of a double dividend from environmental taxes, namely a simultaneous improvement in the environment and in economic efficiency by replacing other distortive taxes with environmental taxes. Our finding that environmental regulations improve resource allocation along the extensive margin is consistent with this literature. \citet{Goulder:1994} and \citet{FullertonMetcalf:1997} review the early work on this question.}

%The present study also provides a novel mechanism that could explain the declining part of the environmental Kuznets curve. One of its key findings is that large firms have lower pollution intensity. Recent empirical studies such as \citet{Poschke:2014} and \citet{BentoRestuccia:2016} find that mean firm size in the manufacturing sector increases with a country's per-capita income, with \citet{HsiehKlenow:2014} and \citet{BentoRestuccia:2016} attributing the cross-country differences in firm size largely to the cross-country differences in distortions. These two facts together imply that the decrease in distortions would contribute to the decline in aggregate pollution as per capita income increases.

This paper also contributes to a growing literature in environmental economics that emphasizes the importance of firm-level heterogeneity. One line of research uses the heterogeneous firms framework to evaluate specific environmental policies. \citet{LiSun:2015} and \citet{LiShi:2017} compare the welfare effects of emission taxes and standards. Their theoretical findings imply that emission taxes will generally be favored when the dispersion of productivity across firms is large. Relatedly, \citet{TombeWinter:2015} use a heterogeneous firms model calibrated to the U.S. to evaluate the productivity losses from output-based intensity standards. They find that these losses can be large. We differ from this line of research by our focus on how correlated distortions affect aggregate pollution and how they interact with environmental policies.
%our emphasis on the tight connection between our theory and firm-level data.

Another line of research focuses on how international trade affects the environment through resource reallocation across heterogeneous firms. \citet{Cherniwchanetal:2017} provide a comprehensive survey of recent contributions.\footnote{Notable examples include \citet{Martin:2013}, \citet{Cherniwchan:2017}, \citet{BarrowsOllivier:2016}, \citet{Forslidetal:2015}, and \citet{ShapiroWalker:2015}. A more detailed comparison between our paper and this line of research can be found in Online Appendix H.} Our paper differs from this literature in two aspects. First, most studies in this literature, with the exception of \citet{ShapiroWalker:2015}, are empirical where models are used primarily to motivate the empirical specification indirectly. Instead, we discipline our general equilibrium model using firm-level data directly in an internally consistent way, and quantitatively assess counterfactual policies using the model. Our quantitative analysis explicitly takes into account the general equilibrium effects of alternative policies, allowing us to investigate novel policy implications that are absent from reduced-form or partial equilibrium analyses. Second, using the firm-level data on treatment technologies, we provide an explanation for the negative correlation between firm size and pollution intensity. It allows us to not only investigate the overall effects of distortions, but also quantify an important channel through which the distortions affect emissions, namely, the adoption of clean technology.
%Recently, a growing literature studies how international trade affects industrial pollution through the reallocation of resources across heterogeneous firms; see \citet{Cherniwchanetal:2017} for a comprehensive survey. Our paper differs from this closely related literature in several ways.\footnote{A more detailed review of the difference between our paper and this literature is deferred to Online Appendix H.} First, while most studies in this literature use either reduced-form or purely theoretical analysis, with the exception of \citet{ShapiroWalker:2015}, we complement these studies by providing a quantitative assessment of the general equilibrium effects of alternative policies on output and emissions, using an internally-consistent model disciplined by firm-level data.\footnote{Some studies in the literature, for example, \citet{Forslidetal:2015}, prove some qualitative results in general equilibrium under certain assumptions on model parameters, while they do not take the model to data directly.} The quantitative nature makes our analysis suitable for evaluating alternative policies. Second, the information on treatment equipment used by firms allows us to provide an explanation for the negative correlation between firm size and pollution intensity, and to investigate not only the overall effects of distortions, but also an important channel through which the distortions affect emissions, namely the adoption of clean technology.\footnote{In this regard, our work shares similar spirits with empirical studies like \citet{BarrowsOllivier:2016} that study how trade liberalization affect carbon emissions through changes in the product-mix of firms, and \citet{Bustos:2011} that study how trade liberalization improves firm productivity by inducing more investment in technology.} Third, while the literature focuses on trade costs, we focus on domestic distortions, and these two types of frictions interact with environmental regulations in different ways. In our model, environmental regulations would worsen resource allocation along the intensive margin, but this may not be the case with trade costs, because the (more productive) exporting firms can use foreign markets as a cushion to buffer the burdens from environmental regulations after paying the trade costs.

%Our paper differ from a recent literature in international trade surveyed in \citet{Cherniwchanetal:2017} that also investigate how resource allocation across heterogenous firms affects industrial pollution in several important ways.\footnote{In the interests of space, we leave a more extensive review of the difference between our paper and this literature to the Online Appendix I.} First, the theoretical implications of our model on resource allocation along the intensive margin under the policy-relevant scenario where environmental policies are changed while keeping trade costs (a form of distortion) fixed are different from those of the models used in these studies. While our model implies that in these situations resource allocation along the intensive margin worsens, in those studies it is improved because exporting firms can use foreign markets as a cushion to buffer the increase in costs.\footnote{Loosely speaking, the counterpart of an increase in environmental regulation could be interpreted as an increase of the fixed entry cost $f$ in the original \citet{Melitz:2003} model. An increase in $f$ perturbs the optimal decisions of all firms by affecting the aggregate price index $P$ through general equilibrium effect. Since the revenues of exporting firms come from both domestic sources that are affected, and foreign sources that are not, they are less affected by the increase in $f$.} Such difference of our model offers a novel policy message compared to these papers and shows the conceptual difference of domestic distortion and trade costs.\footnote{The difference is not caused by our use of a \citet{Lucas:1978b} model instead of a \citet{Melitz:2003} model. Rather, it is because distortions in our model is a pure cost, while in the \citet{Melitz:2003} model, firms pay trade costs to access to foreign markets, which bring additional revenues. If instead we model in the \citet{Melitz:2003} model similar ``deadweight-loss'' distortions, we would get similar results.} Second, the quantitative nature of our study allows us to analyze in general equilibrium how output and emissions respond to distortions and environmental policies simultaneously, making it suitable to evaluate quantitatively scenarios such as whether a decrease in pollution intensity coupled with a larger increase in output would lead to an increase in aggregate pollution. The local nature of the empirical estimates in these studies are not designed to answer questions like this.\footnote{Some of these studies prove some qualitative results in general equilibrium under certain assumptions on model parameters, however they did not take the model to data directly on this respect [\citet{Forslidetal:2015}].} Our paper demonstrates the potential and the importance of studying environmental issues quantitatively. From a methodological perspective, our work is similar to that of \citet{ShapiroWalker:2015}. Third, the fact that we observe directly the emissions and treatment equipments of firms in the data allows us to investigate how distortions affect emission through a specific channel, that is the adoption of clean technology, instead of an overall effect of distortions (trade costs to be specific) to emissions [\citet{Martin:2013}, \citet{Forslidetal:2015}, and \citet{ShapiroWalker:2015}]. In this regard, our work shares similar spirits with empirical studies like \citet{BarrowsOllivier:2016} that study how trade liberalization affect carbon emissions through changes in the product-mix of firms, and \citet{Bustos:2011} that study how trade liberalization improves firm productivity by inducing more investment in technology.

Finally, our results highlight the role of firm size and misallocation for technology adoption in developing countries. As emphasized by \citet{ParentePrescott:1994, ParentePrescott:1999}, an important task in development economics is to understand the slow rate of technology adoption in developing countries. Recent contributions include \citet{Bustos:2011}, who study the role of tariff reductions, and \citet{Acemogluetal:2012}, who focus on taxes and subsidies. %and \citet{ColeGreenwoodSanchez:2016}, who emphasize the importance of contractual frictions and financial market inefficiencies.
Using direct observations on the adoption of pollution treatment technologies by firms, we establish empirically that firm size plays a key role in technology adoption. Our quantitative framework thus allows us to quantify how policy distortions and market inefficiencies that distort firm size would impede technology adoption. %If we view tariffs as a particular type of policy distortion, our results share a similar increasing returns to scale intuition with that of \citet{Bustos:2011}, who shows empirically that a bilateral reduction in tariffs induces more firms to adopt a new technology with a fixed cost. Our study also differs from hers in that the quantitative framework we adopt allows us to evaluate the effect of regulatory policies with the presence of distortions and analyze the economic mechanism. %This conclusion applies generally to cases in which technology adoption exhibits increasing returns to scale.\footnote{In a different setting, \citet{OlmsteadRhode:2001} study the adoption of tractors in the U.S agriculture between 1910--1960, and they also find that larger farm scale induced greater adoption of tractors.}

The remainder of the paper proceeds as follows. Section \rom{1} describes the data sources and empirical findings of this paper. Section \rom{2} introduces the model and analyzes its property. We calibrate the model in Section \rom{3}, and perform counterfactual experiments to study the effects of correlated distortions and environmental policies in Section \rom{4}. Section \rom{5} concludes the paper.

%The remainder of the paper proceeds as follows. Section \rom{1} discusses the two facts that motivate our study in details. We describe the model and use it to estimate the firm-level distortions in Section \rom{2}. We calibrate the model in Section \rom{3} and perform several policy experiments to study the interaction between size-dependent distortions and environmental policies in Section \rom{4}. Section \rom{5} concludes the paper.

\section{Empirical Analysis}

In this section, we describe our empirical findings on the firm-level pollution intensity and the firm size distribution in China. We start by introducing the data sources we use, and then document the key facts on pollution intensity, pollution treatment technologies and employment shares across firms of different sizes. In the end, using a simple accounting exercise, we show that firm size distribution plays a quantitatively important role in aggregate pollution intensity.

%In this section, we describe the key empirical findings that suggest firm size distribution as an important factor affecting aggregate pollution. We start by introducing the data and then explain the two motivating facts mentioned before in details. In the end, by using an accounting exercise, we show that firm size distribution has a sizable effect on aggregate pollution.

\subsection{Data Sources}
\label{sec:data_source}

This study draws upon three major data sources: the 2007 National General Survey of Pollution Sources (NGSPS) in China, the 2004 China National Economic Census (CNEC), and the Statistics of U.S. Businesses (SUSB).

%This study draws upon three major data sources to calculate the pollution intensity of Chinese manufacturing firms and the firm size distributions of manufacturing firms in China and the United States: China's National General Survey of Pollution Sources (NGSPS), China's National Economic Census (CNEC), and the Statistics of U.S. Businesses (SUSB).\footnote{In the interest of brevity, we defer a more detailed description of these data to Online Appendix A.}

\textit{NGSPS.}---The NGSPS was a joint effort of multiple national ministries in China.\footnote{A more detailed description of the dataset can be found in Online Appendix A.} It covers all of the pollution units from the agricultural, industrial, and domestic sources, as well as facilities for the centralized treatment of pollution. In this study, we focus on the {industrial pollution sources}, which consist of businesses generating pollutants from the 39 manufacturing industries. The survey is organized directly by the State Council with data collection and quality control done by specially trained field staffs. According to the \textit{Regulation on the National General Survey of Pollution Sources [Decree of the State Council of the People's Republic of China (No. 508)]}, faking data would lead to financial penalties or even lawsuits. Therefore, the dataset is administrative in nature and is subject to less concerns of systematic under-reporting.\footnote{In fact, uniform under-reporting of emissions by small and large firms will not change the basic conclusions of this paper. Our results will be contaminated only if under-reporting by large firms is more severe than that by small firms, which is less likely in most polluting industries since larger firms are usually inspected more carefully. For instance, many large firms in these industries are identified as \textit{Key Pollution Sources}, and therefore are inspected and monitored by the government routinely and carefully, making it harder for them to fake data during this one-time survey.} While we are not aware of other economic studies that use NGSPS, a number of studies in environmental sciences have used this dataset.\footnote{See \citet{Niuetal:2016} and \citet{Qietal:2017} and the references therein for recent examples.}

%\textit{NGSPS.}---The 2007 NGSPS was a joint effort of multiple national ministries in China. The survey records data for 2007. It is designed to cover {all} entities and self-employed households that emit pollutants in China. The complete survey consists of four components: industrial pollution sources, agricultural pollution sources, domestic pollution sources, and facilities for the centralized treatment of pollution. In this study, we focus on the {industrial pollution data}, which include all polluting production entities that belong to any of the 39 manufacturing industries. The survey is organized directly by the State Council with data collection and quality control done by specially trained field staffs. According to the \textit{Regulation on the National General Survey of Pollution Sources [Decree of the State Council of the People's Republic of China (No. 508)]}, faking data would lead to financial penalties or even lawsuits. Therefore, the dataset is administrative in its nature and is subject to less concerns of the systematic underreporting like those documented in \citet{Dufloetal:2013} and \citet{Currieetal:2015}.\footnote{In fact, universally systematic underreporting of emissions will not change the conclusion of our paper. Our results will be contaminated if underreporting among large firms are more severe than small firms, which is less likely since these firms are usually inspected more carefully. For instance, many large firms are categorized as \textit{Nationally and Provincially Monitored Key Pollution Sources} which have to report their emissions routinely, making it harder for them to fake data during the survey process.} While we are not aware of other studies in economics that use NGSPS so far besides our own, the data in fact have been used by a large number of studies in environmental sciences.\footnote{See \citet{Niuetal:2016} and \citet{Qietal:2017} and the references therein for recent examples.}

The NGSPS collects information on the emissions of multiple air, water, and solid waste pollutants. In this paper, we focus on water pollution mainly because the data are more accurately measured than air pollution.\footnote{After our interview with the field staff, we learned that the differences in data quality was mainly caused by the differences in the sampling environment. For water pollution, firms rarely discharge wastewater directly. Instead, the wastewater is first stored in a sedimentation tank to settle heavier solids to the bottom and let lighter materials to float. This process creates a quiescent environment for sampling. For air pollutants, the sampling tube has to be placed in the chimneys, where the environment is much more volatile. Moreover, if firms were to manipulate their instant emissions when the field staff is on-site sampling, the sedimentation process makes it more difficult, since firms would have to clean the tank prior to the on-site visit of the field staff.} Among all the water pollutants, we focus on \textit{chemical oxygen demand} (COD), a major water pollutant that is commonly used to measure water quality. It measures the overall quantity of contaminants that will eventually cause oxygen loss and thus the death of living creatures. We choose COD because a much larger share of firms report positive emissions of COD compared to other pollutants. If we focus on other pollutants, such as hexavalent chromium, we would limit ourselves to a much smaller share of firms that engage in special production processes like leather tanning or pigment manufacturing.\footnote{See Table A.1 in Online Appendix A for the share of firms that report positive emissions for different water pollutants. We would like to stress that many firms do not report emissions of specific pollutants such as cyanidium or chromium simply because they do not use related materials in production. It is \textit{not} because of under-reporting or other issues related to data quality. In fact, since these pollutants are more hazardous than others, firms that discharge these pollutants are usually monitored more closely.} Moreover, COD emissions are typically highly correlated with the emissions of other pollutants when both are positive.\footnote{In the Paper and Paper Products industry, for example, the correlation between the emissions of COD and $\text{NH}_4^+$ is $\text{corr}(\text{COD},\text{NH}_4^+) = \text{0.82}$ and that between COD and BOD is $\text{corr}(\text{COD},\text{BOD}) = \text{0.94}$.}
%The NGSPS contains information on the emissions of multiple air, water, and solid waste pollutants. Here, we focus on water pollution because the data are more accurately measured than air pollution.\footnote{We have interviewed the field staff responsible for data collection about this issue. The difference in data quality was mainly caused by the differences in the sampling environment. For water pollution, firms rarely discharge wastewater directly; instead, the wastewater is first stored in the sedimentation tank to settle heavier solids to the bottom and let lighter materials to float. This process creates a quiescent environment for sampling. For air pollutants, the sampling tube has to be placed in the chimneys, where the environment is much more volatile. Moreover, if firms were to manipulate their instant emissions when the field staff is on-site sampling, the sedimentation process involved with wastewater treatment makes it more difficult, since firms would have to clean the tank prior to the on-site visit of the field staff.} The variables we use are the amount of major pollutants generated and emitted, total production value, type, book value and annual operating costs of pollutant treatment equipment, firm's industry (four-digit GB/T4574-2002), ownership classification, and province. %However, we expect that the main results of this paper could be applied to other pollution sources as well. %The raw data contain 921,004 firms.

The data reveal that industrial water pollutants are typically concentrated in a handful of industries, including COD. In our empirical analysis, we focus on the top five polluting industries in terms of COD emissions, which account for nearly 80\% of all COD emissions from industrial sources. Within narrowly defined manufacturing industries, the NGSPS divides firms into two large groups---\textit{key sources} and \textit{regular sources}. Firms identified as ``key sources'' are deemed as more polluting, and therefore are inspected and monitored by the government more carefully.\footnote{See Online Appendix A.1 for more information on the definition of key sources.} During this census, the field staff also collected extra information on these firms, and the data from these firms were inspected with greater care. Importantly, the key sources include both small and large firms, and the empirical densities of the log output of the key sources resemble the bell shape of those of all firms.\footnote{Please see Figure B.1 in Online Appendix B.} As a result, when empirically analyzing the relationship of firm size with pollution intensity and treatment technology adoption, we focus on the key sources, which account for more than 90\% of total COD emissions and 80\% of total output from the top five polluting industries (Table \ref{tab:top5summary}). %\footnote{We have also run all the following regressions using a sample including both regular and key sources, and found similar negative correlations.}
The variables we use are the firm's output, the amount of major pollutants generated and emitted, installation costs of pollutant treatment technologies, industry (up to four-digit GB/T4574-2002), ownership classification, and province.

%Environmental sciences studies find that industrial waste is typically concentrated in a handful of sectors. Further, even within narrowly defined manufacturing sectors, pollutant emissions are usually concentrated among firms that engage in some particular manufacturing processes. Grouping firms by sector thus risks blurring important within sector variations. To address this issue, the NGSPS divides the sample into two large groups---\textit{key sources} and \textit{regular sources}---where loosely speaking, firms identified as ``key sources'' are the most polluting ones. We focus on key firms in the study because they produce the majority of the pollutants (Table \ref{tab:top5summary}, which is explained in details shortly after), while most regular firms emit very little pollutants. This means that key firms are more representative of polluting manufacturing firms in China.\footnote{Online Appendix A.1 defines these key sources formally.} Moreover, the key sources include large firms as well as small firms in that the empirical density function of the log output value of the key sources resembles the same hump-shape of that of all firms.\footnote{Please refer to Online Appendix B and the two upper panels of Figure B.1 therein for details.} As a result, we use the key sources to characterize the relationship between emissions and firm size more accurately.

%Among all the pollutants, we analyze \textit{Chemical Oxygen Demand} (COD), a major water pollutant commonly used to measure water quality. COD measures the amount of oxygen consumed when a chemical oxidant is added into a sample of water. It is an indirect measure of the overall quantity of contaminants that will eventually cause oxygen loss and thus the death of living creatures. We choose COD because it is widely emitted by firms across various sectors, which allows us to keep most observations from the data. Other pollutants are emitted by significantly fewer firms because they are usually associated with the production in particular industries, which raises sample selection concerns. See Table \ref{tab:nfirmPollutant}, which lists the percentage of key sources firms that have positive emissions of different pollutants. In other words, if we focus on other pollutants, for instance hexavalent chromium, we are limiting ourselves to a handful of firms that engage in particular production processes like leather tanning and pigment manufacturing.\footnote{We would like to stress that the fact many firms do not report emissions on specific pollutants like cyanidium or chromium is simply because their production process does not involve the use of related materials. It is \textit{not} because of under-reporting or other issues related to data quality. In fact, since these pollutants are more hazardous than others, firms that discharge these pollutants are monitored more closely.} Moreover, COD emissions are highly correlated with the emissions of other pollutants.\footnote{In the Paper and Paper Products industry, for example, the correlation between the emissions of COD and those of $\text{NH}_4^+$ is $\text{corr}(\text{COD},\text{NH}_4^+) = 0.82$ and that between COD and BOD is $\text{corr}(\text{COD},\text{BOD}) = 0.94$. This suggests that if a firm discharges a large amount of COD, it is likely that it also discharges a lot of other pollutants that are specific to its production process.} Finally, we focus on {measured} end-of-pipe emissions. Online Appendix A.2 explains in detail how these data are collected. In the interests of space, we note that the information contained is different from the mix of energy sources and intermediate goods in the production process. We present the results for the top five polluting industries (29,019 firms).
%\begin{figure}[t]
%    \begin{center}
%    \includegraphics[width=0.50\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/pie_R.pdf}
%    %\floatfoot{Source: National General Survey of Pollution Sources}
%    \caption{COD Emissions by Sectors}
%    \label{fig:pie}
%    \end{center}
%\end{figure}
\begin{table}[t]
\footnotesize
\centering
\begin{threeparttable}
\caption{Statistics of the Top 5 Polluting Industries by COD}
\begin{tabular}{lcccccc}
    \hline \hline
%    & Waste & COD & Petrochemicals & Ammonian & BOD & Cyanidum & Hex-Chromium & Phenol & Arsenium & Chromium \\
                                                &  & Paper & Food Processing& Textiles  & Chemical Materials & Beverages  \\ %& Med  & Fer  & Petro & Food & Fib \\
    \midrule
    \% in Total COD Emission                    &  & 33.4  & 15.1 & 14.0 & 10.4 & 4.27  \\ %& 2.98 & 2.49 & 2.32  & 2.30 & 2.15 \\
    \% of Emissions by Key Sources              &  & 99.6  & 91.8 & 91.1 & 99.7 & 65.1  \\ %& 92.9 & 99.9 & 99.9  & 96.4 & 97.8 \\
    \% of Production by Key Sources             &  & 87.2  & 69.3 & 48.3 & 98.6 & 88.1  \\ %&& 95.7 & 99.3 & 99.7  & 98.5 & 91.9 \\
    \bottomrule
\end{tabular}
\begin{tablenotes}
     \footnotesize{
     \item[$\dagger$]
     Data Source: NGSPS. The acronyms respectively refer to (with the two-digit GB/T4547-2002 classification code in parentheses): Paper and Paper Products (C22), Processing of Food from Agricultural Products (C13), Textiles (C17), Raw Chemical Materials and Chemical Products (C26), and Beverages (C15).}
%     \item[a]
%     Relative contribution to total COD emissions by each sector.
%     \item[b]
%     Percentage of total COD emissions accounted for by key firms within each sector.
%     \item[c]
%     Percentage of total production accounted for by key firms within each sector.
\end{tablenotes}
\label{tab:top5summary}
\end{threeparttable}
\end{table}

%Table \ref{tab:top5summary} contains the basic statistics on these industries. The key firms in the top five polluting industries are representative of industrial pollution in China. These industries combined contribute 77\% of total industrial COD emissions, with key firms responsible on average for more than 90\% of the within-sector emissions and more than 80\% of the within-sector output.

\textit{CNEC.}---The CNEC was conducted by the National Bureau of Statistics in China. It was designed to cover {all} businesses that undertake economic activities in the secondary and tertiary industries in China. We use observations that belong to the manufacturing sector. The variables we use are the firm's output, labor compensation, book value of capital stock, number of employees, industry (up to four-digit GB/T4574-2002), ownership classification, and province.\footnote{A widely used firm-level dataset in China is the Annual Surveys of Industrial Firm (ASIF). We use CNEC instead of ASIF because that ASIF does not have information on non-state firms with sales below 5 million CNY, which are crucial for our analysis. If we used the 2004 ASIP, we would have missed 29\% of the employment and 80\% of the firms.}%\footnote{It is important to use the CNEC rather than the commonly used \textit{Annual Surveys of Industrial Production (ASIP)} for which the 2007 data are available (the same year that the NGSPS covers). The reason is that the CNEC surveys firms of {all} sizes as opposed to only firms with a revenue of more than CNY 5 million in the ASIP. In 2004, the number of firms and employees covered by the annual survey were 276,410 and 66,725,059, respectively, while those covered by the census were 1,375,148 and 93,541,923. Therefore, we would be missing 28.6\% of employment and 79.9\% of firms had we used only the annual survey. %The number of firms covered in the NGSPS and CNEC, 921,004 and 1,375,148, are broadly consistent given that the former further requires a production entity to have pollution sources in order to be included.Online Appendix B provides more details. However, the basic features such as the variable definitions are essentially the same in these two datasets. \citet{Brandtetal:2012} describe the annual surveys in detail.}

\textit{SUSB.}---The SUSB from the U.S. Census Bureau, is an annual series that provides national and subnational data on the distribution of economic activities by firm size and industry. We use the data on total employment by firm size and industry from 2004.

\subsection{Firm Size and Pollution Intensity}

\begin{figure}[t]
    \begin{center}
    \includegraphics[width=0.40\textwidth]{./Figures/pooled_intensity_size.pdf}%
%    \includegraphics[width=0.40\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/agri_R.pdf} \\
%    \includegraphics[width=0.40\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/textile_R.pdf}
%    \includegraphics[width=0.40\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/chemmat_R.pdf} \\
%    \includegraphics[width=0.40\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/bever_R.pdf}
%    \includegraphics[width=0.40\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/nitro_R.pdf} \\
%    \vspace{-1em}
    \caption{Firm Size and Pollution Intensity}
    \floatfoot{Source: NGSPS. Line: Least square fit. Residuals after controlling for the covariates are plotted.}
    \label{fig:codscatter}
    \end{center}
\end{figure}

\textit{Firm Size and Pollution Intensity.---}To examine the statistical relationship between pollution intensity and firm size for the polluting sectors in China, we run the following regression using data on firms identified as ``key sources'' in the top five polluting industries:
\begin{equation}
%    \log (\text{COD}_i) = \underset{(0.17)}{-3.75} + \underset{(0.01)}{0.63} \times \log (\text{Output}_i) + \vec{X_s} \gvec{\upgamma}_1 + \vec{X_p} \gvec{\upgamma}_2 + \vec{X_o} \gvec{\upgamma}_3 + \varepsilon_i,
    \log (\text{COD}_i) = \underset{(0.17)}{-3.75} + \underset{(0.01)}{0.63} \times \log (\text{Output}_i) + \vec{X} \gvec{\upgamma} + \varepsilon_i,
\label{eq:poolols}
\end{equation}
where $\vec{X} = \{\vec{X_s}, \vec{X_p}, \vec{X_o}\}$ refers to the dummy variables for two-digit industry, province, and ownership respectively. The estimates are all statistically significant at the 0.1\% level, with the standard errors reported in the parentheses below the estimates.

These estimates imply that as a firm's output increases by 1\%, its COD emissions would on average increase by only 0.63\%. We define pollution intensity as COD emissions per unit of output. Then the above estimates imply an elasticity of $-$0.37 between firm-level pollution intensity and output. That is, other things equal, a 1\% increase in firm output is associated with a 0.37\% decrease in pollution intensity. %\footnote{Notice that regression (\ref{eq:poolols}) is mathematically equivalent to $\log(\text{COD}_i/\text{Sales}_i)= \beta_0 + \beta_1 \log(\text{Sales}_i) + \gvec{\upgamma}\vec{X} + \varepsilon_i$.}
Figure \ref{fig:codscatter} visualizes this negative relationship, where the residuals from regressing intensity and output on the dummies are plotted to remove the influences of other covariates.\footnote{Recall that here by the \textit{Regression Anatomy Formula}, the slope of the least square fit in Figure \ref{fig:codscatter} is precisely one minus the coefficient on $\log(\text{Output})$ in Equation (\ref{eq:poolols}).} The estimation has an $R^2$ of 0.47, suggesting that nearly half of the variations in COD emissions can be explained by the variations in firm output and in the three sets of dummies.\footnote{We also tried other econometric specifications, and found similar negative correlations between pollution intensity and firm size. For instance, we estimated versions of Equation (\ref{eq:poolols}) for each industry and with robust standard errors clustered in provinces. Please see Online Appendix C for more information.} We would also like to point out that the coefficient of the dummy on whether a firm is state-owned or not, is statistically \textit{insignificant}. %\footnote{This finding is different from that of \citet{Jiangetal:2014}. We think that this could be because we are using a different sample of firms. Specifically, while their data cover the Nationally Monitored Key Pollution Sources, ours have a much broader coverage.}
As a result, we do not distinguish between whether a firm is state-owned or not in our later analysis.

%the regression results indicate that pollution intensity decreases as firm sales rise. Specifically, by subtracting $\log (\text{Sales}_i)$ on both sides of Equation (\ref{eq:poolols}), the elasticity between pollution intensity and total sales is $-$0.37, which means that other things being equal, doubling total sales is associated with a 37\% decrease in pollution intensity.\footnote{Notice that regression (\ref{eq:poolols}) is mathematically equivalent to $\log(\text{COD}_i/\text{Sales}_i)= \beta_0 + \beta_1 \log(\text{Sales}_i) + \gvec{\upgamma}\vec{X} + \varepsilon_i$.} Figure \ref{fig:codscatter} shows this negative relationship, where the residuals from regressing intensity and sales on the dummies are plotted to remove the influences from other control variables.\footnote{Recall that here by the \textit{Regression Anatomy Formula}, the slope of the least square fit in Figure \ref{fig:codscatter} is precisely one minus the coefficient on $\log(\text{Sales})$ in Equation (\ref{eq:poolols}).} The estimation has an $R^2$ of 0.47, which suggests that a fair amount of variation can be explained by the variations in total sales and in the three sets of dummies.\footnote{We also estimated the relationship by using other econometric specifications. For instance, we estimated versions of Equation (\ref{eq:poolols}) for each industry and with robust standard errors clustered in different groups. All the regressions suggest the same negative qualitative relationship between intensity and size. The estimation results of the other specifications as well as the interpretations of the coefficients before the dummies are included in Online Appendix C.} It is important to emphasize that the coefficients of the ownership dummies ($\gvec{\upgamma}_3$)---that is whether a firm is state-owned, collectively owned, etc---are statistically \textit{insignificant}. %\footnote{This finding is different from that of \citet{Jiangetal:2014}. We think that this could be because we are using a different sample of firms. Specifically, while their data cover the Nationally Monitored Key Pollution Sources, ours have a much broader coverage.} As a result, in the remaining of the paper, we do not distinguish between whether a firm is state-owned or privately owned.

%We define pollution intensity as follows:
%\begin{equation*}
%    \text{Intensity} = \frac{\text{Total COD Emissions}}{\text{Total Production value}}.
%\end{equation*}
%We group firms into quartiles based on their total output value. For each industry, we calculate the output-weighted average of the pollution intensity of the firms in each quartile. Table [] reports the results. For the Paper and Paper Products industry, the pollution intensity of the firms in the bottom quartile is 6.7 times that of the firms in the top quartile. The difference can be as large as 31.4 times, as is the case for the Beverages industry. Moreover, pollution intensity decreases continuously as firm size rises. This trend can also be seen from the scatterplot of the logarithm of intensity against that of total production value. We plot the Paper and Paper Products industry in Figure \ref{fig:codscatter} as an example. It shows a significant negative correlation between log-intensity and log-production in the data. The scatterplots for the other four industries are deferred to Online Appendix C.%for the top-5 polluting industries are presented  (except the bottom-right one).

%\begin{table}[t]
%\footnotesize
%\centering
%\begin{threeparttable}
%\caption{Firm Size and Pollution Intensity}
%\begin{tabular}{p{120pt}>{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt}}
%    \hline \hline
%        & \multicolumn{4}{c}{Quartile of Firm Sales} \\
%    \cmidrule{2-5}
%    Industry & $\text{QU}_1$ & $\text{QU}_2$ & $\text{QU}_3$ & $\text{QU}_4$ \\
%    \midrule
%    Paper               & 6.7   & 3.2   & 2.0 & 1.0 \\
%    Agricultural Food   & 20.8  & 7.6   & 3.6 & 1.0 \\
%    Textiles            & 8.3   & 3.6   & 2.4 & 1.0 \\
%    Chemical Materials  & 6.7   & 3.8   & 2.7 & 1.0 \\
%    Beverages           & 31.4  & 18.7  & 4.7 & 1.0 \\
%    \bottomrule
%\end{tabular}
%\begin{tablenotes}
%     \footnotesize{
%     \item[$\dagger$]
%     Data Source: NGSPS. $\text{QU}_1$ to $\text{QU}_4$ represent the bottom to top quartiles, respectively. The pollution intensity of the top quartile of each industry is normalized to one.}
%\end{tablenotes}
%\label{tab:intensityquartile}
%\end{threeparttable}
%\end{table}
%\footnote{One exception is Chemistry Materials, the middle right panel. Further investigation into the data reveals that the reason that we do not observe a negative correlation in the Chemistry Material industry as clear as in the other industries is because there are many sub-industry level heterogeneities pooled at the two-digit level. To show that this is indeed the case, we repeat the exercise for the top polluting four-digit industry within the Chemistry Material industry, {Nitrogen Fertilizer} (C2621) which ranks the sixth among all four-digit industries. The result is shown in the lower right panel of Figure \ref{fig:codscatter}, where the negative correlation is clear. Put differently, the ``circle'' we see in the Chemistry case is in fact a combination of several ``ellipses'', each with a downward sloping transverse diameter. We have also plotted these figures using the residuals from regressions (\ref{eq:indols}) and (\ref{eq:poolols}). The results are almost the same and are thus shown Online Appendix. This is confirmed by including four-digit industry dummy in the regression for Chemistry industry specified below. After introducing four-digit industry dummies, the magnitude estimates for $\beta_1$ increases to $-$0.21 and the $R^2$ increases to 0.26. Both numbers are now of comparable magnitudes with the other industries.}

%\begin{equation}
%    y_i = \beta_0 + \beta_1 V_i + \vec{X_p}\beta_2 + \vec{X_o}\beta_3 + \varepsilon_i
%
%\end{equation}
%We then run a pooled regression with :
%\begin{equation}
%    y_i = \beta_0 + \beta_1 V_i + \vec{X_p}\beta_2 + \vec{X_s}\beta_3 + \vec{X_o}\beta_4 + \epsilon_i
%\label{eq:poolols}
%\end{equation}

%Regression results from (\ref{eq:indols}) and (\ref{eq:poolols}) are shown in Table \ref{tab:olsresults}.
%\begin{table}[t]
%\footnotesize
%\centering
%\begin{threeparttable}
%\caption{OLS Regression Results}
%\begin{tabular}{ccccccc}
%    \hline \hline
%    Parameters & Paper           & Agri            & Textile         & Chem            & Bever           & Pooled          \\
%    \midrule
%    $\beta_0$  & $-$4.25$^{***}$ & $-$2.23$^{***}$ & $-$3.73$^{***}$ & $-$7.17$^{***}$ & $-$2.82$^{***}$ & $-$2.50$^{***}$ \\
%               & (0.56)          & (0.26)          & (0.57)          & (0.33)          & (0.38)          & (0.17)          \\
%    $\beta_1$  & $-$0.30$^{***}$ & $-$0.61$^{***}$ & $-$0.31$^{***}$ & $-$0.13$^{***}$ & $-$0.45$^{***}$ & $-$0.37$^{***}$ \\
%               & (0.01)          & (0.01)          & (0.01)          & (0.01)          & (0.02)          & (0.01)          \\
%    $R^2$      & 0.22            & 0.46            & 0.14            & 0.08            & 0.29            & 0.37            \\
%    \midrule
%    $N$        & 5626            & 6883            & 6281            & 8672            & 1557            & 29019 \\
%    \bottomrule
%\end{tabular}
%\begin{tablenotes}
%     \footnotesize{
%     \item[$\dagger$]
%     Note: All regressions include complete sets of provincial and ownership rights dummies. The pooled regression includes in addition two-digit industry dummies. OLS standard errors are reported in parentheses. Regressions with robust standard errors clustered on provinces (industry level regression) and industries (pooled regression) yield very similar results are reported only in Online Appendix for brevity. Statistical significance code: $^{***}$ $p < 0.001$.}
%\end{tablenotes}
%\label{tab:olsresults}
%\end{threeparttable}
%\end{table}
%Estimation results in Table \ref{tab:olsresults} confirm observations in Figure \ref{fig:codscatter}. The intensity-production elasticities $\beta_1$ are all accurately estimated to be negative and they are also of economic significance. The $R^2$ shows that a fair amount of variation observed in the intensity is changing with production scale. As a crude calculation, the coefficient $-$0.37 on $\log V_i$ for the pooled case implies that by doubling the production scale, the COD discharged per unit production will decrease by 37\%.\footnote{The pollution intensity also exhibits rich provincial level variations. We leave the discussion to Online Appendix.}

%The pollution intensity also exhibits rich provincial level variations. We take the regression on the paper industry for example. Using Beijing as the benchmark, of all the remaining 29 provinces and municipalities, only the coefficient of Shanghai is negative and statistically insignificant. For the remainders, all estimates positive and highly significant.\footnote{One province, Hebei is significant at 10\% level. Four others are significant at 1\% level while all others are significant at 0.1\% level. We notice here Hebei, a province that is seriously disturbed by industrial pollution, is in fact an outlier as the paper manufacturing is not its pillar industry.} The bottom-five provinces (those with largest dummy estimates) are Ningxia, Inner Mongolia, Gansu, Shaanxi and Xinjiang, for which the rankings of their provincial GDP in 2007 are respectively 29, 16, 27, 19 and 25 out of 31 provincial administrative regions (not including Hong Kong and Macau). Geographically, these are also inland provinces. In fact, despite that the majority of the paper manufacturers belong to the Zhejiang and Guangdong (GDP ranking 1st and 4th), the pollution problem is of less severity as opposed to those inland less developed regions. These patterns are highly robust across various versions of our regressions. The general message we get from these dummy coefficients are the following. First, Beijing and Shanghai, arguably the political and economic center of China, exert pollution control standard far more stringent than the rest provinces. Second, inland and less developed provinces in general have worse control over and are more prone to the industrial pollution problem. These two patterns are broadly consistent with the EKC and results from previous literature. However, unlike \citet{Jiangetal:2014}, ownership rights do not seem to have significant effect. If anything, the Foreign and Hong Kong, Macau and Taiwan based firms seem to have weakly better pollution controls. But the statistical significance of the estimates never exceed 10\% across all our regressions. We think this could well because their estimates are derived using a limited sample of firms (1931) while our data cover the universe of firms.

%\subsection{Firm Size and Treatment Technologies}

\textit{The Role of End-of-Pipe Treatment Technologies.---}An important question is, why do larger firms have lower pollution intensity? To answer this question, we exploit the detailed information on the end-of-pipe wastewater treatment technologies that firms use in the NGSPS. The NGSPS groups wastewater treatment technologies into five categories: physical, chemical, physio-chemical, biological, and {combined} technologies. In the subsequent analysis, we drop {physio-chemical} technologies because fewer than 0.5\% of the firms adopt this type of technologies. In addition, because {combined} technologies look very similar to biological technologies for all the technical aspects we are interested in, we group them with {biological} technologies.\footnote{Examples of the technologies attributed to these three base categories include filtering, centrifuging, precipitation, and separation (physical); oxidation reduction and neutralization (chemical); and aerobic biological treatment and activated sludge processes (biological). Broadly speaking, physical technology corresponds to the \textit{Preliminary} and \textit{Primary} treatment in the U.S., while chemical and biological technologies correspond to the \textit{Secondary} and \textit{Tertiary} treatment.} In sum, we are left with three broad categories of technologies.%Among all the technical aspects, we are particularly interested in the abatement efficiency and fixed installation costs of these technologies.%\footnote{We do not include the annual {operating costs} in our analysis because we do not model how much a firm uses its abatement equipment. Adding the operating costs to the installation costs does not change the results in this section. Hence we exclude the operating costs here for consistency. See Online Appendix D.2 for more details.}

%\textit{The Role of Treatment Technologies.---}The negative firm size--intensity relationship does not explain why larger firms pollute with less intensity. To answer this question, we exploit the detailed information in the NGSPS on the end-of-pipe wastewater treatment equipment that firms use. The NGSPS groups wastewater treatment technologies into five categories: physical, chemical, physio-chemical, biological, and combined technologies. In the subsequent analysis, we drop {physio-chemical} technologies because fewer than 0.5\% firms adopt this type of equipment. Further, because {combined} technologies are different combinations of biological technologies with other technologies, and for all the technical aspects we are interested in they look very similar to biological technologies, we group them with {biological} technologies.\footnote{Examples of the technologies attributed to these three base categories include filtering, centrifuging, precipitation, and separation (physical); oxidation reduction and neutralization (chemical); and aerobic biological treatment and activated sludge processes (biological). Broadly speaking, physical treatment corresponds to the \textit{Preliminary} and \textit{Primary} treatment in the United States, while chemical and biological treatment relates to the \textit{Secondary} and \textit{Tertiary} treatment and those that are more advanced.} We are interested in the processing efficiency and installation costs of these technologies.\footnote{We do not include the annual {operating costs} in our analysis because we do not model how much a firm uses its abatement equipment. Adding the operating costs to the installation costs does not change the results in this section. Hence we exclude the operating costs here for consistency. See Online Appendix D.2 for more details.}

%We do not include the annual {operating costs} in our analysis because, on average, the ratio of operating costs of the treatment equipment to annual production value is about 1.5\%. Furthermore, the median of this ratio is less than 0.5\%, suggesting that operating costs are almost negligible for more than 50\% of firms. Therefore, operating costs alone are unlikely to affect a firm's decision to adopt treatment technology. Actually, adding the operating costs into the installation costs does not change the results in this section. In practice, operating costs could affect how much a firm uses the abatement equipment though. Since we do not model this intensive margin in our theory, we exclude the operating costs here for consistency. See Online Appendix D.1 for more details.}

To illustrate the difference in technology adoption by firms of different sizes, we use the Paper and Paper Products industry as an example.\footnote{We focus on one industry because of the potential heterogeneity in production processes across industries.} Table \ref{tab:techeffi} shows the mean abatement efficiency, median installation costs, and median output of firms that adopt these technologies, where the {abatement efficiency} is defined as one minus the ratio of {emitted COD} to {generated COD}.\footnote{Online Appendix D.2 provides more details.}
%To show the difference in the technologies adopted by firms of different size, we use the Paper and Paper Products industry as an example. Here we use one industry as an example because we would like to control for the potential heterogeneities in production processes across industries. Table \ref{tab:techeffi} shows the mean processing efficiency, median installation costs, and median sales of firms that adopt these technologies. We proximate for {processing efficiency} by using one minus the ratio of {emitted COD} to {generated COD}.\footnote{Online Appendix D.2 presents more details. Also we would like to stress again that our data contain both emitted and generated quantity of pollutants, see Online Appendix A.1 for details.} We normalize both the median installation costs of physical technologies and the median sales of firms using physical technologies to 100.
\begin{table}[t]
\footnotesize
\centering
\begin{threeparttable}
\caption{Firm Size and Treatment Technologies}
\begin{tabular}{p{50pt} >{\centering\arraybackslash}p{70pt} >{\centering\arraybackslash}p{60pt} >{\centering\arraybackslash}p{60pt} >{\centering\arraybackslash}p{70pt}} %>{\centering\arraybackslash}p{80pt}}%{lcccc}{p{45pt} >{\centering\arraybackslash}p{40pt} >{\centering\arraybackslash}p{40pt} >{\centering\arraybackslash}p{40pt} >{\centering\arraybackslash}p{40pt} >{\centering\arraybackslash}p{80pt}}
%{>{\centering\arraybackslash}p{45pt}>{\centering\arraybackslash}p{40pt}>{\centering\arraybackslash}p{40pt}>{\centering\arraybackslash}p{40pt}
%>{\centering\arraybackslash}p{40pt}>{\centering\arraybackslash}p{80pt}}
    \hline \hline
    Technology  & Mean Efficiency & Adoption Rate & Median Costs\tnote{a} & Median Output\tnote{b} \\
    \midrule
    Physical    & 63.37\%         & 25.89\%       & 100                   & 100                       \\
    Chemical    & 74.96\%         & 33.64\%       & 308                   & 219                       \\
    Biological  & 80.90\%         & 38.60\%       & 923                   & 500                       \\
    \bottomrule
\end{tabular}
\label{tab:techeffi}
\begin{tablenotes}
     \footnotesize{
     \item[$\dagger$]
     Data Source: NGSPS. The numbers reported are for the Paper and Paper Products (C22) industry. Abatement efficiency is defined as $1 - \text{COD Emitted}/\text{COD Generated}$.
     \item[a,b]
     The median installation costs of physical technologies and median output of firms that adopt them are normalized to 100.}
\end{tablenotes}
\end{threeparttable}
\end{table}
We find that for each 100 tons of COD generated, firms with biological technologies on average emit 19.1 tons of COD, compared to 36.6 tons for firms with physical technologies. Meanwhile, the median installation costs of biological technologies are 8 times higher than those of physical technologies. In addition, the median output of the firms using biological technologies are 4 times higher than those of firms with physical technologies. These point to increasing returns to scale in the adoption of cleaner technologies. As a result, small firms lack the profit margins needed to take advantage of these increasing returns to scale, while large firms are more likely to adopt cleaner technologies.
%We find that, on average, biological equipment is 17 percentage points more efficient than physical equipment, but its installation costs are also 11 times more expensive. Further, the sales of the firms using biological technologies are about 8 times those of firms with physical technologies. This evidence points to increasing returns to scale of clean technologies, indicating that small firms lack the profit margins needed to take advantage of the returns to scale imbedded in biological technologies, whereas large firms are more likely to adopt these more advanced technologies.

To see the statistical relationship between firm size and technology adoption, we pool all five polluting industries together and estimate a linear probability model of treatment technology adoption with the same set of covariates as in Equation (\ref{eq:poolols}):
\begin{equation*}
%     y_i = \underset{(0.04)}{0.23} + \underset{(0.001)}{0.05} \times \log (\text{Output}_i) + \vec{X_s} \gvec{\upgamma}_1 + \vec{X_p} \gvec{\upgamma}_2 + \vec{X_o} \gvec{\upgamma}_3 + \epsilon_i,
    y_i = \underset{(0.04)}{0.23} + \underset{(0.001)}{0.05} \times \log (\text{Output}_i) + \vec{X} \gvec{\upgamma} + \epsilon_i,
%\label{eq:poolols}
\end{equation*}
where $y_i$ is a dummy variable which equals to 1 when a firm uses biological technologies. The estimates show that doubling a firm's output leads to a 5 percentage points increase in the adoption rate of biological technologies. This confirms our previous finding with the paper industry that larger firms are more likely to use cleaner technologies.

%\begin{table}[t]
%\footnotesize
%\centering
%\begin{threeparttable}
%\caption{Treatment Efficiencies and Adoption Rates}
%\begin{tabular}{p{45pt} >{\centering\arraybackslash}p{40pt} >{\centering\arraybackslash}p{40pt} >{\centering\arraybackslash}p{40pt} >{\centering\arraybackslash}p{40pt} >{\centering\arraybackslash}p{80pt}}
%%{>{\centering\arraybackslash}p{45pt}>{\centering\arraybackslash}p{40pt}>{\centering\arraybackslash}p{40pt}>{\centering\arraybackslash}p{40pt}
%%>{\centering\arraybackslash}p{40pt}>{\centering\arraybackslash}p{80pt}}
%    \hline \hline
%    Technology  & 25\%      & Median    & 75\%      & Mean      & Adoption Rates \\
%    \midrule
%    Physical    & 39.54\%   & 77.81\%   & 87.83\%   & 63.37\%   & 25.79\%        \\
%    Chemical    & 74.96\%   & 81.29\%   & 86.78\%   & 69.77\%   & 34.50\%        \\
%    Biological  & 78.87\%   & 86.77\%   & 91.27\%   & 80.90\%   & 39.71\%        \\
%    \bottomrule
%\end{tabular}
%\label{tab:techeffi}
%\begin{tablenotes}
%     \footnotesize{
%     \item[$\dagger$]
%     Note: The numbers reported are for the Paper and Paper Product (C22) industry. Treatment Efficiencies is defined as $1 - \text{COD Emitted}/\text{COD Generated}$.}
%\end{tablenotes}
%\end{threeparttable}
%\end{table}
%The {designed processing capacity} (in tons) and {actual installation costs} (in 2007 CNY) can be retrieved directly from the data. We calculate the average processing cost by dividing the installation cost with capacity. In Figure \ref{fig:techdiff}, in clockwise order we plot the density functions of processing capacity, installation costs, total value of industrial output, and average processing cost by technology type. For all panels, log-scale is used in the horizontal axes.
%\begin{figure}[ht]
%    \begin{center}
%    \includegraphics[width=0.4\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/abate_cap.pdf}%
%    \includegraphics[width=0.4\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/abate_inv.pdf} \\
%    \includegraphics[width=0.4\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/abate_unit.pdf}
%    \includegraphics[width=0.4\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/abate_prod.pdf}
%    \caption{Technical Features of Different Treatment Technologies}
%    \floatfoot{Source: National General Survey of Pollution Sources. In all panels, the horizontal axes are in log-scale.}
%    \label{fig:techdiff}
%    \end{center}
%\end{figure}
%Broadly speaking, biological technologies have the {best} processing efficiency, the {largest} processing capacity, the {highest} installation costs, but the {lowest} average processing cost. Therefore, although biological technologies are more advanced in terms of processing capacity and efficiency, they are also more costly. Combined with the lower average processing cost that is displayed by the biological technologies, the evidence points to increasing returns to scale of the clean technologies. In fact, in the bottom-right panel, we find that the clean technologies are more likely to be installed by larger firms. %For example, the mean (median) processing efficiency of biological technologies is 17\% (10\%) higher than the physical technologies.
%The density functions of processing capacity, installation costs and production scale for physical, chemical and biological equipments order from left to right while those of the unit capacity costs order from right to left.
%It indicates that small firms lack the profit margins that are needed to take advantage of the increasing returns to scale exhibited by biological technologies, while at the same time, large firms are more likely to adopt these more advanced technologies. %This could also be seen by calculating the ratio of the \textit{summed} value of treatment equipments to the \textit{summed} value of production using firms of different scales. While this ratio is equal to 18\% when only firms with production level below the median are included, the number decreases dramatically to 4\% when up to 95\% firms are included which suggests strongly for the existence of fixed costs.

%\textit{Environmentally Friendly Production Technology.---}Note that the above results all relate to end-of-pipe treatment technologies; we have ignored those factors that could lead to less generation of COD. Indeed,
\textit{Emissions Generated During Production.---}Another reason behind the negative correlation between pollution intensity and firm size is that large firms also generate less COD per output during the production stage. In particular, if we divide our sample into two groups according to whether a firm is using biological technologies or not and repeat regression (\ref{eq:poolols}) again, we find significantly negative correlations between COD per output and firm size within each group:
\begin{align}
%\label{eq:poolols_physical}
%    \text{Physical:}   \quad \log (\text{COD}_i) &= \underset{(0.40)}{-3.50} + \underset{(0.01)}{0.58} \times \log (\text{Output}_i) + \vec{X_s} \gvec{\upgamma}_1 + \vec{X_p} \gvec{\upgamma}_2 + \vec{X_o} \gvec{\upgamma}_3 + %%\varepsilon_i \\
%\label{eq:poolols_chemical}
%    \text{Chemical:}   \quad \log (\text{COD}_i) &= \underset{(0.49)}{-4.59} + \underset{(0.01)}{0.77} \times \log (\text{Output}_i) + \vec{X_s} \gvec{\upgamma}_1 + \vec{X_p} \gvec{\upgamma}_2 + \vec{X_o} \gvec{\upgamma}_3 + %\varepsilon_i \\
%\label{eq:poolols_biological}
%    \text{Biological:} \quad \log (\text{COD}_i) &= \underset{(0.20)}{-4.37} + \underset{(0.01)}{0.67} \times \log (\text{Output}_i) + \vec{X_s} \gvec{\upgamma}_1 + \vec{X_p} \gvec{\upgamma}_2 + \vec{X_o} \gvec{\upgamma}_3 +\varepsilon_i, \\
%\label{eq:poolols_others}
%    \text{Others:} \quad \log (\text{COD}_i) &= \underset{(0.31)}{-3.41} + \underset{(0.01)}{0.64} \times \log (\text{Output}_i) + \vec{X_s} \gvec{\upgamma}_1 + \vec{X_p} \gvec{\upgamma}_2 + \vec{X_o} \gvec{\upgamma}_3 +\varepsilon_i,
\label{eq:poolols_biological}
    \text{Biological:} \quad \log (\text{COD}_i) &= \underset{(0.20)}{-4.37} + \underset{(0.01)}{0.67} \times \log (\text{Output}_i) + \vec{X} \gvec{\upgamma} + \varepsilon_i, \\
\label{eq:poolols_others}
    \text{Others:} \quad \log (\text{COD}_i) &= \underset{(0.31)}{-3.41} + \underset{(0.01)}{0.64} \times \log (\text{Output}_i) + \vec{X} \gvec{\upgamma} + X_e \gamma_e + \varepsilon_i,
\end{align}
where $X_e$ is the dummy for chemical equipment.\footnote{In our model and quantitative analysis below, we assume two types of treatment technologies. This is why we divide firms into two groups here. Running regression (\ref{eq:poolols}) separately for firms using physical and chemical equipments yield similarly significantly negative correlations between COD per output and firm size for each technology. Online Appendix C.3 reports the results in details.}

Such within group negative correlations could be because that larger firms use \emph{cleaner} or \emph{fewer} inputs to produce the same amount of ouput. For one example, the paper pulp industry uses two different types of inputs: bagasse and wood. While bagasse, which generates 140--180 kg COD per ton, is mostly used by firms with an annual production of {\color{black}less} than 100 kilotons, wood, which generates 30--55 kg COD per ton, is mostly used by firms with an annual production of {\color{black}more} than 100 kilotons.\footnote{See the \textit{Handbook of Emission Coefficients} published by the \textit{Chinese Academy of Sciences}.} As another example, \citet{Bloometal:2010} find that better managed firms use less energy to produce the same amount of output in the U.K. Unfortunately, we cannot test these hypotheses directly using our data. Instead, we focus on firms' decisions to adopt treatment technologies, and model the emissions reduction during the production stage in a reduced-form way based on Equations (\ref{eq:poolols_biological}) and (\ref{eq:poolols_others}).

%Such within group relationship could be because that larger firms use environmentally friendlier production technologies, sell products with a higher mark-up, or produce products that are technologically less polluting. An example from the \textit{Handbook of Emission Coefficients} published by the \textit{Chinese Academy of Sciences} is as follows. The paper pulp manufacturing use two different types of inputs: bagasse and wood. While bagasse, which generates 140--180 kg COD per ton, is mostly used by firms with an annual production of {\color{black}less} than 100 k-tons, wood, which generates 30--55 kg COD per ton, is mostly used by firms with an annual production of {\color{black}more} than 100 k-tons. Alternatively, it could also be that more productive firms simply use less resource to produce [\citet{Bloometal:2010} and \citet{ShapiroWalker:2015}]. Unfortunately, we cannot test these hypotheses directly with our data. Therefore, in this study, we focus on firms' decisions to adopt treatment equipment and model the intensity reduction during the production stage by using the reduced-form characterization of (\ref{eq:poolols_dirty}) and (\ref{eq:poolols_clean}).

\subsection{Firm Size Distribution and Aggregate Pollution Intensity}

The negative correlation between pollution intensity and firm size suggests that firm size distribution plays a potentially important role in aggregate pollution intensity. To illustrate the point, we document facts on the firm size distribution in China and in the U.S., and compute the change in the aggregate pollution intensity if the firm size distribution for the top five polluting industries in China were replaced with that in the U.S. We choose the U.S. as the benchmark because it is a relatively less distorted economy, and China and the U.S. are both large economies with a complete set of industries so that we can find comparable industries in both economies.\footnote{It is a commonly used strategy in the macro-development literature to take the U.S. economy as an undistorted benchmark. See \citet{Guneretal:2008}, \citet{HsiehKlenow:2009, HsiehKlenow:2014}, \citet{Bartelsmanetal:2013}, and \citet{AdamopoulosRestuccia:2014}, among others.}

%The negative correlation between pollution intensity and production scale implies that, {ceteris paribus}, the shares of total output by large and small firms could influence aggregate industrial pollution markedly. An important task then is to understand the firm size distribution in China. To gain some intuitions, we compare the firm size distribution in China with that in the United States. We choose the United States as the benchmark because the U.S. economy is relatively distortion-free compared with other economies, and China and the United States are both large economies with complete sets of industrial sectors. Naturally, firm size in the United States and China differs for many reasons. Yet we do not intend to answer why they are different in this paper. The primary purpose of this section is to motivate the investigation of size-dependent distortions that disproportionately affect large firms.

Since the SUSB does not report information on output, we focus on the shares of employment accounted for by firms of different sizes, since firm-level employment is strongly correlated with output. We use the \textit{International Standard Industrial Classification of All Economic Activities, Revision 3.1} (ISIC Rev. 3.1) published by the United Nations to bridge the different industrial classification systems adopted by China (GB/T4574-2002) and the United States (NAICS 2002). Specifically, we use the crosswalks of GB/2002 at the four-digit level and those of NAICS/2002 at the six-digit level to the ISIC Rev. 3.1 issued respectively by the National Bureau of Statistics of China and the U.S. Census Bureau, respectively. %The results presented in this section are from matching at the disaggregated level (four-digit GB with six-digit NAICS).\footnote{Matching at a more aggregated level (two-digit GB with three-digit NAICS) yields similar results.}

\begin{figure}[t]
    \begin{center}
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/paperES_R.pdf}%
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/agriES_R.pdf} \\
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/textileES_R.pdf}
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/chemmatES_R.pdf} \\
    \includegraphics[width=0.45\textwidth]{./Figures/poolES_R_1.pdf}
    \includegraphics[width=0.45\textwidth]{./Figures/allES_R_1.pdf}
    %\vspace{-1em}
    \caption{Employment Distributions}
    \floatfoot{Sources: CNEC and SUSB.}
    \label{fig:es}
    \end{center}
\end{figure}

%\begin{figure}[htbp]
%    \begin{center}
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/paperdist_R.pdf}%
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/agridist_R.pdf} \\
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/textiledist_R.pdf}
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/chemmatdist_R.pdf} \\
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/beverdist_R.pdf}
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/pooldist_R.pdf}
%    %\vspace{-1em}
%    \caption{Firm Size Distribution}
%    %\floatfoot{Source: China National Economic Census and the Statistics of U.S Businesses}
%    \label{fig:dist}
%    \end{center}
%\end{figure}

The left panel of Figure \ref{fig:es} depicts the firm size distribution for the top five polluting industries pooled together. %\footnote{Online Appendix E explains the calculation.}
We find that approximately 70\% of employment in the United States is accounted for by firms with 400+ employees, while in China the number is only 20\%.\footnote{In Online Appendix E, we compare the firm size distributions between China and the U.S. for each individual polluting industry. We find that while there is substantial heterogeneity across individual industries, large firms in China account for a much smaller employment shares than those in the U.S. for each industry.} %\footnote{Some readers may concern that the large difference of population size between the United States and China could drive the results. However, as shown in \citet{BentoRestuccia:2016}, there is no systematic relationship between average firm size and population size. In addition, because China has a much larger population size, the difference in population size between the two countries is more likely to make our case stronger.}
Compared with the U.S., a much larger share of production is carried out by small and medium-sized firms in China. %This implies that could be a candidate for explaining the high industrial pollution emissions in China. %\footnote{These results are consistent with those of \citet{WangWhalley:2014}, who compare the manufacturing concentration ratios (the share of market occupied by the largest firms) for China and the United States. According to Table 1 in their paper, the ratios of the concentration indicators of the United States over China for the top five polluting industries are higher than the {overall average}, which suggests that in polluting industries, large firms in the United States take a larger share of the markets.}
From the right panel of Figure \ref{fig:es}, we find a similar pattern for the whole manufacturing sector.

%\subsection{Size Distribution and Aggregate Pollution}

\textit{An Accounting Exercise.---}To see the potential importance of firm size distribution, we conduct the following accounting exercise. For the top five polluting industries in China, we fix the total output and  the estimated firm size--pollution intensity relationship, and replace the firm size distributions with those in the U.S. Then we compute the resulting average pollution intensity for these polluting industries pooled together.\footnote{Online Appendix F provides more details.} This simple exercise is complicated by the fact that the NGSPS reports firm-level output instead of number of employees. Therefore, we construct an employment--output relationship from a linear regression using the CNEC data.\footnote{There are different methods to construct the employment--output relationship using the CNEC data, and all of them provide similar results. See Online Appendix F for details.} %We use CNEC to overcome this problem. There are many ways to construct the employment-production relationship using CNEC, and each method has its own advantages and disadvantages. In this section, we report results where the employment-production relationship is constructed by linear regression. Calculation using alternative methods gives similar results. We leave the details of these alternative methods in Appendix A.

%\textit{An Accounting Exercise.---}To see how the firm size distribution affects aggregate pollution quantitatively, we conduct an accounting exercise in which, for the pooled polluting industries in China, we fix total output, but replace the employment distribution with that from the United States in order to calculate the implied level of average pollution intensity by using the firm size--intensity relationship estimated in Section \rom{1}.B.\footnote{We show the results for each of the five industries in Online Appendix F.} This simple exercise is complicated by the fact that the NGSPS only reports firm-level production values but not the number of employees. Therefore, we construct the employment--production relationship by using a linear regression with CNEC data.\footnote{There are many ways to construct the employment--production relationship by using CNEC data; however, these alternative methods all provide similar results. See Online Appendix F for details.} %We use CNEC to overcome this problem. There are many ways to construct the employment-production relationship using CNEC, and each method has its own advantages and disadvantages. In this section, we report results where the employment-production relationship is constructed by linear regression. Calculation using alternative methods gives similar results. We leave the details of these alternative methods in Appendix A.

We find that the average pollution intensity for the pooled top five polluting industries drops by 33\%, when the firm size distributions in these industries are changed to those in the U.S. This would cause a 25\% decrease in the average pollution intensity for the whole manufacturing sector if we leave other manufacturing industries untouched. Admittedly a crude approximation, this exercise nevertheless shows that distortions to firm size distribution could potentially have a large impact on aggregate pollution intensity.

%By doing so, we calculate that on average, for the top five polluting industries, the effect of changing the firm size distribution reduces emissions to 83\% of the original level. Changing the firm size distributions of these five industries while keeping those of all the other industries untouched leads to a reduction in total emissions of about 17\%. Although the exercise here is a crude approximation, it nevertheless shows that the firm size distribution can have a significant impact on aggregate industrial pollution.
%The numbers reported in Table \ref{tab:accounting} are the ratio of the aggregate pollution produced with the U.S. employment share distribution over that with the original Chinese distribution.
%\begin{table}[t]
%\footnotesize
%\centering
%\begin{threeparttable}
%\caption{Size Distribution on Pollution}
%\begin{tabular}{lccccccc}
%    \hline \hline
%                        & Paper  & Agricultural Food   & Textiles    & Chemical Materials    & Beverages        & Average  \\
%    \midrule
%    Average Intensity   & 43.5\% & 61.1\%              & 97.5\%      & 101.2\%               & 89.0\%           & 67.0\% \\
%    \bottomrule
%\end{tabular}
%\begin{tablenotes}
%     \footnotesize{
%     \item[$\dagger$]
%     Note: The numbers reported are aggregate pollution from the artificial U.S. production structure as a percentage of that of China. Column 6 (``Average'') calculates the weighted average of these ratios by using the percentage contributions in row one of Table \ref{tab:top5summary} as the weights.}
%\end{tablenotes}
%\label{tab:accounting}
%\end{threeparttable}
%\end{table}
%These results imply that by changing the employment share distribution to that of the United States, while keeping production at the same level, aggregate emissions in the Paper and Paper Products industry reduce to 43.5\% of the original level. On average, for the top five polluting industries, the effect of changing the firm size distribution reduces emissions to 67\% of the original level. The average is calculated by using the relative pollution levels of each industry as the weights, which is the first row of Table []. %Changing the firm size distributions of these five industries while keeping those of all the other industries untouched leads to a reduction in total emissions of about 25.5\%.

\section{The Model}

%The accounting exercise presented in the last section has several limitations. First, aggregate output is fixed. Indeed, when the size distribution changes, although pollution intensity decreases, aggregate pollution may increase as a result of the larger increase in aggregate output. In the language of scale, composition, and technique decomposition commonly used in the environmental literature [\citet{CopelandTaylor:2004} and \citet{Levinson:2009}], the accounting exercise evaluates only the technique effect, while instead we would like to consider the trade-off between scale and technique effects. Second, the firm size distribution is mechanically changed to that in the United States. From the accounting exercise alone, the factors that drive the difference between the firm size distribution of China and that of the United States remain unclear. We also do not know whether the implied employment distribution will in fact become that of the United States. Third, the relationship between firm size and emissions intensity is taken as exogenous and invariant. However, changes in the factors that affect the employment distribution may also influence the technology choices of firms, which makes the firm size--intensity relationship endogenous. Therefore, to better evaluate the environmental consequences of distortions to firm size, we need a model that (i) contains economic factors affecting both aggregate output and pollution, (ii) reveals which factors affect firm size and how, and (iii) explains the firm size--intensity relationship.

The above accounting exercise is illuminating but has several limitations. First, the firm size distribution is mechanically changed to that in the U.S. Although the U.S. economy is often taken as a benchmark in the development literature, it is unclear whether the differences in firm size distribution between China and the U.S. are mainly caused by distortions or by some other factors. Second, the accounting exercise captures the technique channel only, while distortions may operate through the scale and composition effects as well. Third, the relationship between firm size and emissions intensity is taken as exogenous in the accounting exercise. However, distortions to firm size may also influence firm's incentive to adopt different treatment technologies, which endogenizes the firm size-intensity relationship.

To address these issues, we build a two-sector (polluting and non-polluting) model of heterogeneous firms based on the classic \citet{Lucas:1978b} span-of-control model.\footnote{A widely used alternative  in the misallocation literature is the closed-economy version of \citet{Melitz:2003}. \citet{HsiehKlenow:2009} establish that these two frameworks are isomorphic for the effects of misallocation on aggregate output when the number of firms is fixed. In Appendix G.3, we show the qualitative equivalence between the two family of models.} The model features heterogeneity in firm productivities in both sectors, heterogeneity in pollution intensity and endogenous choice of pollution treatment technologies in the polluting sector. Motivated by the data, firms in both sectors are subject to correlated distortions [\citet{RestucciaRogerson:2008}] in the form of progressive output taxes. Throughout, we refer to the sectors as polluting and non-polluting, and technologies as dirty and clean. In the context of our model, the polluting sector is mapped to the top five polluting industries and the non-polluting sector to the rest of the manufacturing sector, while clean and dirty technologies are interpreted as biological and other (physical plus chemical) equipments as specified in Section \rom{1}.B.

%The model allows us to assess structurally the scale, composition and technique channels, through which counterfactual policies affect aggregate pollution. Moreover, the model allow us to further decompose the technique channel into: within sector reallocation across firms, and the change in clean technology adoption.
%For this purpose, we consider a two-sector neoclassical growth model with heterogeneous firms within each sector. We refer to the two sectors as polluting and non-polluting sector, which corresponds respectively to the five polluting industries and the rest of the manufacturing sector. Both sectors feature size-dependent distortions, while firms in the polluting sector face in addition imperfect environmental regulation and endogenous treatment technology choice. We assume two types of treatment technologies: dirty and clean. In the context of our model, these two technologies are interpreted as the {physical} and {biological} technology classifications discussed in Section \rom{2}.B.

\subsection{Setup}

We assume that the products of the two sectors in the model economy are perfect substitutes to each other.\footnote{In Online Appendix J, we consider the case in which the final good is a CES aggregation of the goods produced in two sectors, and conduct counterfactual experiments under different values of the elasticity of substitution in the CES function. For the counterfactual experiments we discuss in the main text, the quantitative results in the CES cases are similar to those in the perfect substitute case, largely because the experiments induce very limited reallocation between sectors.} The polluting sector generates emissions as a by-product of production as in \citet{CopelandTaylor:1994}, while the non-polluting sector does not. The top five polluting industries (hence the polluting sector) account for nearly 80\% of total industrial COD emissions and 20\% of total industrial output, with the rest accounted for by the non-polluting sector. Firms in both sectors behave competitively.

\textit{Household.}---%Assuming that there is perfect risk sharing within the household, the preference of the household could then be characterized by
%\begin{equation}
%    \sum_{t=0}^{\infty} \beta^t U(C_t)
%\end{equation}
%where $C_t$ is the consumption at time $t$ and $\beta$ is the discount factor.
There is a representative household with a continuum of members that maximizes the life-time utility $\sum_{t=0}^{\infty} \beta^t U(C_t)$,
%\begin{equation*}
%    \sum_{t=0}^{\infty} \beta^t U(C_t)
%\end{equation*}
where $C_t$ is the consumption at time $t$ and $\beta$ is the discount factor.

A fraction $\mu$ of the members are endowed with managerial talent that can be used in the polluting sector, drawn from a distribution with cumulative distribution function (CDF) $G^{d}(z)$. Meanwhile, a fraction $1-\mu$ of the members are endowed with managerial talent for the non-polluting sector with CDF $G^{c}(z)$. The support for both distributions are $Z \triangleq [0,\overline{z}]$. We assume that $z$ is fixed once drawn. Household members face an occupational choice between {worker} and {manager}. A worker supplies one unit of labor inelastically in exchange for wage income, and is perfectly mobile between sectors, while a manager in each sector runs a neoclassical firm and earns profits. As we see in Figure \ref{fig:es}, the firm size distributions in the polluting sector and the whole manufacturing sector are almost identical. Based on this observation, we assume that $G^{d}(z) = G^{c}(z) \triangleq G(z)$ from now on.

In addition, we assume that the household owns all the firms and capital in this economy. The goods produced in the non-polluting sector are taken as the numeraire. With perfect substitution between goods, the price of the goods produced in the polluting sector is also 1. $R$ and $W$ are the capital rental price and wage rate, respectively.

%There is a representative household with a continuum of members. Each household member is endowed with $z$ units of managerial talent, $z \sim G(z)$ with support $Z \triangleq [0,\overline{z}]$, where $G(z)$ is the cumulative distribution function and $g(z)$ is the probability density function. We assume that the support and distribution of $z$ are exogenous, and $z$ is fixed once drawn. Household members face an occupational choice between {worker} and {entrepreneur}. A worker supplies one unit of labor inelastically in exchange for wage income, whereas an entrepreneur rents capital and labor to run a neoclassical firm and earns profits. In the latter case, $z$ is also interpreted as firm-level TFP and with the understanding from the readers, we take the liberty to the term managerial talent and firm-level TFP interchangeably depending on whether we emphasize the household member or the firm.

%We assume that the managerial talent $z$ is sector specific in the sense that a share $\mu$ of the members can only choose to open firms in the polluting sector, while the rest $1-\mu$ have to enter the non-polluting sector. Workers are perfect mobile between sectors. Notice that by the law of large numbers, the ex ante distributions of $z$ in the two sectors are both $G(z)$. This is inspired by the fact that the firm size distributions in the polluting and non-polluting sectors are very similar (Figure \ref{fig:es}). We assume that the goods produced by the two sectors are perfect substitutes and are the numeraire.%\footnote{In Online Appendix J, we consider the case where there is a final good producer that aggregates the two intermediate goods following a CES production function. Because in the two quantitative experiments we study in the main text, there is very little cross-sector reallocation of production, the quantitative results from a CES setup are very similar. As a result, for simplicity, we choose the perfect substitute setup.} Let $R$ and $W$ be the capital and labor rental price respectively.

\textit{Firms.}---The production functions in the two sectors take the same form. Production requires managerial talent $z$, capital $k$, and labor $l$. For a firm run by a manager with managerial talent $z$, its output is
\begin{equation*}
    y = F(z,k,l) = z^{1-\gamma} (k^{\alpha}l^{1-\alpha})^{\gamma},
\end{equation*}
where $0<\gamma<1$ is the \emph{span-of-control} parameter, which governs the returns to scale at the firm level. Due to diminishing returns to scale at the firm level, a non-degenerate distribution of firm size can be supported, and firms earn positive profits as the payment to managerial talent $z$. %\footnote{We build our model based on \citet{Lucas:1978b}. However, all the qualitative properties of our model remain valid if instead we use a model with monopolistic competition [\citet{Melitz:2003}] since the two models are isomorphic. In the Melitz model, the decreasing returns to scale come from the concavity of the utility function. Similarly, firms in the Melitz model earn profits because of the monopolistic competition market structure, while the profits in our model are the factor payment to managerial talent $z$. We show the math details in Appendix G.3 [see also Appendix \rom{1} of \citet{HsiehKlenow:2009}].}
As is typically the case with studies that follow \citet{Lucas:1978b}, managerial talent $z$ is the only source of firm-level TFP in this model. In what follows, we also refer to $z$ as firm-level TFP.

In the non-polluting sector, firms do not discharge pollutants and are not subject to environmental regulations.  In the polluting sector, pollutants $e$ are generated as a by-product of production. Total emissions $e = \mathcal{E}(i,y)$ depend on the amount of output $y$, and the pollution treatment technology firms use, with $i = 1$ refers to the {clean} technology and $i = 0$ the dirty technology. Based on the empirical findings from Section \rom{1}, we assume that
\begin{equation}
\label{eq:intensity_size}
    \log\left(\frac{e}{y}\right) = \psi_0^i + \psi_1^i \log y, \quad i = 0,1,
\end{equation}
where $\psi_1^i<0$. This formulation indicates that larger firms emit fewer pollutants per unit of output, which is consistent with our empirical findings in Section \rom{1}.B.

Adopting the clean technology incurs fixed cost $R k_E$. %\footnote{Adding the variable costs of adopting the clean technology to our model will not change our quantitative results fundamentally. Specifically, the introduction of the variable costs would lead to an increase in calibrated value of $k_E$ but would not lead to materials changes in our quantitative results.See more related discussions in Online Appendix D.2.} %Furthermore, we assume that firms only invest in one treatment technology. We make this assumption because in the data only 7\% firms have more than one set of treatment equipment, and only 2\% have more than two sets.}
We assume that the environmental authority punishes firms for using the dirty technology. Similar in spirit to \citet{ShapiroWalker:2015}, we model the environmental regulations in a parsimonious way as a penalty which equals to a proportion $\xi$ of the firm's profits. The penalty is later redistributed to the household as lump-sum transfers.\footnote{In practice, China employs both economic incentives and command-and-control instruments in its current industrial pollution management and control system. A powerful tool used by the Chinese government is to randomly audit and then shut down firms that do not comply with environmental regulations for a certain period of time, which would cost these firms a fraction of their profits. See Trevor Nace, ``China shuts down tens of thousands of factories in widespread pollution crackdown,'' Forbes, October 24th 2017, for an anecdotal example.  We present a short summary of China's current environmental regulations based on \citet{Lin:2013} in Online Appendix D.3.} Therefore, the central trade-off in a firm's technology choice is weighing the adoption cost $R k_E$ against the profit losses from the environmental penalties. To the extent that firms' profits increase with firm size, we implicitly capture the size-dependency of the environmental regulations, in the sense that larger firms are usually monitored and inspected more closely. A key prediction from these assumptions is that, only firms above a certain size threshold would choose the clean technology, which is qualitatively consistent with our empirical findings on technology adoption in Section \rom{1}.B.\footnote{In Online Appendix D.2, we provide further evidence that the ratio of a firm's expenditures on treatment technology to firm output decreases with firm size, which supports the fixed cost assumption.}

%Installing the clean equipment incurs fixed cost $R k_E$, where we assume that the equipment is also rented from the market, just as the production capital $k$.\footnote{We choose to model the installation costs as one-time fixed costs as opposed to fixed costs plus operating costs or size-dependent fixed costs because the latter two are not supported by the empirical evidence. We also assume that the fixed cost is only associated with the clean technology. Qualitatively, assuming that the dirty technology also requires a fixed cost does not affect the property of the model. This is equivalent to a decrease in the cost of the clean technology, since what matters for the firm's decision is the difference between the two costs. See Online Appendix D.2 for further details.} %Furthermore, we assume that firms only invest in one treatment technology. We make this assumption because in the data only 7\% firms have more than one set of treatment equipment, and only 2\% have more than two sets.}
%We assume that the environmental authority penalizes firms for using dirty technology. We take a similar approach as \citet{ShapiroWalker:2015} to capture the regulations in a parsimonious way by modeling the penalty as a proportion of $\xi$ of the total profits of these firms.\footnote{In practice, China employs both economic incentives and command-and-control instruments in its current industrial pollution management and control system. A pollution levy system requires that firms pay levies according to how much pollutants they discharge. In addition, local authorities penalize firms for fraudulent report of pollutants emissions and temporarily suspend firms using obsolete treatment technology from production under some occasions. \citet{Lin:2013} provides a detailed description of China's current environmental regulations. We present a short summary in the Online Appendix D.3 based on his work. The $\xi$ is used to summarize all these regulations jointly.} The penalty is distributed back to the household as lump-sum transfers.

%As a result, the core trade-off of firm's technology adoption decision is between the installation cost $R k_E$ and the profit loss. Because firm's profits and size increase with its productivity $z$, by assuming a constant fraction $\xi$, we are implicitly capturing the fact that environmental regulation is also size-dependent, that is large firms are being watched more closely. Also since our model is deterministic and firms are risk neutral, $\xi$ captures the deterrence effects of the regulations as well.

%\textit{Regulators.}---We assume that the environmental authority monitors the adoption of the clean technology by firms with probability $p$. When a firm using the dirty technology is inspected, we assume that a proportion $\xi$ of its total profits is confiscated by the regulating agency. As a result, firms using the dirty technology lose a proportion of $p\xi$ of their total profits. The confiscated profits are distributed to the household as lump-sum transfers, meaning that they do not affect the decision problem of household members. This reduced-form way of modeling monitoring policy could, for instance, be rationalized by a mixed strategy Nash equilibrium of a behind-the-scenes ``monitoring game.''
%
%The current industrial pollution management and control system in China consists of {economic incentives} and {command-and-control instruments}.\footnote{See Chapter 5 of \citet{WorldBankChinaPollution:2001} for a detailed description.} The pollution levy system is the most widely used economic instrument in China. However, it places limited constraints on the emissions of firms because the penalty imposed is very low. Firms only have to pay for the emissions that go {beyond} the national standard. Moreover, these emissions are {self-reported} and the truthfulness of the reported emissions is imperfectly examined by the regulators.\footnote{This may be a common problem for developing countries. For example, \citet{Dufloetal:2013} conduct a field experiment in India and find that the emissions reporting system is largely corrupted, with auditors systematically reporting plant emissions just below the standard.} Further, for firms that discharge multiple pollutants and when the levels of more than one of the pollutants are above the national standards, firms only have to pay for the one that leads to the highest penalty. We calculate from the CNEC the pollution fees levied on firms as a proportion of total labor compensation. We find that for firms with strictly positive emissions fees, these fees only account for 0.06\% (median) and 0.3\% (mean) of the labor compensation.
%
%Therefore, in practice, environmental agencies rely mostly on command-and-control instruments. To implement the regulation, field inspections are conducted by the staff of local environmental agencies. %Using data from the Environmental Protection Bureau in Zhenjiang, \citet{Dasguptaetal:2001} showed that at the firm level, compared to variations in pollution levies, variations in the number of field inspections better explains the variations in pollution levels.
%Field staff typically check the type of treatment equipment that firms have installed and test the emissions intensity of major pollutants. Firms found to be at fault are usually suspended from production until the issues are resolved.\footnote{See \citet{Dasguptaetal:2001} for a case study of Zhenjiang.} In our model, the proportion $\xi$ of the profits confiscated is used to approximate these costs. Since, according to Table \ref{tab:techeffi}, the treatment technology used by firms is highly correlated with pollution intensity, we assume that the regulator in our model checks only treatment technologies. Although the local environmental agencies also monitor total emissions, these regulations are usually based on a {provincial-level} aggregation. They are thus less relevant to the firm-level decision that we study here.\footnote{Firm inspections in the United States also target the adopted treatment technologies. See \citet{BeckerHenderson:2000} and the references therein for more details.}
%%\footnote{Firm level inspections in the U.S are also targeted mainly on the adopted treatment technologies. For example, as is stated in the 1977 Amendment of the Clean Air Act, each July every county in the U.S will be classified as either an {attainment} or a {non-attainment} county according to their overall pollutants emissions level. Firms in {non-attainment} counties are subject to substantially stronger environmental regulations. For instance, newly established firms in these counties are required to meet the standard of Lowest Achievable Emission Rate (LAER) which demands the installation of the cleanest possible technology supposedly regardless of costs. While on the other hand, in {attainment} counties only Best Available Control Technology (BACT) which incorporates cost considerations is required. Similarly, existing firms are also subject to stricter regulations on production and end-of-pipe treatment technologies in non-attainment counties than in the attainment counties. See \citet{BeckerHenderson:2000} and the references therein for more details.}

%\subsection{Firm-level Distortions}
%
%Recent studies of the Chinese economy have described large distortions at both the sector and the firm levels, which negatively affect aggregate productivity and output considerably.\footnote{See \citet{HsiehKlenow:2009}, \citet{Songetal:2011}, \citet{Brandtetal:2013}, and \citet{TombeZhu:2015}, among others.} Following the seminal approach developed by \citet{HsiehKlenow:2009,HsiehKlenow:2014}, we model and estimate firm-level distortions by using variations in the average products of capital and labor across firms.\footnote{Though average factor product dispersion could arise due to dynamic capital adjustment [\citet{Askeretal:2014}], for the case of China, \citet{DavidVenkateswaran:2017} show that dynamic adjustment costs only accounts for about 10\% of the variation in average factor product in the data. Naturally, we also cannot separate out variations in average factor product caused by firm-specific characteristics, for instance firm-specific markup or capital share. However, we argue that it would require empirical implausible variations of such parameters to generate the patterns of the data we observe.} %An important feature of Cobb-Douglas production function is that the ratio of average factor products to marginal factor products is constant. Therefore, the average factor products also provide information on marginal factor products.\footnote{In particular, using the notations later in this section, we can verify that $\partial y/\partial k = \alpha \gamma z^{1-\gamma} k^{\alpha \gamma - 1} l^{(1-\alpha)\gamma} = \alpha \gamma (y/k)$. Similarly, we can show that $\partial y/\partial k = (1-\alpha) \gamma (y/l)$.}
%More specifically, if we let $\tau_{z_i}, \tau_{k_i}$, and $\tau_{l_i}$ be the wedges that firm $i$ faces in the product, capital, and labor markets, respectively, the profit maximization problem of firm $i$ is
%\begin{equation*}
%    \pi_i = \max_{k_i,l_i} \left\{ (1-\tau_{z_i}) z_i^{1-\gamma} (k_i^{\alpha} l_i^{1-\alpha})^{\gamma} - (1+\tau_{k_i})Rk_i - (1+\tau_{l_i})W l_i \right\}.
%\end{equation*}
%By using the first-order conditions, the firm level average product of capital $\phi_{k_i}$, labor $\phi_{l_i}$, and the capital/labor ratio $\kappa_i$ can be expressed as
%\begin{align}
%    \label{eq:phik}
%    \phi_{k_i} &= \frac{y_i}{k_i} = \frac{(1+\tau_{k_i})R}{\alpha \gamma (1-\tau_{z_i})}, \\
%    \label{eq:phil}
%    \phi_{l_i} &= \frac{y_i}{l_i} = \frac{(1+\tau_{l_i})W}{(1-\alpha) \gamma (1-\tau_{z_i})}, \\
%    \label{eq:kappa}
%    \kappa_i &= \frac{k_i}{l_i} = \frac{\alpha}{1-\alpha} \cdot \frac{(1+\tau_{l_i})W}{(1+\tau_{k_i})R}.
%\end{align}
%The above equations show that in the absence of any distortions ($\tau_{z_i} = \tau_{k_i} = \tau_{l_i} = 0$), $\phi_{k_i}, \phi_{l_i}$, and $\kappa_i$ should be equalized across firms. Equations (\ref{eq:phik}) and (\ref{eq:phil}) state that firms that face higher distortions in the capital (labor) and/or product markets demonstrate a higher average product of capital (labor). In addition, according to Equation (\ref{eq:kappa}), the capital/labor ratio increases with the {relative} size of labor to the capital market wedge. By using firm-level data on total production value, the book value of capital stock, and labor compensation from the CNEC, we calculate $z_i, \phi_{k_i}, \phi_{l_i}$, and $\kappa_i$ for each firm in our sample. Anticipating the calibration later, here we set $\gamma = 0.93$. Figure \ref{fig:phiklkappa} shows on log scales the scatterplots of $\phi_{k_i}, \phi_{l_i}$, and $\kappa_i$ against firm-level productivity $z_i$ for the Paper and Paper Products industry. We plot the Paper industry for demonstration purpose, qualitatively the results for all five polluting industries combined are the same.
%\begin{figure}[t]
%    \begin{center}
%    \includegraphics[width=0.4\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/arpk.pdf}%
%    \includegraphics[width=0.4\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/arpl.pdf} \\
%    \includegraphics[width=0.4\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/klratio.pdf}
%    \caption{Factor and Product Market Distortions}
%    \floatfoot{Source: CNEC. All panels are plot in log scale. Lines are least square fit.}
%    \label{fig:phiklkappa}
%    \end{center}
%\end{figure}
%
%Two patterns emerge from Figure \ref{fig:phiklkappa}. First, from the two upper panels, we see that both $\phi_{k_i}$ and $\phi_{l_i}$ are positively correlated with $z_i$, which suggests that more productive firms have higher average products of both capital and labor. Expressed in wedges, this means that both $(1+\tau_{k_i})/(1-\tau_{z_i})$ and $(1+\tau_{l_i})/(1-\tau_{z_i})$ are higher for more productive firms.\footnote{Generally speaking, the estimated elasticities of $\phi_k$ and $\phi_l$ to $z$ depend on the value of $\gamma$, since $z$ is calculated for a given $\gamma$. For $\gamma = 0.93$ used here, the elasticities of $\phi_k$ and $\phi_l$ to $z$ equal 0.03, which taken at its face value seem well below the estimates in the literature. For instance, by using comprehensive micro data, \citet{HsiehKlenow:2014} find that the elasticity is 0.1 for the United States and between 0.5 and 0.6 for India and Mexico. By using data from the World Bank's Enterprise Surveys, \citet{BentoRestuccia:2016} find that the elasticities range from 0.22 to 0.74, averaging 0.52. However, we argue that this is because in the calibration of our general equilibrium model, as is explained in details in Section \rom{3}, in order that our model generates a wide range of firm size, a value of $\gamma$ that is higher than commonly used in the literature is needed. If we assume $\gamma = 0.85$ following \citet{AtkesonKehoe:2005}, then the elasticities of $\phi_k$ and $\phi_l$ to $z$ are both about 0.25, which are quantitatively consistent with those in the literature. The value of $\gamma$ used, however, will not affect the qualitative conclusion that large firms are subject to more distortions.}
%%If we assume $\gamma = 0.85$ following \citet{AtkesonKehoe:2005}, then the elasticities of $\phi_k$ and $\phi_l$ to $z$ are both about 0.25, meaning that doubling firm productivity is associated with a 25\% increase in the average revenue product of factor inputs.\footnote{Generally speaking, the estimated elasticity depends on the value of $\gamma$, since $z$ is calculated for a given $\gamma$. However, because we are using a \citet{Lucas:1978b} model, while \citet{HsiehKlenow:2014} and \citet{BentoRestuccia:2016} both use the \citet{Melitz:2003} model, our results are not directly comparable with theirs. Yet broadly speaking, our estimates are quantitatively consistent with theirs. By using comprehensive micro data, \citet{HsiehKlenow:2014} find that the elasticity is 0.1 for the United States and between 0.5 and 0.6 for India and Mexico. By using data from the World Bank's Enterprise Surveys, \citet{BentoRestuccia:2016} find that the elasticities range from 0.22 to 0.74, averaging 0.52.}
%This could be because more productive firms are subject to higher factor or product market distortions or both. Second, from the lower panel, we see that the capital/labor ratio is at best weakly negatively correlated with $z$. The least squares estimate of the elasticity is $-$0.0057 and the $R^2$ is only 0.053. This finding indicates that the {relative} wedge firms face in the capital and labor markets does not depend strongly on the idiosyncratic productivity of firms, which in the context of our model implies $1+\tau_{k_i} \approx 1+\tau_{l_i}$.\footnote{This also explains the reason why $\phi_{k_i} \approx \phi_{l_i}, \forall i$.} Since we cannot separately identify the three wedges, for simplicity, we assume $\tau_{k_i} = \tau_{l_i} = 0$ and attribute all the variations in the average product of factors to the wedges in the product market $\tau_{z_i}$. Whether we assume $\tau_{k_i} = \tau_{l_i} = 0$ or alternatively $\tau_{z_i} = 0$ does not affect our results, but the interpretations need to be changed accordingly.\footnote{For example, we cannot distinguish between the data-generating process we use here and another process where $\tau_{k_i}$ and $\tau_{l_i}$ increase simultaneously, while $\tau_{z_i}$ is equal to zero. Moreover, when estimating the elasticity, since $y,k$ and $l$ enter the regression on both sides, when any of the variables are measured with measurement error, such error may drive spurious correlations. While we cannot rule out this possibility completely, we argue that this does not seem to be the case here. In particular, if $y$ is measured with extreme measurement error, the regression coefficient of $\phi_k$ over $z$ is $1-\gamma$. Similarly, if instead $k$ is measured with extreme measurement error, the regression coefficient is $(1-\gamma)/\gamma$. We calculate $\phi_{k_i}$ and $z_i$ by using different values of $\gamma$ and the regression coefficients do not vary as predicted by either case.}
%
%\begin{figure}[t]
%    \begin{center}
%    \includegraphics[width=0.50\textwidth]{D:/PaperDrafts/Pollution/PrelimResults/tauz.eps}
%    \caption{The Generic Tax Function of $\tau_z$}
%    \label{fig:tauz}
%    \end{center}
%\end{figure}
%In the spirit of \citet{BentoRestuccia:2016}, we implement these idiosyncratic wedges in the model by positing a generic ``tax'' function that specifies the wedges as a function of a firm's productivity $z$:\footnote{This tax function has a long history in macro-public finance literature, for instance \citet{Benabou:2002}. We thank a referee for pointing this out. Recent studies that adopt the same tax functions include \citet{AdamopoulosRestuccia:2014}, \citet{HsiehKlenow:2014}, and \citet{Guneretal:2017}.}
%\begin{equation}
%    \tau_z = \max \left\{ 0, 1 - \phi_0 z^{\phi_1} \right\}.
%\label{eq:tax}
%\end{equation}
%%From now on, we use $\tau_z$ instead of $\tau_y$ to denote the product market frictions.We assume that the taxes collected are returned to the household as lump-sum transfers.
%Anticipating the benchmark calibration in the next section, the wedge function specified in Equation (\ref{eq:tax}) is increasing and concave in $z$, with the lower and upper bounds being 0 and 1, respectively. The shape of the function captures the size dependency of the product market distortions where the wedges are higher for larger firms. The actual tax function used in the quantitative exercises is plotted against log productivity in Figure \ref{fig:tauz}. %\footnote{There is one difference between Equation (\ref{eq:tax}) and the tax function used by \citet{AdamopoulosRestuccia:2014}. To model the size dependency, in their specification, the authors use an exponential function as opposed to the power function here. We choose the power function because it is consistent with the log-linearity of $\phi_k$ ($\phi_l$) and $z$, while the exponential function implies a tax scheme that increases much sharper with respect to productivity than the empirical counterpart for our case.} %We rule out subsidies ($\tau_z < 0$) because $\lim_{z \downarrow 0} \tau_z = -\infty$ which leads to unrealistic solutions.}
%
%The idiosyncratic $\tau_z$ is meant to capture implicitly a variety of policies and institutions that allocate factors away from large productive firms. For example, this is consistent with the fact that large productive firms face transportation costs, additional management costs, and/or local protectionism and trade barriers that impede the inter-regional flow of goods [\citet{Young:2000}, \citet{HsiehKlenow:2014}, \citet{TombeZhu:2015}, and \citet{Eberhardtetal:2016}]. It could also be that smaller firms are subject to preferential tax treatment.\footnote{For instance, value added taxes for firms with an annual industrial output value of less than CNY 1 million is 3\%, while firms with a production scale larger than CNY 1 million are subject to a 13\% tax rate.} Some of these examples, for instance transportation costs and management costs, are not exactly the ``distortions'' narrowly defined as inefficiencies from policy interventions. Hence a more accurate interpretation of the distortions captured by our model is any factors that prevent firms from reaching their full potentials in a frictionless world. We stick to the word \textit{distortion} mainly for convenience. From a policy perspective, it would be of great interest to follow the so-called direct approach in \citet{RestucciaRogerson:2013} and study the specific policies and regulations that we observe in the economy. We leave this important task to future work.\footnote{\citet{AdamopoulosRestuccia:2014} and \citet{AGpaper} are two recent examples where observable policy distortions are investigated. \citet{AdamopoulosRestuccia:2014} study the observable farm-level price distortions while our own work \citet{AGpaper} focus on distortions created by migration and land distribution policies in China.}

\textit{Correlated Distortions.---}Finally, we introduce the correlated distortions at the firm level to both sectors as implicit progressive output taxes which increase with a firm's productivity levels. The functional form of the tax is:
\begin{equation}
    \tau^i_z = \max \left\{ 0, 1 - \phi_0^i z^{\phi_1^i} \right\}, i=c,d,
\label{eq:tax}
\end{equation}
where the parameter $\phi_0^i$ determines the mean level of the taxes, and the parameter $\phi_1^i$ determines the progressiveness of the taxes. When $\phi_1^i<0$, the tax rates are increasing in firm TFP $z$, and the progressiveness of the taxes is decreasing in the value of $\phi_1^i$.\footnote{This tax function has a long history in macro-public finance literature, which dates back to at least \citet{Benabou:2002}. We thank a referee for pointing this out. Recent studies that use this function include \citet{HsiehKlenow:2014}, \citet{BentoRestuccia:2016} and \citet{Guneretal:2017}. In Online Appendix D.1, we discuss distortions that take more general forms, and present empirical evidence in support of this specific form. See Figure D.2 in Online Appendix D for an illustration of this tax function.}

The tax function (\ref{eq:tax}) is consistent with the positive relationship between firm TFP and average product of factors we find in the CNEC data. This can be seen from Figure \ref{fig:apf}, which depicts the geometric average of the average product of labor and capital $(y/k)^\alpha(y/l)^{1-\alpha}$ against firm TFP $z$ for both the polluting and non-polluting sectors.\footnote{To estimate firm TFP and average product of factors, we need to assign the values to the parameter that determines the income share of capital, $\alpha$, and the span-of-control parameter, $\gamma$. We take these values from our calibration in Section \rom{3}, which implies that $\alpha = \text{0.54}$ and $\gamma = \text{0.93}$. The positive relationship between firm TFP and average product of factors, however, hold for a wide range of values for $\alpha$ and $\gamma$, in addition to the values used here. Moreover, similar patterns are also observed in other countries. For instance, \citet{HsiehKlenow:2014} find a strong and positive relationship between firm TFP and average product of factors in India and Mexico, while \citet{BentoRestuccia:2016} document similar patterns for dozens of developing countries.} It can also be seen that the relationships between the average product of factors and firm TFP in these two sectors share very similar slopes. Based on this observation, we assume $\tau^c_z = \tau^d_z \triangleq \tau_z$ from now on.

The firm-level distortions $\tau_z$ are meant to capture a variety of policies and institutions that reallocate production factors from large productive to small unproductive firms, following the \emph{indirect} approach in the misallocation literature [\citet{RestucciaRogerson:2013}]. These policies and institutions include size-dependent taxes and subsidies that favor small firms, imperfect tax enforcement that induces tax evasion among small firms, financial constraints and contractual frictions that impede the growth of productive firms [\citet{CooleyQuadrini:2001}, \citet{ClementiHopenhayn:2006} and \citet{Akcigitetal:2018}], and local protectionism and internal trade barriers that disproportionately affect large productive firms [\citet{Eberhardtetal:2016} and \citet{TombeZhu:2015}]. While it is of great policy interest to study specific policies and institutions underlying $\tau_z$, we leave this important task to future work.

\begin{figure}[t]
    \begin{center}
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/paperES_R.pdf}%
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/agriES_R.pdf} \\
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/textileES_R.pdf}
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/R_Implementation/chemmatES_R.pdf} \\
    \includegraphics[width=0.45\textwidth]{./Figures/apf_pol_weight.pdf}
    \includegraphics[width=0.45\textwidth]{./Figures/apf_npol_weight.pdf}
    %\vspace{-1em}
    \caption{Average Product of Factors versus Firm Productivity in the Cross-Section}
    \floatfoot{Source: CNEC. Both variables have been normalized to reflect the relative values to the sectoral averages, which are similar across sectors.}
    \label{fig:apf}
    \end{center}
\end{figure}

\subsection{Firm's Profit Maximization Problem}

%Recent studies of the Chinese economy have described large distortions at both the sector and the firm levels, which negatively affect aggregate productivity and output considerably.\footnote{See \citet{HsiehKlenow:2009}, \citet{Songetal:2011}, \citet{Brandtetal:2013}, and \citet{TombeZhu:2015}, among others.}
%We assume that all firms face an idiosyncratic distortions $\tau_z$ depending on their productivity level. In the spirit of \citet{BentoRestuccia:2016}, we implement these idiosyncratic wedges in the model by positing a generic ``tax'' function:\footnote{This tax function has a long history in macro-public finance literature, for instance \citet{Benabou:2002}. We thank a referee for pointing this out. Recent studies that adopt the same tax functions include \citet{AdamopoulosRestuccia:2014}, \citet{HsiehKlenow:2014}, and \citet{Guneretal:2017}. Online Appendix D.1 presents detailed empirical evidence that motivates us using this specification.}\textsuperscript{, }\footnote{The idiosyncratic $\tau_z$ is meant to capture implicitly a variety of policies and institutions that allocate factors away from large productive firms. For example, this is consistent with the fact that large productive firms face transportation costs, additional management costs, and/or local protectionism and trade barriers that impede the inter-regional flow of goods [\citet{Young:2000}, \citet{HsiehKlenow:2014}, \citet{TombeZhu:2015}, and \citet{Eberhardtetal:2016}]. It could also be that smaller firms are subject to preferential tax treatment. Some of these examples, for instance transportation costs and management costs, are not exactly the ``distortions'' narrowly defined as inefficiencies from policy interventions. Hence a more accurate interpretation of the distortions captured by our model is any factors that prevent firms from reaching their full potentials in a frictionless world. We stick to the word \textit{distortion} mainly for convenience. From a policy perspective, it would be of great interest to follow the so-called direct approach in \citet{RestucciaRogerson:2013} and study the specific policies and regulations that we observe in the economy. We leave this important task to future work.}
%\begin{equation}
%    \tau_z = \max \left\{ 0, 1 - \phi_0 z^{\phi_1} \right\}.
%\label{eq:tax}
%\end{equation}
%Following the evidence in the Online Appendix D.1, we assume that the the tax function (\ref{eq:tax}) is the same for both sectors.

Firms in the non-polluting sector solve the standard profit maximization problem:
\begin{equation*}
    \pi^c(z) = \max_{k,l} \left\{ (1-\tau_z) z^{1-\gamma}(k^{\alpha}l^{1-\alpha})^{\gamma} - Wl - Rk \right\},
\end{equation*}
where the superscript $c$ stands for ``clean.''

Firms in the polluting sector choose first the optimal treatment technology and then the optimal production plan. The profit function $\pi^d(z)$ is the envelope of the profits when using the dirty technology $\pi_0^d(z)$ and clean technology, $\pi_1^d(z)$:
\begin{equation*}
    \pi^d(z) = \max \left\{ \pi_0^d(z), \pi_1^d(z) \right\},
%\label{eq:piz}
\end{equation*}
where the superscript $d$ stands for ``dirty,'' and the subscript denotes the treatment technologies. Firms using the clean technology are not subject to environmental penalties, hence the profit $\pi_1^d(z)$ is:
\begin{equation*}
    \pi_1^d(z) = \max_{k,l} \left\{ (1-\tau_z) z^{1-\gamma} (k^{\alpha} l^{1-\alpha})^{\gamma} - Wl - R (k+k_E) \right\}.
%\label{eq:pi1z}
\end{equation*}
However, when a firm chooses the dirty technology, it will be punished by the environmental authority, and lose a fraction $\xi$ of its gross profit. Its net profit $\pi_0^d(z)$ is thus:
\begin{equation*}
    \pi_0^d(z) = \max_{k,l} \left\{ \left( 1 - \xi \right)
        \left[ (1-\tau_z) z^{1-\gamma} (k^{\alpha} l^{1-\alpha})^{\gamma} - Wl - Rk \right] \right\}.
%\label{eq:pi0z}
\end{equation*}

The firm's problem implies that there exists a firm size threshold $\tilde{l}^d$ above which firm will choose the clean technology, which is consistent with our empirical finding that larger firms are more likely to use the clean technology. This can be seen by comparing the potential profits from using clean versus dirty technology:
\begin{equation*}
    \pi_1^d(z) - \pi_0^d(z)  = \frac{\xi(1-\gamma)}{(1-\alpha)\gamma}Wl^d(z) - Rk_E.
\end{equation*}
In addition, when $1-\gamma+\phi_1 > 0$, which is satisfied given our calibrated values for $\gamma$ and $\phi_1$, $\pi_i^d(z)$ and $l^d(z)$ are both increasing $z$. Then there exists a threshold $\tilde{z}_d$ such that only firms with TFP higher than $\tilde{z}_d$ would choose clean technology.\footnote{Please see Lemma 2 in Online Appendix G.1 for details.} There also exist two thresholds $\hat{z}_i, i=c,d,$ such that in sector $i$, only members with $z > \hat{z}_i$ choose to be managers.

%The first-order conditions from this problem imply that
%\begin{equation*}
%    \pi_1^d(z) = \frac{1-\gamma}{(1-\alpha)\gamma}Wl^d(z)-R k_E.
%%\label{eq:pi1z}
%\end{equation*}
%Note that here treatment equipment $k_E$ cannot be used to produce the final product. This assumption is based on the empirical finding of \citet{ShadbegianGray:2005}.
%\begin{equation*}
%    \pi_0^C(z) = (1-\xi) \left[ (1-\tau_z) z^{1-\gamma} (k^{\alpha} n^{1-\alpha})^{\gamma} - Wn - Rk \right],
%\end{equation*}
%where the superscript $C$ indicates ``caught.'' If the firm succeeds in evading inspection, the profit function is
%\begin{equation*}
%    \pi_0^{E}(z) = (1-\tau_z) z^{1-\gamma} (k^{\alpha} n^{1-\alpha})^{\gamma} - Wn - Rk,
%\end{equation*}
%where the superscript $E$ indicates ``evaded.'' Because we assume perfect risk sharing within the household, these entrepreneurs will {not} have precautionary motives and will simply maximize the expected profits over $\pi_0^C$ and $\pi_0^{E}$:
%\begin{equation*}
%    \pi_0(z) = \max_{k,n} \left\{ (1-p)\pi_0^{E}(z) + p\pi_0^{C}(z) \right\}.
%\end{equation*}
%Some algebra yields

%and the first-order conditions from this problem imply that
%\begin{equation*}
%    \pi_1^d(z) = \frac{(1-\gamma)(1 - \xi)}{(1-\alpha)\gamma}Wl^d(z).
%%\label{eq:pi1z}
%\end{equation*}

%The standard result from the Lucas model implies that there exist two thresholds $\hat{z}_d$ and $\hat{z}_c$ such that in each sector, all members with $z$ higher than the corresponding thresholds choose to become entrepreneurs. In addition, when $\pi_d^0(z)$ and $\pi_d^1(z)$ satisfy single-crossing condition, there also exists a third threshold $\tilde{z}_d$ such that all firms in the polluting sector with TFP higher than $\tilde{z}_d$ use clean technology. See Result \ref{prop:friction} in Section \rom{2}.D and Lemma 2 in Online Appendix G.1 for details.

\subsection{Stationary Equilibrium}

In this section, we close the model by specifying the household's optimization problem, and then define the stationary equilibrium.

The household solves the standard consumption-saving problem:
\begin{align}
\label{eq:household}
    & \max_{C_t, K_{t+1}} \sum_{t=0}^{\infty} \beta^t U(C_t) \\
    & s.t.  \nonumber \\
    & \qquad C_t + K_{t+1} - (1-\delta)K_t = I_t, \nonumber
\end{align}
where $K_t$ is aggregate capital, $\delta$ is the depreciation rate, and $I_t$ is household income, which we specify in detail shortly.%\footnote{We assume here that the household values only consumption and not environmental quality. This assumption is innocuous in the competitive equilibrium, since no individual household member has control over aggregate environmental quality. However, the assumption does affect the results if a planner's problem is studied or if we want to evaluate the welfare effect of different policies.}
\footnote{Here we do not explicitly model the household's preference over water quality. Adding water quality to the household's utility function will not change the firm's choices in a competitive equilibrium, as no individual household member has any control over the aggregate environmental quality. However, if we study the social planner's problem instead, the household's preference over water quality will become a major modeling choice.} The solution to (\ref{eq:household}) is the standard intertemporal Euler equation:
\begin{equation*}
    U'(C_t) = \beta U'(C_{t+1})(1+R_{t+1}-\delta),
%\label{eq:euler}
\end{equation*}
which pins down the equilibrium interest rate.%\footnote{There is no uncertainty at the household level since we assume perfect risk sharing and a continuum of household members which wash out all idiosyncratic uncertainty. As a result, on the right hand side of (\ref{eq:euler}) there is no expectation operator.}

Household income $I_t$ comes from three sources: wage and interest income, firms' profits, and lump-sum transfers from taxes $\tau_z$ and environmental penalties $\xi$. %To characterize $I_t$, we need some additional notation. We denote $Z_0 = \{z \in [\hat{z}_t, \overline{z}] | \pi_0(z) \geq \pi_1(z) \}$ as the set of firms operating with the dirty technology and $Z_1 = \{z \in [\hat{z}_t, \overline{z}] | \pi_0(z) < \pi_1(z) \}$ as the set of firms using the clean technology. Note that for the intermediate case, where $0 < \hat{z} < \tilde{z} < \overline{z}$, Result \ref{prop:friction} implies $Z_0 = [\hat{z}, \tilde{z})$, and $Z_1 = [\tilde{z}, \overline{z}]$. %Since the ``taxes'' are distributed to household as lump-sum transfers, in the budget constraint taxes and transfers cancel each other out.
If we let $T$ denote the transfers, then household income is\footnote{We have implicitly assumed that firms using clean and dirty technologies coexist in the equilibrium, meaning that $\hat{z}_d<\tilde{z}_d$. This is also the case in our quantitative exercises later.}
\begin{align*}
    I &= R K + W [\mu G(\hat{z}_d)+(1-\mu)G(\hat{z}_c)] + T \\
      & \qquad + \mu\left[\int_{\hat{z}_d}^{\tilde{z}_d} \pi^d_0(z) dG(z) + \int_{\tilde{z}_d}^{\overline{z}} \pi^d_1(z)dG(z)\right] + (1-\mu)\int_{\hat{z}_c}^{\overline{z}}\pi^c(z)dG(z),
\end{align*}
where the transfer $T$ is given by
\begin{equation*}
    T = \mu\xi\int_{\hat{z}_d}^{\tilde{z}_d}\pi^d(z)dG(z) + \tau_z\left[\mu \int_{\hat{z}_d}^{\overline{z}} \pi^d(z) dG(z) + (1-\mu)\int_{\hat{z}_c}^{\overline{z}} \pi^c(z) dG(z) \right].
\end{equation*}%the five terms are capital rental income, wage income, profits from dirty and clean firms, and government transfers, respectively. %The law of large numbers guarantees that the {ex ante} probability of being inspected equals the {ex post} proportion of firms that actually are inspected.

Let $Y$ be aggregate output and $E$ be aggregate pollution, then the stationary equilibrium of the model is defined as follows.
\begin{definition}
\label{def:ge}
A stationary equilibrium of the model consists of prices $\{W, R \}$, allocations $\{C, K, Y \}$, firms' policy functions $\{k^j(z),l^j(z),y^j(z),\pi^j(z)\}, j = c,d$, thresholds for household's occupational choices $\{\hat{z}_c, \hat{z}_d\}$ and polluting firms' technology adoption $\tilde{z}_d$, as well as aggregate pollution $E$, such that:
\begin{enumerate}[(i)]
    \item
    Given factor prices $\{W, R\}$, $\{C, K, \hat{z}_c, \hat{z}_d\}$ solve the household's optimization problem;
    \item
    Given factor prices $\{W, R\}$, $\{k^j(z),l^j(z),y^j(z),\pi^j(z)\}$, $j = c,d$, and $\tilde{z}_d$ solve the firms' optimization problems;
    \item
    Factor prices $\{W, R\}$ clear all markets:
    \begin{itemize}
        \item
        Labor Market:
        \begin{equation*}
            \mu G(\hat{z}_d) + (1-\mu) G(\hat{z}_c) = \mu \int_{\hat{z}_d}^{\overline{z}} l^d(z)dG(z) + (1-\mu)\int_{\hat{z}_c}^{\overline{z}} l^c(z)dG(z),
        \end{equation*}
        \item
        Capital Market:
        \begin{equation*}
            K = \mu \left[\int_{\hat{z}_d}^{\overline{z}} k^d(z)dG(z) + k_E [G(\overline{z})-G(\tilde{z}_d)]\right] + (1-\mu)\int_{\hat{z}_c}^{\overline{z}} k^c(z)dG(z),
        \end{equation*}
        \item
        Product Market (Walras' Law):
        \begin{equation*}
            C + K - (1-\delta)K = \mu \int_{\hat{z}_d}^{\overline{z}} y^d(z)dG(z) + (1-\mu)\int_{\hat{z}_c}^{\overline{z}} y^c(z)dG(z);
        \end{equation*}
    \end{itemize}
    \item
    Aggregate Pollution:
    \begin{equation*}
        E = \mu\left[\int_{\hat{z}_d}^{\tilde{z}_d} e\left(0,y^d(z)\right)dG(z) + \int_{\tilde{z}_d}^{\overline{z}} e\left(1,y^d(z)\right) dG(z)\right].
    \end{equation*}
\end{enumerate}
\end{definition}
\noindent The algorithm to compute the stationary equilibrium is sketched in Online Appendix I.

\subsection{Correlated Distortions, Technology Adoption, and Aggregate Pollution}
\label{sec:analytical}

In this section, we show qualitatively how correlated distortions modeled as Equation (\ref{eq:tax}) affect the adoption of clean technology across firms and the aggregate pollution. First, we show that correlated distortions have a negative impact on firms' incentives to adopt the clean technology. Second, using a simplified one-sector model with a fixed number of firms and \emph{exogenous} firm-level pollution intensity, we characterize the conditions under which correlated distortions lead to higher aggregate pollution. The core trade-off here is between the scale and technique effects, because the correlated distortions generally cause a decrease in average pollution intensity but an increase in aggregate output.\footnote{All the proofs and additional analytical results can be found in Appendices G.1 and G.2.}

%In this section, we illustrate the basic mechanisms of the model using a simplified version of the model. The core concept in our model is the trade-off between the scale and technique effects. Because the standard result from the misallocation literature where distortions reduce output is inherited by our model [\citet{Guneretal:2008}], the key here is hence to understand how distortions affect the technique effect, and how they shape the trade-off between the scale and technique effects. Specifically, we first show that size-dependent distortions discourage the technology adoption of medium size firms. We then characterize the condition under which reducing distortions lead to lower pollution, or equivalently, the technique effect dominates the scale effect. We leave all the proofs as well as additional analytical properties of the model to Appendices G.1 and G.2.

\textit{Technology Adoption.---}Proposition \ref{prop:friction} shows that correlated distortions discourage firms from adopting of clean technology, especially those with medium levels of TFP.
\begin{proposition}
\label{prop:friction}
When $1-\gamma+\phi_1>0$, a firm chooses to install clean technology only when its productivity is higher than a threshold. Moreover, the threshold when the firm faces distortions $\tilde{z}_f$ is higher than that when the firm is not subject to any distortions $\tilde{z}_n$.
\end{proposition}
%\begin{proposition}
%\label{prop:friction}
%Consider two economies, one (denoted as $f$) with correlated distortions as specified in Equation \ref{eq:tax}, and one (denoted as $n$) without distortions. When $1-\gamma+\phi_1^d>0$, there are two thresholds $\tilde{z}_f$ and $\tilde{z}_n$ such that:
%\begin{enumerate}[(i)]
%    \item
%    For each economy $j = f,n$, firms with $z \leq \tilde{z}_j$ choose the dirty technology, while those with $z > \tilde{z}_j$ choose the clean technology.
%%    \item
%%    In the economy with size-dependent distortions, entrepreneurs with $z \leq \tilde{z}_f$ produce by using the dirty technology, while those with $z > \tilde{z}_f$ produce by using the clean technology.
%    \item
%    When the equilibrium wages $W^f = W^n$, the threshold is higher for the distorted economy, that is, $\tilde{z}_f > \tilde{z}_n$.
%\end{enumerate}
%\end{proposition}

\begin{figure}[t]
\begin{center}
    \includegraphics[width=0.60\textwidth]{./Figures/prod_tax_effect.pdf}
    \caption{Correlated Distortions and Treatment Technologies}
    \label{fig:constantp}
    \floatfoot{Note: There are four profit functions in the figure, $\pi_0^n, \pi_0^f, \pi_1^n$, and $\pi_1^f$. The superscripts $n$ and $f$ denote whether firms face correlated distortions in the economy or not, and the subscripts $0$ and $1$ denote whether firms choose the dirty or clean technology, respectively.}
\end{center}
\end{figure}
\noindent Figure \ref{fig:constantp} illustrates the intuition behind Proposition \ref{prop:friction}. Recall that while the adoption cost $Rk_E$ is independent of $z$, the penalty for not using the clean technology $\xi \pi_0(z)$ is increasing in $z$. %Whether it is worthwhile for firms to pay the installation costs is determined by contrasting the costs with the expected profits loss from inspections by the regulator.
Therefore, the adoption of the clean technology pays off for firms with higher $z$. On the other hand, the elasticity of profits to productivity $z$ is $1-\gamma$ when there are no distortions, and is $1-\gamma+\phi_1$ when distortions exist. Since $\phi_1 < 0$, the slope of the profit functions are flatter, and the threshold where firms switch technologies is higher when they face distortions. As a result, while firms with TFP in the range $[\tilde{z}_n, \tilde{z}_f]$ would choose the clean technology when they are not subject to distortions and choose the dirty technology when they are.\footnote{Of course, when the correlated distortions apply to all firms in the economy, they would affect the equilibrium wage $W$ and hence $\tilde{z}_f$ in equilibrium. We find later in our quantitative exercises that $\tilde{z}_f > \tilde{z}_n$ even with general equilibrium effects considered.} %which may partially offset the direct impact of distortions on technology adoption. However, in our quantitative analysis, this general equilibrium response is not large enough to invalidate the conclusion of Proposition \ref{prop:friction}.

%\noindent Figure \ref{fig:constantp} illustrates Result \ref{prop:friction}. There are four profit functions in the figure, $\pi_0^n, \pi_0^f, \pi_1^n$, and $\pi_1^f$, where the superscripts $n$ and $f$ indicate whether there are size-dependent distortions, and the subscripts $0$ and $1$ represent firms using the dirty and clean technology, respectively. Note that although installation cost $Rk_E$ is fixed, expected loss $\xi \pi_0(z)$ is increasing in $z$. %Whether it is worthwhile for firms to pay the installation costs is determined by contrasting the costs with the expected profits loss from inspections by the regulator.
%Therefore, despite that the fixed installation costs of the clean technology are not justified for firms with lower $z$, they will eventually pay off for those with higher $z$. Because the elasticity of profits to productivity is $1-\gamma$ when there are no distortions, which is larger than that in a market with distortions $1-\gamma-\phi_1$, the distortions decrease the rate at which profits increase with $z$ by $\phi_1$. As a result, the threshold where firms switch technology is higher when there are distortions, causing a positive measure $\Delta z(\tau) = G(\tilde{z}_f) - G(\tilde{z}_n)$ of firms using the clean technology when there is no distortions to switch to dirty technology when facing the distortions.

\textit{Aggregate Pollution.---}To derive a closed-form expression for the relationship between correlated distortions and aggregate pollution, we build a simplified one-sector model that abstracts from the endogenous occupation choice and treatment technology choice.\footnote{Please see Online Appendix G.2 for a detailed description of the simplified model and additional analytical results.} This complements our analysis in Proposition \ref{prop:friction}, which focuses on how distortions affect firms' decision in that the general equilibrium effects are captured. Proposition \ref{prop:agge} summarizes our main findings.
%\textit{Aggregate Pollution.---}To derive a closed-form relationship between aggregate pollution and the size-dependency of the distortions $\phi_1$, we simplify the model along several key dimensions. In particular, we abstract from capital in the economy. We also shut down the two discrete choice decisions in the model by assuming first that household members are exogenously assigned to the roles of entrepreneurs and workers, and second that the firm-level pollution intensity mechanically decreases in production scale by elasticity $-1 < \psi_1 < 0$. Notice that here $\psi_1$ captures both the reduction in intensity from adopting clean technology and during the production stage. Formally, we have the following result.
\begin{proposition}
\label{prop:agge}
In the simplified model, there exists a threshold
\begin{equation*}
    \hat{\phi}_1 = \frac{\psi_1(1-\gamma)}{1-\gamma(1+\psi_1)},
\end{equation*}
such that $E$ is increasing in $\phi_1$ when $\phi_1 < \hat{\phi}_1$, and decreasing when $\phi_1 \geq \hat{\phi}_1$.
\end{proposition}

Proposition \ref{prop:agge} says that the relationship between $E$ and the progressiveness of the distortions $|\phi_1|$ is inverted-U shaped with the turning point $|\hat{\phi}_1|$ increasing in the elasticity of emission intensity to output $|\psi_1|$.\footnote{Here because $\phi_1$ and $\psi_1$ are both negative, the progressiveness and elasticity are stronger when $|\phi_1|$ and $|\psi_1|$ are larger, meaning that the raw values $\phi_1$ and $\psi_1$ are smaller.} Because correlated distortions generally decreases the total output while increasing the average pollution intensity, the intuition behind Proposition \ref{prop:agge} is a race between the scale and technique effects. In particular, the scale effect is larger when the progressiveness of the distortions is stronger (a smaller negative $\phi_1$), while the technique effect is larger when the elasticity of emission intensity to output is larger (a smaller negative $\psi_1<0$). As a result, when $|\phi_1|$ is not too large relative to $|\psi_1|$, the technique effect dominates. Hence an increase in $\phi_1$ (which means a decrease in the progressiveness) causes higher aggregate pollution $E$. Otherwise, the scale effect dominates, and an increase in $\phi_1$ reduces $E$. The average distortions $\phi_0$ (which is positive) always reduces $E$, however, since it does not affect the firm size distribution, and hence only has the scale effect. The overall effect of the correlated distortions is thus a combination of those from $\phi_0$ and $\phi_1$.

%Proposition \ref{prop:agge} basically says that  the aggregate pollution $E$ is decreasing in average distortion $\phi_0$, and the relationship between $E$ and the progressiveness of distortions $-\phi_1$  (notice that $\phi_1<0$) is inverted-U shaped, where the turning point $-\hat{\phi}_1$ is decreasing in $\psi_1$. The intuition goes as follows. Correlated distortions generally decreases the total output while increasing the average pollution intensity. Therefore, the change in aggregate pollution $E$ depends on the race between the scale and technique effects.  An increase in $\phi_0$ leads to a uniform increase in distortions facing all firms. This would increase aggregate output, but induce little reallocation of production factors across firms, and thus have little impact on the average pollution intensity. Therefore, an increase in $\phi_0$ would cause a decrease in $E$.
%
%What is more interesting and more relevant to our core mechanism-the reallocation across firms, is the relationship between $E$ and the progressiveness of distortions $-\phi_1$. Intuitively, the scale effects are larger when the progressiveness of the distortions $-\phi_1$ is larger, and the technique effects are larger when the elasticity of emission intensity to output $\psi_1$ is smaller (notice that $\psi_1<0$). When $-\phi_1$ is not too large relative to $-\psi_1$ (or $-\phi_1< -\hat{\phi}_1$), the technique effect dominates, and an increase in $-\psi_1$ leads to an increase in aggregate pollution $E$. Otherwise, the scale effect would dominate, and an increase in $-\psi_1$ leads to a decrease in $E$.
%
%To sum up, the key message from Proposition \ref{prop:agge} is, when both average distortion $\phi_0$ and progressiveness of distortions $-\phi_1$ are not too large relative to the absolute value of the elasticity of emission intensity to output $-\psi_1$, the technique effect of correlated distortions would dominate the scale effect, which lead to an increase in aggregate pollution.

%The intuition of the above result goes as follows. We have shown that reducing the distortions would increase the production scale and decrease the pollution intensity at the same time. Hence whether aggregate pollution $E$ increases or decreases depends on whether the scale or the technique effect dominates. For a given level of the technique effect $\psi_1$, if output is very responsive to a marginal decrease in distortions, total emissions are likely to increase as a consequence. Because more severe size-dependency of the distortions implies that more productive firms are subject to higher marginal tax, a marginal decrease of the size-dependency of the distortions at a higher level would increase output by more, and hence the increase in aggregate emissions. The reverse is true for lower level of the size-dependency. Since output is more responsive to the size-dependency of the distortions when returns to scale are high, and the technique effect is stronger when $\psi_1$ is lower (remember $\psi_1 < 0$, hence a larger $\psi_1$ implies weaker technology), the scale effect is more likely to dominate when returns to scale are high, and when the technique effect is weak.

%Proposition \ref{prop:agge} basically says that the relationship between the progressiveness of distortions $-\phi_1$ ($\phi_1<0$) and the aggregate pollution $E$ is inverted-U shaped. When $-\phi_1$ is below a threshold -$\hat{\phi}_1$, which itself is decreasing in $\psi_1$, the aggregate pollution $E$ is increasing in $-\phi_1$. Otherwise, the aggregate pollution $E$ is decreasing in $-\phi_1$. The intuition goes as follows. Correlated distortions would generally decrease the total output as well as increase the average pollution intensity, and the change in aggregate pollution $E$ depends on the race between the scale and technique effects.  The scale effects are larger when $-\phi_1$ is larger, and the technique effects are larger when $-\psi_1$ is larger ($\psi_1<0$). When $-\phi_1$ is not too large relative to $-\psi_1$, the technique effects dominate. Otherwise, the scale effects would dominate, and aggregate pollution $E$ is increasing in $-\phi_1$.  In addition,  when $\gamma$ is larger, the scale effects are also larger as the output is more responsive to distortions, and therefore the threshold -$\hat{\phi}_1$ would move to the left.

%Part (i) of Proposition \ref{prop:agge} basically says that when the progressiveness of distortions ($-\phi_1$) is below a threshold (-$\hat{\phi}_1$), the aggregate pollution $E$ is increasing in the progressiveness. Correlated distortions would generally decrease the total output as well as increase the average pollution intensity. As a result, the change in aggregate pollution $E$ depends on the race between the scale and technique effects. If average pollution intensity is very responsive to distortions while the aggregate output is not, total emissions would increase as a consequence. This is true if the progressiveness of distortions is not too high. Since output is more responsive to the distortions when returns to scale ($\gamma$) are high, and the average pollution intensity is less responsive when $\psi_1$ is higher ($-\psi_1$ is lower; in the extreme case of $\psi=0$, the average pollution intensity is not responsive to distortions at all), the threshold -$\hat{\phi}_1$ moves to the left when  $\psi_1$ or $\gamma$ are higher.

%Similar results hold for the average pollution intensity as well, which we summarize formally in the following corollary.
%\begin{corollary}
%\label{prop:aggi}
%In the simple model, the relationship between the average pollution intensity $E/Y$ and the size-dependency of the distortions $\phi_1$ is such that
%\begin{enumerate}[(i)]
%    \item
%    when $\psi_1 < (1-2\gamma)/\gamma$, the average pollution intensity is always increasing in $\phi_1$;
%    \item
%    when $\psi_1 \geq (1-2\gamma)/\gamma$, there exists a threshold
%    \begin{equation*}
%        \overline{\phi}_1 = \frac{(1+\psi_1)(1-\gamma)}{1-2\gamma+\gamma\psi_1},
%    \end{equation*}
%    such that when $\phi_1 < \overline{\phi}_1$, $E/Y$ is increasing in $\phi_1$; and when $\phi_1 \geq \overline{\phi}_1$, $E/Y$ is decreasing in $\phi_1$;
%    \item
%    $\overline{\phi}_1$ is increasing in both $\psi_1$ and $\gamma$;
%%    \item
%%    $\overline{\phi}_1 < \hat{\phi}_1$.
%\end{enumerate}
%\end{corollary}
%Notice that the threshold in Result \ref{prop:agge} %and Corollary \ref{prop:aggi}
%provides a quick way to predict that whether for a certain economy, reducing the size-dependency of the distortions would lead to decrease in aggregate pollution.% and average intensity.

%To clarify the basic mechanics of the model, in this section we analyze a firm's optimization problem when $R$ and $W$ are given exogenously. We prove two results. First, we show that there exists a threshold $\tilde{z}$ such that firms with $z>\tilde{z}$ adopt the clean technology, while firms with $z \leq \tilde{z}$ do not. Second, if we denote the previous threshold in environments with and without size-dependent distortions as $\tilde{z}_f$ and $\tilde{z}_n$ respectively, we show that $\tilde{z}_f > \tilde{z}_n$. The first result states that the returns to scale embedded with the clean technology are only exploited when firms are sufficiently large. The second result states that by introducing size-dependent distortions, some firms that adopt the clean technology when there are no distortions do not have the profit margin to benefit from the clean technology and hence choose to enter the market with the dirty technology. Throughout, we assume $0<\alpha<1$, $0<\gamma<1$, $\phi_0 = 1$, and $1 - \gamma - \phi_1 > 0$. We impose the last inequality because the tax specified in [reference to tax function] is imposed on firm-level TFP $z$. For the benefits of higher talent $z$ (the elasticity of profits to TFP is $1-\gamma$) to always outweigh the costs (the elasticity of tax to TFP is $\phi_1$), $1-\gamma-\phi_1 > 0$ must be satisfied. All proofs are deferred to Online Appendix G.
%
%Lemma \ref{lemma:nofrictionpi} characterizes firms' profit functions in the absence of size-dependent distortions.
%\begin{lemma}
%\label{lemma:nofrictionpi}
%In an economy with no size-dependent distortions, $\pi_0(z)$ and $\pi_1(z)$ are both increasing and linear with respect to $z$. In addition, the slope of $\pi_1(z)$ is steeper than that of $\pi_0(z)$:
%\begin{equation}
%    \frac{\partial \pi_0 (z)}{\partial z} = (1 - \xi)\frac{\partial \pi_1 (z)}{\partial z}, \qquad \forall z \in Z.
%\label{eq:lemma1}
%\end{equation}
%\end{lemma}
%\noindent Lemma \ref{lemma:nofrictionpi} highlights the core trade-off of adopting the clean technology in our model. Although the upfront fixed costs shift the overall profit function down by $R k_E$, the profits of firms using the clean technology will not be confiscated by the regulators. Owing to the constant elasticity between capital and labor, the optimal capital/labor ratio is constant in the absence of factor market distortions. Further, entrepreneurs reap economic rents from managerial talents $z$, and these economic rents increase linearly in $z$ because due to the constant returns to scale production function.
%
%Since the tax in [reference to tax function] is size-dependent in the sense that more talented entrepreneurs are subject to higher distortions, it can be shown that in an economy {with} size-dependent distortions, both $\pi_0(z)$ and $\pi_1(z)$ are concave.
%\begin{corollary}
%\label{lemma:frictionpi}
%Suppose size-dependent distortions are specified as $\max\left\{ 0, 1 - z^{\phi_1} \right\}$ with $1 - \gamma + \phi_1 > 0$; then, $\pi_0(z)$ and $\pi_1(z)$ are both increasing and concave with respect to $z$. In addition, the slope of $\pi_1(z)$ is steeper than that of $\pi_0(z)$:
%\begin{equation*}
%    \frac{\partial \pi_0 (z)}{\partial z} = (1 - \xi)\frac{\partial \pi_1 (z)}{\partial z}, \qquad \forall z \in Z.
%\end{equation*}
%\end{corollary}
%\noindent On the contrary, if the taxes are uniform---meaning that for all $i$, $\tau_{z_i}$s are the same---then $\pi_0(z)$ and $\pi_1(z)$ remain linear.
%
%Since the wage income associated with being a worker is fixed at $W$, the monotonicity of the profit functions implies a threshold $\hat{z}$ for which all household members with talents higher than $\hat{z}$ choose to become entrepreneurs. Put differently, household members choose their occupation according to their comparative advantages. This is the standard result from the Lucas model. We summarize this in Result \ref{prop:threshold}.
%\begin{result}
%\label{prop:threshold}
%There exists a unique threshold $\hat{z}$ such that all household members with $z < \hat{z}$ choose to be workers and those with $z \geq \hat{z}$ become entrepreneurs. Further, $\hat{z}$ is the solution of $W = \pi(\hat{z})$.
%\end{result}

%To explain the intuition, we solve a simplified version of the model here.

\section{Calibration}
\label{sec:calibration}

In this section, we calibrate our model to the Chinese data. We first describe our key calibration strategy, and then explain how our choice of the model parameters matches the key aspects of the Chinese economy at both the aggregate and firm levels.

%\textit{Calibration.}---We calibrate our model to the Chinese data. The model period is set to be one year. Motivated by the empirical evidence in Section \rom{1}, we assume that the pollution intensity function of firms with treatment technology $i$ and production level $y$ is log-linear:
%\begin{equation}
%    \log \left(\frac{e}{y}\right) = \psi_0^{(i)} + \psi_1^{(i)} \log y.
%\label{eq:emsn_cali}
%\end{equation}
%This specification implies that {conditional} on the treatment technology adopted, a ``within-group'' intensity reduction remains as production scale increases. Here, $\psi_1^{(i)}$ captures, for instance, the intensity reduction during the production stage mentioned in Section.\rom{1}.C. Equation (\ref{eq:emsn_cali}) implies that actual emissions are
%\begin{equation}
%    e = E(i,y) = \mathrm{e}^{\psi_0^{(i)}} y^{1+\psi_1^{(i)}}.
%\label{eq:emsn_firm}
%\end{equation}
%Our model features only two broad categories of treatment technologies as opposed to five in the data.
%Recall that we interpret the clean technology in our model as {biological} technology and dirty technology as the remaining types. We assume only the installation of biological technology incurs the fixed cost $R k_E$.\footnote{Qualitatively, assuming that dirty technology also requires a fixed cost will not affect the property of the model. It is equivalent to a decrease in the cost of clean technology, since what matters for firm's decision is the difference of the two costs.}
\subsection{Calibration Strategy}

The main task of our quantitative analysis is to study how correlated distortions affect aggregate pollution by changing the firm size distribution. Therefore, it is important that our model matches well the pollution intensity by firm size, the relative distortions faced by large versus small firms, and the firm size distribution in China. As we mentioned in the previous section, the polluting sector is mapped to the top five polluting industries and the non-polluting sector to the rest of the manufacturing sector, while clean and dirty technologies are interpreted as biological and other (physical plus chemical) technologies.

%\textit{Calibration Strategy.---}The purpose of our quantitative exercise is to study how firm size distribution affects aggregate pollution. Therefore, it is important that our model matches well the firm size distribution in China, the relative distortions faced by large versus small firms, and the pollution intensity of different sized firms. We calibrate our model to the whole manufacturing sector.

Following the evidence in Section \rom{1}.B, the relationship between pollution intensity and firm size is given by Equation (\ref{eq:intensity_size}), and the emissions by firms using different treatment technologies are
\begin{equation*}
    e = \mathcal{E}(i,y) = \mathrm{e}^{\psi_0^{(i)}} y^{1+\psi_1^{(i)}}, i=0,1,
\end{equation*}
where $\{\psi_0^{(i)}, \psi_1^{(i)}\}$ are based on the reduced-form estimates from Equations (\ref{eq:poolols_biological}) and (\ref{eq:poolols_others}).
%where $\{\psi_0^{(i)}, \psi_1^{(i)}\}$ are given by the regression coefficients from (\ref{eq:poolols_dirty}) and (\ref{eq:poolols_clean}).

The firm size distribution in the model is determined by {both} the talent distribution $G(\cdot)$ and the progressiveness of distortions $-\phi_1$. This can be seen from Equation (\ref{eq:zandl}), derived from firms' first order conditions:
\begin{equation}
\label{eq:zandl}
    (1-\tau_z)z^{1-\gamma} = \Phi l^{1-\gamma},
   % z^{1-\gamma+\phi_1} = \Phi l^{1-\gamma}
\end{equation}
where $\Phi$ is a constant. As a result, the progressiveness of the distortions $\phi_1$ has to be calibrated together with the talent distribution $G(\cdot)$. To do this, if we let $\theta_l = y/l$ be the average product of labor, the firm's first order conditions imply
\begin{equation*}
    \theta_l = \frac{y}{l} = \frac{W}{(1-\alpha)\gamma(1-\tau_z)}.
\end{equation*}
Substituting the tax function (\ref{eq:tax}) into the above equation, we have
\begin{equation}
\label{eq:avgl}
    \left(\frac{z_i}{z_j} \right)^{\phi_1} = \frac{\theta_{l,j}}{\theta_{l,i}}.
\end{equation}
Equation (\ref{eq:avgl}) can then be used to identify $\phi_1$ using the variations in average product of labor of firms with different TFP.

%Given a choice of the parameters $\alpha$ and $\gamma$, we can bring our model to the data, and back out the progressiveness of distortions $-\phi_1$. Then we choose $G(\cdot)$ to match the observed firm size distribution. Specifically, let $\theta_l = y/l$ be the average product of labor, the firm's first order conditions imply
%%Because the firm size distribution in the model is affected by {both} the talent distribution $G(\cdot)$ and distortions $\tau_z$, we cannot use only the distribution to calibrate these two elements. To find an additional moment to calibrate $\tau_z$, we follow a similar strategy as the one used by \citet{AdamopoulosRestuccia:2014} when they calibrate the elasticity of distortions to farm size. Specifically, if we let $\theta_l = y/l$ be the average product of labor, then the first order conditions of the firm give us

%Given a choice of the parameters $\alpha$ and $\gamma$, we can compute both $z$ and $\theta_l$ for each firm in the data. Then Equation (\ref{eq:avgl}) can be used to pin down $\phi_1$.
%Since $z$ and $\theta_l$ can all be measured from the data, Equation (\ref{eq:avgl}) can be used to back out $\phi_1$. Notice that because the value of $z$ changes with $\gamma$, and $\gamma$ naturally affects how $G(z)$ shapes the size distribution by altering the level of returns to scale, Equation (\ref{eq:avgl}) needs to be used together with the firm size distribution to calibrate jointly $\phi_1, \gamma$ and $G(z)$.\footnote{Another commonly used calibration strategy in the literature takes the United States as an undistorted economy (which means $\tau_z = 0$ in our model) and calibrates $G(z)$ by matching the size distribution from the tax-free economy to the U.S. data. Taxes are then introduced such that certain moments of the size distribution of the country in interest are matched. See, for example, \citet{Guneretal:2008}, \citet{Bartelsmanetal:2013}, \citet{AdamopoulosRestuccia:2014}, and \citet{BentoRestuccia:2016}, among others. If we calibrate our model in this way, the underlying assumption would be that the talent distribution of the entrepreneurs in China is the same as that in the United States and that all the difference between the size distributions of China and the United States is due to taxes. As is suggested by Figure 2 in \citet{BloomVanReenen:2010}, the evidence regarding this point is mixed. In our opinion, both strategies have their strengths and weaknesses, and face different data limitations. The quantitative results of our study are largely driven by the size and shape of $\tau_z$. Therefore, as long as the alternative calibrations imply similar tax schemes to that used in our benchmark, the quantitative aspects of our results would continue to hold.}

%Given a choice of $\alpha$, $\gamma$ and $\phi_1$, we then choose  $G(\cdot)$ to match the observed firm size distribution in China. Again, from the firm's first-order conditions, we get another useful equation that bridges our model and the data
%We ask the model to match two aspects of the firm size distribution---the total number and share of employment of firms of a certain size. Figure \ref{fig:es} suggests that the distribution of small and medium-sized firms follows approximately a log-normal distribution. However, a log-normal distribution cannot generate the concentration of employment among large firms. The heavy right tail is crucial to our evaluation because these are the firms that are producing with the clean technology. To capture both characteristics, we assume that the distribution consists a log-normal part with total mass $1-g_{max}$ that accounts for the lower end, and an atom with measure $g_{max}$ that accounts for those very large firms.\footnote{This strategy follows \citet{Guneretal:2008} and is popular among macroeconomic studies featuring distributions with fat tails. See, for example, \citet{Castanedaetal:2003}.} The first order conditions of the firms yield

%\begin{equation}
%\label{eq:zandl}
%    (1-\tau(z))z^{1-\gamma} = \Phi l^{1-\gamma}
%\end{equation}
%where $\Phi$ is a constant. Since by the property of log-normal distribution, if $l$ is log-normally distributed, $l^{1-\gamma}$ follows also a log-normal distribution, Equation (\ref{eq:zandl}) suggests that the ``after-tax'' productivity $(1-\tau_z)z^{1-\gamma} \triangleq z'$ follows the same distribution of $l^{1-\gamma}$, which we assume to be a combination of a log-normal part with mean $\mu_z$ and standard deviation $\sigma_z$, and an atom with value $z'_{max}$ and measure $g_{max}$. With $z'$ suitably calibrated, the ``before-tax'' productivity $z$ can then be calculated according to

To calibrate $G(\cdot)$, we cannot use Equation (\ref{eq:zandl}) and the firm-level data to back out the distribution semi-parametrically. The reason is that the endogenous occupation choice in our model makes the distribution of firm TFP in equilibrium also endogenous, making the semi-parametric way of calibration untractable in practice. This is unlike other studies that assume a fixed distribution of firm TFP in equilibrium, for instance \citet{BentoRestuccia:2016}. To keep the calibration process tractable, motivated by Equation (\ref{eq:zandl}), we assume that the distribution of the ``after-tax'' TFP $z' = (1-\tau_z)z^{1-\gamma}$ takes the following form: the majority on the bottom side characterized by a truncated log-normal distribution with log mean $\mu$ and standard deviation $\sigma$, and a top value $z'_{max}$ with a total mass $g_{max}$ representing the large firms. This mass of top managers allows our model to match the observed concentration of employment among large firms, which have lower emission intensities than other firms.\footnote{This strategy is popular among macroeconomic studies on distributions with fat tails. See for example, \citet{Castanedaetal:2003} and \citet{Guneretal:2008}.} We then choose the values of $(\mu, \sigma, z'_{max}, g_{max})$ together with $\gamma$ and $\phi_1$ to match the observed firm size distribution and Equation (\ref{eq:avgl}). The true firm-level TFP $z$ and its distribution $G(z)$ can be backed out by
\begin{equation*}
    z = \left[z'\phi_0^{1/(\gamma-1)}\right]^{\frac{1-\gamma}{1-\gamma+\phi_1}},
\end{equation*}
given a choice of $\phi_0$.
%\begin{equation*}
%    z = \left(z'\phi_0^{1/(\gamma-1)}\right)^{\frac{1-\gamma}{1-\gamma+\phi_1}}
%\end{equation*}

%Since $\tau_z$ is levied based on productivity $z$, we assume that the distribution of ``after-tax'' productivity $z' = (1-\tau_{z})z^{1-\gamma}$ is a combination of two components. The first is a log-normal distribution with mean $\mu$, standard deviation $\sigma$, and total probability mass $1-g_{max}$ that accounts for the majority of small and medium-sized firms. The second is an atom with value $z'_{max}$ and measure $g_{max}$, which accounts for those very large firms.\footnote{This strategy follows \citet{Guneretal:2008} and is popular among macroeconomic studies of wealth distribution. See, for example, \citet{Castanedaetal:2003}.} The ``before-tax'' productivity $z$ can then calculated by
%\begin{equation*}
%    z = \left(z'\phi_0^{1/(\gamma-1)}\right)^{\frac{1-\gamma}{1-\gamma+\phi_1}},
%\end{equation*}
%which gives us $G(z)$.

\subsection{The Choice of Parameter Values}

In total, we need to choose values for 17 parameters: the discount factor $\beta$, depreciation rate $\delta$, production technology parameters $\{\alpha, \gamma \}$, the relative size of the polluting sector $\mu$, treatment technology parameters $\{\psi_0^{(0)}, \psi_1^{(0)}, \psi_0^{(1)}, \psi_1^{(1)}, k_E, \xi \}$, correlated distortions $\{ \phi_0, \phi_1 \}$, and distributional parameters $\{\mu_z, \sigma_z, z'_{max}, g_{max}\}$.

Six of the seventeen parameters can be determined exogenously. We set the depreciation rate $\delta$ to 10\% [\citet{Songetal:2011}]. $\{\psi_0^{(0)}, \psi_1^{(0)}, \psi_0^{(1)}, \psi_1^{(1)} \}$ are taken from the reduced-form estimates from regressions (\ref{eq:poolols_biological}) and (\ref{eq:poolols_others}) in Section \rom{1}.B, giving us $\psi_0^{(0)} = -$3.41, $\psi_1^{(0)} = -$0.36, $\psi_0^{(1)} = -$4.37, and  $\psi_1^{(1)} = -$0.33. In the context of our model, these estimates suggest that, on average, for two firms with the same level of output but different treatment technologies, the firm using the clean technology emits 40\% to 60\% less pollutants than the one with the dirty technology. We set $\mu =$ 0.20, which is the share of manufacturing firms accounted for by the five polluting industries in the CNEC data.\footnote{The share of manufacturing output accounted for by these five industries is 18\%, which is very close as well.}

The remaining parameters have to be calibrated jointly due to general equilibrium feedback. Still, some parameters can be calibrated more or less independently from others, which substantially improves the tractability of the calibration process. In particular, we choose $\beta$ and $\alpha$ to match the capital to output ratio of 1.65, and capital share of 0.5 in China respectively [\citet{Baietal:2006}]. We set $\xi$ and $k_E$ such that the model reflects the average cost to install clean technology in the data. For the five polluting industries, we find in the NGSPS that 57\% of the firms use biological technology, and the average adoption cost of biological equipments is approximately 2.5\% of the average firm output \textit{for these firms}. These two targets help us pin down the values of $\xi$ and $k_E$.

We then calibrate $\{\gamma, \mu_z, \sigma_z, z'_{max}, g_{max}, \phi_1 \}$ such that the model generates a firm size distribution comparable to the data and Equation (\ref{eq:avgl}) holds. To link Equation (\ref{eq:avgl}) to the data, we set $\{z_i, z_j\}$ and $\{\theta_{l,i}, \theta_{l,j}\}$ to their average levels in the top and bottom quintiles respectively. Because the calculation of $z$ depends on $\gamma$, Equation (\ref{eq:avgl}) also defines a relationship between $\phi_1$ and $\gamma$. For the size distribution, we focus on the number of firms and the share of employment in each of the five size groups (Figure \ref{fig:fitness}). In practice, we adjust $\{\gamma, \mu_z, \sigma_z, z'_{max}, g_{max} \}$. For each $\gamma$, we compute the corresponding $\phi_1$ using Equation (\ref{eq:avgl}). We then compute the size distributions in the equilibrium and repeat the process for different combinations of the five parameters until the size distributions in the equilibrium are close to those in the data.

Finally, following \citet{BentoRestuccia:2016}, we calibrate $\phi_0$ such that the average tax burden in the economy matches the value added tax rate Chinese manufacturing firms face, which is 13\%. This gives us the value of $\phi_0 =$ 1.15.\footnote{In general, it is difficult to estimate the average distortions in an economy, because many factors that lead to resource misallocation are either poorly measured or not measured at all. Reassuringly, a change in $\phi_0$ causes a uniform change in firm-level distortions, and therefore has very limited impact on firm size distribution and aggregate pollution intensity. As a result, a different calibration target for $\phi_0$ will not affect the main insight of our paper.} Table \ref{tab:calibration} lists the model parameters, their targets, and the calibrated values.

%We then calibrate $\{\gamma, \mu_z, \sigma_z, z'_{max}, g_{max}, \phi_1 \}$ such that the model generates a firm size distribution comparable to the data and Equation (\ref{eq:avgl}) holds. To map Equation (\ref{eq:avgl}) to the data, we set $\{z_i, z_j\}$ and $\{\theta_{l,i}, \theta_{l,j}\}$ to their average levels in the top and bottom quintiles respectively. Because the calculation of $z$ depends on $\gamma$, Equation (\ref{eq:avgl}) also defines a relationship between $\phi_1$ and $\gamma$. For the size distribution, we focus on the number of firms and the share of employment in each of the five size groups (Figure \ref{fig:fitness}). In practice, we adjust $\{\gamma, \mu_z, \sigma_z, z'_{max}, g_{max} \}$. For each $\gamma$, we compute the corresponding $\phi_1$. Given the value of all other parameters, we then compute the size distributions in the equilibrium. We repeat the process for different combinations of the five parameters until the size distributions in the equilibrium are close to those in the data. Notice that Equation (\ref{eq:avgl}) holds by construction.

%Finally, we calibrate $\phi_0$ such that the average tax burden in the economy equals the 13\% value added tax imposed on Chinese manufacturing firms in the real life. This gives us the value of $\phi_0 =$ 1.15. In general, it is difficult to estimate the average distortions in an economy because many of the factors that lead to resource misallocation are unobservable [\citet{RestucciaRogerson:2013}] and they do not affect the firm size distribution [\citet{Guneretal:2008}]. Our choice of the average tax rate as the calibration target of $\phi_0$ follows \citet{BentoRestuccia:2016}.\footnote{As shown in Result \ref{prop:friction}, the gains in output (and capital and consumption as well) from eliminating size-dependent distortions are increasing in the average level of the distortions. Hence, targeting a higher average ``tax'' rate increases such gains monotonically, which does not affect the results qualitatively. However, quantitatively, the removal of extremely large distortions may lead to a larger increase in output than the decrease in pollution intensity, thereby raising aggregate pollution. Moreover, since Section \rom{4}.B examines the effect of a uniform tax, the results therein could be used to infer how qualitatively the results would change compare to the benchmark case.} Table \ref{tab:calibration} presents the model parameters, their targets, and the calibrated values.

%We use the information on the average products of capital to calibrate the tax function.\footnote{Using the average product of labor yields the same results, since $\phi_i \approx \phi_l$ as shown in Section \rom{2}.B.} Equation [reference to distortion] suggests that the elasticity of $\phi_k$ to $1-\tau_z$ is equal to unity. Therefore, $\phi_1$ is equal to the elasticity of $\phi_k$ to $z$. We thus calculate $\phi_k$ and $z$ according to Section \rom{2}.B, with $R = 0.1$ [\citet{HsiehKlenow:2009}] and the same $\gamma$ used later when we calibrate the model to match the firm size and employment distributions. We then run a regression
%\begin{equation*}
%    \log \phi_{k_i} = \phi_0 + \phi_1 \log z_i + \varepsilon_{z_i},
%\end{equation*}
%which gives us the value for $\phi_1 = -0.03$. %\footnote{The elasticity here is lower than that in Section \rom{2}.B because $z$ is calculated by using $\gamma = 0.93$, while previously we set $\gamma = 0.85$.}
%Given $\phi_1$, we then calibrate $\phi_0$ such that the average tax burden in the economy equals the value added tax imposed on Chinese manufacturing firms in the data, which is 13\%. This gives us the value of $\phi_0 = 1.15$. In general, it is difficult to estimate the average distortions in an economy because many of the factors that lead to resource misallocation are unobservable [\citet{RestucciaRogerson:2013}] and they do not affect the firm size distribution [\citet{Guneretal:2008}]. Our choice of the average tax rate as the calibration target of $\phi_0$ follows \citet{BentoRestuccia:2016}.\footnote{As shown in Result \ref{prop:friction}, the gains in output (and capital and consumption as well) from eliminating size-dependent distortions are increasing in the average level of the distortions. Hence, targeting a higher average ``tax'' rate increases such gains monotonically, which does not affect the results qualitatively. However, quantitatively, the removal of extremely large distortions may lead to a larger increase in output than the decrease in pollution intensity, thereby raising aggregate pollution. Moreover, since Section \rom{4}.B examines the effect of a uniform tax, the results therein could be used to infer how qualitatively the results would change compare to the benchmark case.}
%We choose to target only the value added tax rate and not include other frictions because as is shown in Proposition \ref{prop:friction}, the gains in output (and capital and consumption as well) from eliminating the product market friction are increasing in the average level of frictions. Targeting a higher average ``tax'' rate increases such gains monotonically, and will not affect the results qualitatively. We therefore choose a conservative target for the average level of frictions. We set $A=1$ as normalization.

%The remaining parameters are calibrated jointly by using two layers: an outer layer loops over the parameterization of $G(z)$ and an inner layer solves the model given $G(z)$. In the inner layer, we first approximate $G(z)$ with 5,000 grid points. We then choose $\beta$ and $\alpha$ to match the capital/output ratio of 1.65 and capital share of 0.5 in China, respectively [\citet{Baietal:2006}]. We set $\xi$ and $k_E$ such that the total investment in treatment equipment is equal to 1\% of total output and the proportion of firms adopting clean technologies equals the adoption rate of biological equipment 57\% for the five industries we focus on. The value of $\gamma$ is set such that the difference between the numbers of firms falls into each bin of the employment and firm size distributions generated by the model and those in the data are minimized. More specifically, if we let $s_q$ and $\hat{s}_q$ be the number of firms in each bin $q$ (10 in total) calculated from the data and from the model respectively, $\gamma^*$ solves
%\begin{equation}
%\label{eq:innergamma}
%    \gamma^* = \argmin_{\gamma} \sum_{q=1}^{10} \left( s_q - \hat{s}_q \right)^2.
%\end{equation}
%Note that for each combination of $\{ \mu, \sigma, z_{max}, g_{max} \}$, there is one corresponding $\gamma^*$. Therefore, in the outer layer, we choose the combination of $\{ \mu, \sigma, z_{max}, g_{max} \}$ that minimizes the minimum distances from the inner layer. Furthermore, we require that $\gamma^*$ is the same as that used to calculate $z$. %Put differently, the outer layer picks the minimum of the right hand side values of Equation (\ref{eq:innergamma}) when evaluated at $\gamma^*$.
\begin{table}[t]
\footnotesize
\centering
\begin{threeparttable}
\caption{Calibration}
\begin{tabular}{p{75pt}p{45pt}p{60pt}p{185pt}}
    \hline \hline
    Parameter     &                   & Value         & Target                                                      \\
    \midrule
    Production    & $\delta$          & 0.1000        & Depreciation Rate                                           \\
                  & $\alpha$          & 0.5376        & Capital Share (0.5)                                         \\
                  & $\gamma$          & 0.9300        & Size Distribution\tnote{$\dagger$}                          \\
    \midrule
    Treatment     & $\psi_0^{(0)}$    & $-$3.4144     & Estimates from Regression (\ref{eq:poolols_biological})     \\
                  & $\psi_1^{(0)}$    & $-$0.3636     &                                                             \\
                  & $\psi_0^{(1)}$    & $-$4.3748     & Estimates from Regression (\ref{eq:poolols_others})         \\
                  & $\psi_1^{(1)}$    & $-$0.3288     &                                                             \\
                  & $k_E$             & 4.6000        & Clean Firms: Fixed Cost/Output Ratio (2.5\%)                \\
                  & $\xi$             & 0.2300        & Clean Technology Adoption Rate                              \\
    \midrule
    Distortions   & $\phi_0$          & 1.15          & Average Value Added Tax Rate(13\%)                          \\
                  & $\phi_1$          & $-$0.03       & Elasticity of Average Product to TFP Eq. (\ref{eq:avgl})\tnote{$\dagger$}    \\%Relative Distortions Equation (\ref{eq:avgl})\tnote{$\dagger$}   \\
    Preference    & $\beta$           & 0.8750        & Capital/output Ratio (1.65)                                 \\
    \midrule
    Productivity  & $\mu_z$           & $-$2.4567     & Size Distribution\tnote{$\dagger$}                          \\
                  & $\sigma_z$        & 4.0020        &                                                             \\
                  & $z'_{max}$        & 24855         &                                                             \\
                  & $g_{max}$         & 0.00048       &                                                             \\
                  & $\mu$             & 0.20          & Fraction of Firms in the Polluting Sector                 \\
    \bottomrule
\end{tabular}
\begin{tablenotes}
     \footnotesize{
     \item[$\dagger$]
     Note: The parameters are jointly calibrated.}
\end{tablenotes}
\label{tab:calibration}
\end{threeparttable}
\end{table}

\textit{Discussions.---}Given the calibrated values of the parameters, our model matches well the key aspects of the Chinese economy at both the aggregate and firm levels. Figure \ref{fig:fitness} plots the distributions of the fractions (left panel) and employment shares (right panel) of firms with different sizes for the whole manufacturing sector in the model and in the data.\footnote{Since the distributions of managerial talent are the same in the polluting and non-polluting sectors, and the calibrated value of  $k_E$ is not too large, the firm size distributions for the two sectors in the model are close to each other, which is consistent with data patterns in Figure \ref{fig:es}.} Overall, the model matches the two distributions well. The mean and median of the firm size distribution 53.74 and 19.56, which are not directly targeted in the calibration, match their empirical counterparts of 52.19 and 20 respectively, suggesting that the top managers captures well the upper tail of the firm size distribution.

%\textit{Discussion.}---The calibrated model matches the capital share, capital/output ratio, ratio of average clean equipment cost to average output value for firms using clean technology, proportion of firms adopting the clean technology. %The implied discount rate $\beta$ is less than the values commonly used in the literature. The reason is that we calibrate our model to the steady state as opposed to the balanced growth path. If we assume that the household has log-utility function and the secular growth rate equals to 8\%, the counterpart of the discount rate in this situation is equal to 0.95. We choose to calibrate the model to the steady state because we do not have to make additional assumptions on the functional form of the utility function. Nevertheless, all of the qualitative and quantitative results in this paper hold in the balanced growth path calibration.
%The calibrated value of $\xi$ should be interpreted as the implied ``tax'' rate for using the dirty technology, which combines the effect of many disparate and overlapping policies and adds up to 20.5\% of a firm's annual output value.
%\footnote{While empirical counterpart of the value is difficult to find, the number lies within the range of percentage of plants inspected reported in \citet{BeckerHenderson:2000}. In particular, \citet{BeckerHenderson:2000} report the percentage of firms inspected in four industries in the U.S: Plastics, Wood Furniture, Industrial Organic Chemicals and Metal Containers. The numbers are respectively 2.4\%, 6.8\%, 20\% and 27\%.}

%Figure \ref{fig:fitness} plots the firm size (left panel) and employment share distributions (right panel) in the model and in the data. We plot the match for the whole manufacturing sector. However, since the ex ante distributions of productivity are the same in both the polluting and non-polluting sectors, and that the calibrated level of $k_E$ is small, the matches of the model for the polluting and non-polluting sectors look very similar to those in Figure \ref{fig:fitness}.

%Overall, the model matches the two distributions well given that there are five degrees of freedom. The mean (53.74) and median (19.56) of the firm size distribution, which are not directly targeted in the calibration, match their empirical counterparts of 52.19 and 20 respectively. This suggests that the atom $(z'_{max}, g_{max})$ captures well the upper tail of the firm size distribution.

\begin{figure}[t]
    \begin{center}
    \includegraphics[width=0.45\textwidth]{./Figures/benchfit_fs.eps}
    \includegraphics[width=0.45\textwidth]{./Figures/benchfit_es.eps}
    \caption{Firm Size Distribution: Model versus Data}
    \floatfoot{Note: The firm size and employment distributions of the whole manufacturing sector are reported.}
    \label{fig:fitness}
    \end{center}
\end{figure}

The calibrated value of the span-of-control parameter $\gamma$, which determines the firm-level returns to scale, lies close to the upper end of the range identified by the literature.\footnote{The values of firm-level returns to scale used in the macro literature range from 0.80 [\citet{Guneretal:2008}] to 0.95 [\citet{Bartelsmanetal:2013}]. Estimates from micro studies differ across industries. For example, \citet{OlleyPakes:1996} estimate the value to be between 0.8 and 0.9 for the U.S. telecommunications equipment industry, while the estimates from \citet{Bailyetal:1992} point towards close to constant returns to scale for many manufacturing industries in the U.S.. See also the discussions in \citet{AtkesonKehoe:2005}.} A relatively high value of $\gamma$ is needed to match the employment share of large firms (in particular those with 400+ employees) given the existence of correlated distortions, because we need high returns to scale so that the right amount of resources will be concentrated among firms with high TFP. However, if $\gamma$ is too large, the high demand for factors from the large firms would push up the factor prices, and therefore drive many small firms with low TFP out of the market through endogenous occupation choices. This would jeopardize our model's fit of the fraction of small firms to the data. Our choice of $\gamma =$ 0.93 is a compromise of the two countervailing forces.

%Specifically,  To match the observed employment shares among large firms in the presence of correlated distortions,  %\footnote{Models where these two forces are not antagonistic to each other usually have much better fit. For instance, \citet{Guneretal:2008} study the whole U.S. business sector, which has a narrower employment span and lighter employment concentration. In particular, the largest group they are targeting is firms with more than 100 employees. On the contrary, the largest group we are targeting is 400+ employees, which is significantly larger. In another study, \citet{AdamopoulosRestuccia:2014} assume away the selection mechanism in the model, which, put differently, mutes the general equilibrium feedback through wages. This relaxes the restriction on average talents considerably.}

%The calibrated value of returns to scale $\gamma$ warrants more explanation. The parameter value lies within the empirically estimated range, though close to the upper bound.\footnote{The values previously used in the macro literature range from 0.85 [\citet{AtkesonKehoe:2005}] at the lower end to 0.95 [\citet{Bartelsmanetal:2013}] at the upper end, although in $\gamma$ is not the only source of decreasing returns to scale in [\citet{Bartelsmanetal:2013}]. Please see also the discussion on page 319 of \citet{Bartelsmanetal:2013}. Estimations from micro-level data yield similar results. For example, \citet{OlleyPakes:1996} estimate the value to be between 0.8 and 0.9 for the U.S. telecommunications equipment industry, depending on the particular econometric specification.} The challenges of calibrating the model to simultaneously match the two aspects of the firm size distribution are as follows. To create the concentration of employment among large firms, we need entrepreneurs with very high productivity to face low wages and experience high returns to scale ($\gamma$). Due to the standard property of the \citet{Lucas:1978b} model, low prevailing wage implies that the profits of small firms are also low (Lemma 2 in Online Appendix G). However, higher $\gamma$ raises factor demand from large firms, which pushes up factor prices and drives out small firms with low profits. The latter jeopardizes the fit of the number of small firms in the model. The performance of our model is thus a compromise of these two forces.\footnote{Models where these two forces are not antagonistic to each other usually have much better fit. For instance, \citet{Guneretal:2008} study the whole U.S. business sector, which has a narrower employment span and lighter employment concentration. In particular, the largest group they are targeting is firms with more than 100 employees. On the contrary, the largest group we are targeting is 400+ employees, which is significantly larger. In another study, \citet{AdamopoulosRestuccia:2014} assume away the selection mechanism in the model, which, put differently, mutes the general equilibrium feedback through wages. This relaxes the restriction on average talents considerably.}

%The identification assumption of our benchmark calibration is that {given} taxes $\tau_z$, the talent distribution is identified by the size distribution in the data. Another commonly used calibration strategy in the literature takes the United States as an undistorted economy (which means $\tau_z = 0$ in our model) and calibrates $G(z)$ by matching the size distribution from the tax-free economy to the U.S. data. Taxes are then introduced such that certain moments of the size distribution of the country in interest are matched.\footnote{See, for example, \citet{Guneretal:2008}, \citet{Bartelsmanetal:2013}, \citet{AdamopoulosRestuccia:2014}, and \citet{BentoRestuccia:2016}, among others.} If we calibrate our model in this way, the underlying assumption would be that the talent distribution of the entrepreneurs in China is the same as that in the United States and that all the difference between the size distributions of China and the United States is due to taxes.\footnote{The evidence regarding this point is mixed. See Figure 2 in \citet{BloomVanReenen:2010}.} In our opinion, both strategies have their strengths and weaknesses, and face different data limitations. The quantitative results of our study are largely driven by the size and shape of $\tau_z$. Therefore, as long as the alternative calibrations imply similar tax schemes to that used in our benchmark, the quantitative aspects of our results would continue to hold.\footnote{In an earlier version of this study where endogenous treatment technology choice is not explicitly modeled, we find that this is indeed the case.} We choose to calibrate our model directly to the Chinese economy directly because better empirical evidence (Section \rom{2}.B) is available.

\section{Quantitative Exercises}

In the section, we use our calibrated model to evaluate the effects of alternative policies on aggregate output and pollution. We conduct two sets of counterfactual experiments.  In Experiment (i), we follow the \emph{indirect} approach in the misallocation literature and remove correlated distortions from both sectors by changing $\tau_z$ to $0$.\footnote{In Online Appendix I, we build and calibrate a similar model in which the final consumption is a CES aggregation of the products from two sectors, and use the model to evaluate the impact of removing the correlated distortions from \emph{the polluting sector only}. The main difference between this experiment and Experiment (i) in the main text is that this one induces large reallocation of factors from the non-polluting to the polluting sector. As a result, the reduction in aggregate pollution from removing distortions will be smaller. } This is meant to capture the overall effects of a variety of policies and institutions that reallocate resources from high to low TFP firms. To isolate the importance of the {progressiveness} of distortions, later we replace the progressive taxes with flat taxes by setting $\phi_1=0$, and change $\phi_0$ to keep the total implicit tax revenue constant. In Experiment (ii), we tighten the environmental regulations by increasing $\xi$ to the extent that the fraction of firms using the clean technology is the same as in Experiment (i). We then compare the results from these two experiments, and illustrate the key mechanisms that drive these different results.
%We use the calibrated model as a framework for understanding the effects of the firm size distribution on industrial pollution. We conduct two experiments. In experiment (i), we eliminate all distortions by setting $\tau_z = 0$. This experiment could be interpreted as reductions in inter-regional trade barriers, improvements in transportation infrastructure, decreases in tax burdens, and so on.\footnote{In the main text, we do not study asymmetric changes of $\tau_z$ across sectors, because we feel that it is unlikely that for instance an improvement in infrastructure benefits only the polluting sector. This also explains the reason we model the goods produced by the two sectors are perfect substitutes. However, for the case of asymmetric changes, the assumption of perfect substitutes would lead to overly elastic response in sectoral output. For such scenarios, a moderate elasticity of substitution is needed. We report the results from reducing only the distortions in the polluting sector in a model with CES aggregation (and hence price adjustment) in Online Appendix I. As expected, the main mechanism in our paper remains intact.} Since we are following the \textit{indirect} approach, we are not able to map the size of decrease in $\tau_z$ to observable policy changes. %empirical evidence of the size of the decrease in $\tau_z$ caused by a reduction in some observable distortions of a certain degree is lacking.
%However, as an approximation, the effects of the policy can be assessed by examining the changes in average firm size across steady states, as is done by for example \citet{Guneretal:2008} and \citet{BentoRestuccia:2016}. In experiment (ii), we increase regulation $\xi$ such that the proportion of firms using the clean technology reaches the same level as in experiment (i). In this experiment, we aim to approximate the current environmental policy that punishes firms for using the dirty technology. We then contrast the results of the two experiments to illustrate the different mechanisms and effects of these two types of policies.

%Since along most dimensions, the response from the two sectors are very similar, for each experiment, we will first focus on the polluting sector alone. We then turn to the non-polluting sector by analyzing where its response differs. The overall impact on the whole manufacturing sector is the weighted average of the changes in the two sectors.

\subsection{Counterfactual Experiments}

%\begin{table}[t]
%\footnotesize
%\centering
%\begin{threeparttable}
%\caption{The Effects of Distortions and Regulations}
%\begin{tabular}{p{125pt} >{\centering\arraybackslash}p{75pt} >{\centering\arraybackslash}p{75pt} >{\centering\arraybackslash}p{75pt}}
%    \hline \hline
%    Statistics                      & Benchmark             & (i)                   & (ii)    \\
%    \midrule
%    Aggregate Output                & 100.00                & 129.72                & 100.12  \\
%    Capital                         & 100.00                & 161.25                & 100.14  \\
%    Consumption                     & 100.00                & 123.45                & 100.11  \\
%    Output per Worker (Wage)        & 100.00                & 128.49                & 99.97   \\
%    Output per Firm                 & 100.00                & 298.78                & 109.76  \\
%    Average Talent                  & 100.00                & 224.67                & 109.49  \\
%%    TFP                             & 100.00                & 102.00                & 100.14  \\
%    \midrule
%    Number of Firms                 & 100.00                & 43.42                 & 91.22   \\
%    Mean Size                       & 59.27                 & 137.81                & 65.07   \\
%    Median Size                     & 22.95                 & 41.83                 & 26.72   \\
%    \midrule
%    Aggregate Pollution             & 100.00                & 78.95                 & 90.75 \\
%    Average Intensity               & 100.00                & 60.86                 & 90.65 \\
%    Clean Share                     & 57.40                 & 85.29                 & 85.15  \\
%    Monitoring                      & 20.50                 & 20.50                 & 32.50  \\
%    \bottomrule
%\end{tabular}
%\begin{tablenotes}
%     \footnotesize{
%     \item[$\dagger$]
%     Note: All values are in percentage points except mean and median size, which are the absolute numbers of workers.}
%\end{tablenotes}
%\label{tab:aggregates}
%\end{threeparttable}
%\end{table}

\textit{Removing the Correlated Distortions.}---In Experiment (i), we remove correlated distortions from both sectors by setting $\tau_z = 0$ for all $z$. Column (i) of Table \ref{tab:aggregates} summarizes the results. Overall, removing correlated distortions from both sectors increases aggregate output by about 30\% in both sectors (scale effect), while lowering the pollution intensity of the polluting sector by 42\% (technique effect). Since both sectors share the same distributions of firm-level productivity and distortions, removing correlated distortions induces little reallocation across sectors (composition effect). Therefore, the technique effect from eliminating correlated distortions dominates the scale and composition effects, and aggregate pollution drops by 23\%.\footnote{The decrease in aggregate pollution is consistent with the prediction from Proposition \ref{prop:agge}, given the calibrated values for $\phi_1$, $\gamma$ and the estimated relationship between pollution intensity and firm size.}

%\textit{Removing All Distortions.}---We begin with the effects on the resource allocation in the polluting sector. The core mechanism that generates the results of experiment (i) is the change in size distributions driven by the {general equilibrium wage effect}. Since the tax scheme $\tau_z$ is assumed to be size-dependent, which is larger for more productive firms, the market share of these firms is restrained in the benchmark economy. The elimination of $\tau_z$ removes these constraints. As a result, the previously suppressed factor demand increases considerably. In column (i) of Table \ref{tab:aggregates}, this shows up as a 61\% increase in aggregate capital and a 60\% increase in wages. The size dependency of $\tau_z$ also implies that the situation of small unproductive firms improves to a lesser extent than that of large productive ones. Many small unproductive firms that previously survived because of the low prevailing wage now lose their profit margins. The owners of these firms therefore find it more profitable to work for more productive firms. This {selection} mechanism explains the 121\% increase in the average productivity of firms and 132\% increase in the mean firm size. Furthermore, production is also more concentrated in firms with high productivity {among} the remaining firms. We define the \textit{extensive} margin as the selection of firms into production and the \textit{intensive} margin as the production distribution among the active firms. The first two rows of Table \ref{tab:talentdist}, which report the output share of firms with productivity in different quintiles, show the intensive margin.

\begin{table}[t]
\footnotesize
\centering
\begin{threeparttable}
\caption{The Effects of Correlated Distortions and Environmental Regulations}
%\begin{tabular}{lccccccc}%{p{85pt}>{\centering\arraybackslash}p{50pt}>{\centering\arraybackslash}p{50pt}>{\centering\arraybackslash}p{40pt}}%>{\centering\arraybackslash}p{35pt}}%{lccc}
\begin{tabular}{p{90pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt}>{\centering\arraybackslash}p{40pt} >{\centering\arraybackslash}p{5pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{40pt}}
    \hline \hline
                         & \multicolumn{3}{c}{Polluting}    & & \multicolumn{3}{c}{Non-polluting} \\
    \cmidrule{2-4} \cmidrule{6-8}
    Statistics           & Benchmark & (i)      & (ii)      & & Benchmark     & (i)    & (ii) \\
    \midrule
    Output               & 100.00    & 131.16   & 98.99     & & 100.00        & 129.62 & 100.28 \\
    Capital              & 100.00    & 163.06   & 99.00     & & 100.00        & 161.26 & 100.28 \\
    Consumption\tnote{a} & 100.00    & 123.63   & 100.02    & & 100.00        & 123.63 & 100.02 \\
    Wage\tnote{a}        & 100.00    & 160.00   & 99.96     & & 100.00        & 160.00 &  99.96 \\
    Output per Worker    & 100.00    & 128.25   & 100.15    & & 100.00        & 128.60 &  99.95 \\
    Output per Firm      & 100.00    & 297.62   & 110.17    & & 100.00        & 311.85 & 100.28 \\
    Average Productivity & 100.00    & 221.24   & 111.11    & & 100.00        & 235.39 & 100.00 \\
    \% in Output         & 19.73     & 19.91    & 19.52     & &  80.27        & 80.09  &  80.48 \\
    \midrule
    \# of Firms          & 100.00    & 44.07    & 89.86     & & 100.00        & 41.56  & 100.00 \\
    Mean Size            & 59.98     & 139.19   & 65.98     & & 52.18         & 126.53 &  52.35 \\
    Median Size          & 23.67     & 43.09    & 27.88     & & 18.61         & 34.31  & 18.67  \\
    \midrule
    Pollution            & 100       & 76.67    & 85.82     \\
    Intensity            & 100       & 58.46    & 86.73    \\
%    Pollution            & 100       & 81.66    & 90.60     \\
%    Intensity            & 100       & 62.27    & 91.56     \\
    Clean Share          & 57.77     & 85.61    & 85.10     \\
    Regulation           & 23.00     & 23.00    & 35.50     \\
    \bottomrule
\end{tabular}
\begin{tablenotes}
     \footnotesize{
     \item[$\dagger$]
     Note: All values are percentages except for mean size and median size, which are the numbers of workers.
     \item[a]
     Consumption and wage are not sector specific.}
\end{tablenotes}
\label{tab:aggregates}
\end{threeparttable}
\end{table}

\begin{table}[t]
\footnotesize
\centering
\begin{threeparttable}
\caption{Resource Allocation on the Intensive Margin}
\begin{tabular}{p{65pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt}}
    \hline \hline
    Economy           & $\text{QU}_1$ & $\text{QU}_2$ & $\text{QU}_3$ & $\text{QU}_4$ & $\text{QU}_5$ \\
    \midrule
    \multicolumn{3}{l}{\textbf{Polluting Sector:}} \\
    \ \ \ Benchmark   & 2.79          & 4.26          & 7.45          & 16.81         & 68.70          \\
    \ \ \ Case (i)    & 1.53          & 2.89          & 6.46          & 18.25         & 70.87          \\
    \ \ \ Case (ii)   & 3.02          & 4.63          & 8.01          & 17.79         & 66.55          \\
    \midrule
    \multicolumn{3}{l}{\textbf{Non-polluting Sector:}} \\
    \ \ \ Benchmark   & 2.42          & 3.78          & 6.65          & 15.39         & 71.75          \\
    \ \ \ Case (i)    & 1.30          & 2.49          & 5.68          & 16.97         & 73.57          \\
    \ \ \ Case (ii)   & 2.42          & 3.78          & 6.65          & 15.39         & 71.75          \\
    \bottomrule
\end{tabular}
\begin{tablenotes}
     \footnotesize{
     \item[$\dagger$]
     Note: $\text{QU}_1$ to $\text{QU}_5$ represent the first to fifth quintiles, respectively.}
\end{tablenotes}
\label{tab:talentdist}
\end{threeparttable}
\end{table}

Removing correlated distortions causes large reallocation of production factors across firms, and leads to improved allocation along both the extensive and intensive margins. The \textit{extensive} margin is defined as the selection of firms into production, and the \textit{intensive} margin as the production distribution among the active firms. Specifically, since the distortions $\tau_z$ are higher for high TFP firms in the benchmark economy, a large share of the production factors would flow to these firms after the removal of distortions (intensive margin). The increased labor demand from these firms would drive up the equilibrium wage by 60\%, which would in turn drive many unproductive firms out of the market through the endogenous occupation choices (extensive margin). Due to the changes along the extensive margin, the total number of firms is cut by more than half, and the mean firm size more than doubles. Along the intensive margin, as we see from Table \ref{tab:talentdist}, surviving firms with TFP in the top two quintiles expand at the expense of those with lower TFP. The two panels on the left of Figure \ref{fig:compare_size} show the changes in firm size distributions following the removal of the correlated distortions.
%Firms in the top two groups expand considerably at the expense of firms in the bottom three groups. Therefore, the elimination of $\tau_z$ improves the resource allocation on both the extensive and the intensive margins. For the polluting sector, the overall changes in the size and employment distributions can be seen in the two left-hand panels of Figure \ref{fig:compare_size}.
\begin{figure}[t]
    \begin{center}
    \includegraphics[width=0.45\textwidth]{./Figures/bench_notax_fs.eps}
    \includegraphics[width=0.45\textwidth]{./Figures/bench_reg_fs.eps} \\
    \includegraphics[width=0.45\textwidth]{./Figures/bench_notax_es.eps}
    \includegraphics[width=0.45\textwidth]{./Figures/bench_reg_es.eps}
    \caption{The Effects of Distortions and Regulations on Firm Size Distributions}
    \floatfoot{Note: The distributions of the polluting sector are reported. }
    \label{fig:compare_size}
    \end{center}
\end{figure}

%By defining aggregate $\text{TFP} = {Y}/{(K^{\alpha} N^{1-\alpha})^{1-\gamma}}$,
%%\begin{equation*}
%%    \text{TFP} = \frac{Y}{(K^{\alpha} N^{1-\alpha})^{1-\gamma}},
%%\end{equation*}
%which is the Solow residual following standard growth accounting studies [\citet{HallJones:1999}], we find that the size distributions in models {with} firm selection exert a limited influence over TFP. This finding is consistent with that in \citet{Guneretal:2008}.

The large increase in aggregate output is caused by both an improved allocation of production factors across firms, and a large increase in capital accumulation by about 60\% due to decreased average tax rate from 13\% to 0\%. The increase in capital accumulation plays a much larger role than improved allocation of factors in the increase in aggregate output. Our quantitative exercises in Section \rom{4}.B below illustrate this point further. The large decline in average pollution intensity is also caused by two factors: the improved allocation of production factors, and increased adoption of the clean technology. After the removal of correlated distortions, more production factors move to large productive firms which also have lower pollution intensity. This reallocation lowers the average pollution intensity. Meanwhile, as we discuss in Section \rom{3}.D, correlated distortions lower the elasticity of firm profits to firm TFP $z$, and therefore reduce their incentives to adopt the clean technology. Removing correlated distortions would do exactly the opposite and encourage the adoption of clean technology.

%Now, we examine the effects on pollution. Although aggregate output increases by 31\%, because average pollution intensity decreases by a larger extent of 38\%, aggregate pollution {decreases} by 19\%. In other words, quantitatively, the technique effect dominates the scale effect. The decline in average intensity comes both from the production stage and from the treatment stage. In both stages, changes in the size distribution play a key role. Since production in the no-distortions economy shows a higher degree of concentration in large firms, the fact that the two {within-group} elasticities $\psi_1^{(0)}$ and $\psi_1^{(1)}$ are negative implies a decrease in pollution intensity. %As discussed in Section \rom{1}.C, this could be because larger firms use more productive technologies which just happen to be environmentally friendly. Given that the focus of this paper is at the treatment stage, we capture this relationship in a reduced-form way and we leave a the full-fledged discussion for future research.
%The effect of the size distribution on a firm's choice of end-of-pipe treatment technology works exactly as described in Result \ref{prop:friction}. The elimination of $\tau_z$ increases firms' profits significantly, which strengthens the economic incentive to adopt the clean technology. %\footnote{If we treat international trade costs as one particular example of $\tau_z$, our findings are consistent with those in \citet{Bustos:2011}. More specifically, she found empirically that reduction in tariffs leads to more export revenues which induces upgrade of production technology.} %It is important to bear in mind that both the elimination of $\tau_z$ itself and the subsequent {decrease} of wages contribute to the strengthening of the economic incentive.
%Importantly, by setting the $\psi_1$ in Result \ref{prop:agge} to $\psi_1^{(1)}$, the upper bound of $\hat{\phi}_1$ under our current calibration is $-$0.06. With our calibrated $\phi_1 = -$0.03, which is higher than $\hat{\phi}_1$, the qualitative prediction from Result \ref{prop:agge} is consistent with the quantitative results.

To evaluate the relative contributions of the improved allocation of factors and improved adoption of clean technologies, we consider a hypothetical economy that is identical to our benchmark economy, except that the clean technology shows no improvement in abatement efficiency over the dirty technology, that is, $\psi_0^{(1)} = \psi_0^{(0)}$ and $\psi_1^{(1)} = \psi_1^{(0)}$. First, we remove correlated distortions from this hypothetical economy, which reveals the importance of the improved allocation of factors. We then re-introduce the differences between clean and dirty technologies to the hypothetical economy, so that the resulting equilibrium is exactly the same as that in Experiment (i). This further reveals the importance of the improved adoption of clean technologies. We find that the improved adoption of clean technologies account for 48\% of the decrease in average intensity.

%To evaluate the relative contribution of the reduction in the production and treatment stages, consider a hypothetical world where the clean and dirty technologies are the same, meaning that $\psi_0^{(0)} = \psi_0^{(1)}$ and $\psi_1^{(0)} = \psi_1^{(1)}$. Removing distortions in this economy reduces the aggregate pollution and average intensity by 20\% and 39\%. Restoring the technic specifications of the clean technology further reduces the pollution level and intensity by 17\% and 13\%. Hence on average, about 46\% and 25\% of the decrease in aggregate pollution and average intensity occur in the treatment stage.\footnote{One caveat here is that because we do not model explicitly the technology adoption decision of the firms during the \textit{production} stage due to data limitation, we have implicitly assumed that the relationship between emission and firm size conditional on treatment technology estimated from the data can be extrapolated to other scenarios. In the context of our experiments, this means that the relationships in the world with distortions provide reasonable approximations to those in a world without distortions. As long as this is the case, modeling the technology choice decisions in the production stage explicitly or not would not change the quantitative results.}

Finally, the polluting and non-polluting sectors respond in similar ways to the removal of correlated distortions, and little reallocation is induced across sectors. Recall that the distributions of managerial talent and distortions are the same across these two sectors, and the only material difference between these two sectors is the presence of the fixed adoption cost of clean technology $k_E$ and environmental regulations. Since the calibrated value of $k_E$ is not too large, the differences in the reaction of these two sectors to the removal of distortions are rather limited.

One notable difference is that the mean firm size is somewhat larger in the polluting sector, because the environmental regulations work as an additional selection mechanism. The interaction between the correlated distortions and environmental regulations leads to further efficiency losses in the polluting sector, because average firm size and hence the effective distortions increase. As a result, compared to the non-polluting sector, the removal of correlated distortions leads to a 1.5 percentage points larger increase in output in the polluting sector.

%The non-polluting sector responds to the removal of $\tau_z$ in more or less the same way as the polluting sector. Recall that the only material difference between the polluting and non-polluting sector is the presence of fixed cost $k_E$ and environmental regulations. Hence any difference observed between the two sectors has to stem from how the interaction of these two features affects the decisions of firms. There are two notable differences. First is that the efficiency loss caused by $\tau_z$ is smaller in the non-polluting sector. This can be seen from the fact that the output increase in the polluting sector is 1.5 percentage points smaller in the non-polluting sector. Notice that on average, firms are larger in the polluting sector, because very small firms cannot afford the cost of clean technology. This combines with the size-dependency of $\tau_z$ suggests that on average, the polluting sector is taxed heavier under the same tax scheme, leading to higher efficiency loss. This demonstrates an explicit channel of how environmental regulations amplifies existing distortions in the economy. Second is that although the polluting sector is more selective on the extensive margin, the output per firm is in fact lower. This is because that although the productivity of the very small firms are extremely low, by operating at very small scale, the marginal productivity of labor is in fact higher than other active firms. Leaving them off the market hence reduces the overall efficiency of how the economy uses production factor. Further, Table \ref{tab:talentdist} shows that the additional selection also worsens resource allocation along the intensive margin, which also lowers output per firm.

\textit{Tightening Environmental Regulations.}---In Experiment (ii), we increase environmental penalties $\xi$ to the extent that the fraction of firms using the clean technology increases to the same 56\% as in Experiment (i). The results are shown in Column (ii) of Table \ref{tab:aggregates}. Overall, in the polluting sector, the increase in $\xi$ leads to small decreases in aggregate output, capital stock and equilibrium wage, a sizable change in firm size distribution, and substantial decreases in aggregate pollution and pollution intenisty. On the other hand, the increase in $\xi$ has very limited spillover on both the aggregate outcomes and firm size distribution to the non-polluting sector.

%\textit{Tightening Environmental Regulations.}---The tightening of environmental regulations affects the decisions of firms {directly} through the requirement to adopt technology and {indirectly} through the general equilibrium wage feedback. Unlike in experiment (i), these two effects affect the size and employment distributions at the equilibrium by little, as shown in the two right-hand panels in Figure \ref{fig:compare_size}. There are two reasons behind this. First, large firms that have already adopted the clean technology in the benchmark case are not directly affected by changes in the policy. Second, the installation costs of the clean technology in our benchmark calibration are small. As a result, they do not divert a significant proportion of firms' resources from productive use and hence hardly affect the optimal operating scale of firms. Therefore, in column (ii) of the top panel of Table \ref{tab:aggregates}, the macro aggregates barely change compared with the benchmark case.

The increase in $\xi$ affects the resource allocation in the polluting sector along both the extensive and intensive margins. The most notable changes come along the extensive margin. In our model as well as in the data, it is the small unproductive firms that choose the dirty technology. Therefore, an increase in environmental penalties would mostly fall on small unproductive firms and lower their profits. This would drive a substantial fraction of small firms out of the market through the endogenous occupation choices, and lead to a sizable increase in average firm size. Meanwhile, labor supply increases as the managers of these exiting firms now choose to be workers, which suppresses the equilibrium wage. These changes along the extensive margin are efficiency-improving in the presence of correlated distortions, since the direction of the factor reallocation is opposite to that induced by the correlated distortions.

%However, despite the small {decrease} in wages (0.04\%), selection on the extensive margin {increases} the average size and productivity of active firms. The underlying mechanism here is that an increase in environmental regulations decreases the expected returns from being an entrepreneur, which drives out the least productive firms that cannot afford to install the clean technology. %\footnote{This prediction matches the policy practice in China. For example, during 2004 to 2008, the emissions of major air pollutants as well as industrial production declined significantly. The reductions in per-unit GDP emissions for these pollutants were 35\% for sulfur dioxide, 29\% for black carbon, and 31\% for carbon monoxide [\citet{Linetal:2014pnas}]. During that period, 34 million kW of coal-burning electricity-generating facilities were shut down, which amounts to 6.18\% of total electric production in 2013 (\textit{National Development and Reform Commission} [2009] Decree 4).}
%These household members then choose to become workers, which increases labor supply and suppresses equilibrium wage rates. The remaining large firms benefit from the reduction in wage rates and expand their operating scale, which explains the increases in size and productivity. In Figure \ref{fig:compare_size}, this is graphically reflected as contraction of the smallest group in the firm size distribution and expansion in the other four groups.

While small firms account for a large fraction of firms, they account for only a small fraction of production factors, which implies that the total amount of factors affected by the extensive changes is small. In addition, the polluting sector only account for 20\% of production factors in the benchmark economy. Therefore, the induced decline in equilibrium wage is very small (0.04\%). In addition, our calibrated value of the adoption cost of clean technology $k_E$ is not too large. Therefore, the impact on surviving firms in the polluting sector is also limited, which can be seen in Table \ref{tab:talentdist}. Despite this, the allocation of factors along the intensive margin worsens, in the sense that the share of factors used by the most productive group decreases, because more distorted productive firms benefit less from the wage decline. This offsets the improvement in efficiency from the reallocation along the extensive margin.

%The firms in the four expansion groups are not affected equally, however, as shown in row (ii) in Table \ref{tab:talentdist}. As experiment (i) shows, the most efficient way of allocating talents is to shift production to the most productive firms, increasing the share of output accounted for by firms with productivity in the top quintile. However, row (ii) shows the opposite. Compared with the benchmark case, the proportion of production accounted for by the lower quintiles {increases}, while that of the top quintile {decreases}. Therefore, although strengthening regulation improves the resource allocation on the {extensive} margin, the benefit is countered by exactly the same two channels documented above, namely the amplification of distortions and the loss in efficiency along the intensive margin. %because size-dependent $\tau_z$ limits the extent to which declining wages can benefit firms in the top quintile. %Nevertheless, both these effects are small here because in our benchmark calibration $k_E$ is small.

Since the total amount of factors involved in reallocation along the two margins is small, most of the decrease in average pollution intensity comes from the increased adoption of the clean technology. We repeat the same decomposition exercise as we did in Experiment (i), and find that 94\% and 96\% of the reductions in aggregate pollution and average intensity, respectively, stem from the increased adoption of the clean technology. Since the output of the polluting sector only decreases only slightly in this experiment, the decreases in aggregate pollution and intensity are quantitatively similar at around 14\%.

%Since aggregate output only decreases slightly in this experiment, the decreases in aggregate pollution and intensity are quantitatively similar (i.e., around 10\%). Moreover, because the size and employment distributions change little, most of the decrease comes from the treatment stage. If we repeat the decomposition exercise we just did for experiment (i), 88\% and 91\% of the reductions in aggregate pollution and average intensity stem from the adoption of the clean technology.

The non-polluting sector is indirectly affected in this experiment by the general equilibrium reponse in wage. Since the induced change in equilibrium wage is very small, the increase in $\xi$ has very small effects on both the aggregate outcomes and firm size distribution, even when the products of the two sectors are perfect substitutes.

%The non-polluting sector is barely affected in this experiment, because the only difference in environment it faces is a tiny drop in wage rate. In addition, the distributions of the polluting versus non-polluting sector in Table \ref{tab:talentdist} provide further evidence showing that environmental regulation reduces the efficiency along the intensive margin. The results suggest that when the size of the polluting sector is small, even with perfect substitutes between the goods and hence infinity elasticity of cross-sector substitution, the side effects are negligible.

\textit{Comparing the Two Experiments.}---Both of the above experiments cause the reallocation of factors across firms, but in different ways. Removing the correlated distortions from both sectors improve resource allocation along both the extensive and intensive margins, which push the economy closer to the production possibility frontiers. In contrast, while tightening environmental regulations leads to tougher selection of firms, which pushes the extensive margin closer to its first-best level, the resource allocation across firms along the intensive margin worsens.\footnote{In reality, tightening environmental regulations may cause other undesirable outcomes. For example, extra resources are needed to enforce tightening regulations, and more rent-seeking activities may possibly be induced by the regulations. Due to data limitations, we leave a complete welfare analysis of tightening environmental regulations to future research.} As a result, although the fraction of firms using clean technology is set to be the same in the two experiments, the reduction of aggregate pollution intensity is larger in Experiment (i), because  the larger (and cleaner) firms take a much larger market share compared to Experiment (ii).

%\textit{Comparing the Two Policies.}---These two experiments show that both removing distortions and intensifying regulations affect the resource allocation, but in different ways. Naturally, eliminating the distortions improve resource allocation along both the extensive and intensive margin, since it moves the economy to its first best. On the contrary, although environmental regulation intensifies the selection of firms, which pushes the threshold closer to its level in the first best equilibrium, the average distortions in the economy get amplified the resource allocation along the intensive margin worsens.\footnote{In reality, tightening environmental regulations may cause other undesirable outcomes. For example, extra resources are needed to enforce tightening regulations, and more rent-seeking activities may possibly be induced by the regulations. Due to data limitations, we leave a complete welfare analysis of tightening environmental regulations to future research. }As a result, although in both experiments the share of firms using clean technology is the same, the reduction of aggregate pollution during the treatment stage is larger in experiment (i) because firms adopting clean technology takes a larger market share.

%In particular, removing distortions increases labor {demand} from firms, while intensifying regulations increases labor \textit{supply} from household members. Because distortions are size-dependent, large firms benefit more from their removal. On the contrary, an increase in labor supply suppresses equilibrium wages, thereby potentially raising firms' profits. However, the actual extra profits firms earn need to take out the effect of distortions. As before, owing to the size dependency of the distortions, large firms benefit less from the decrease in wages. This explains the reason why the intensive margin improves when distortions are eliminated but worsens when stronger regulations are imposed. %Put differently, distortions shape the effect of environmental policies on different firms. %This finding complements the analysis by \citet{Bustos:2011}, who empirically studies the extent to which the elimination of tariffs---a particular form of size-dependent distortion---influences technology upgrading. The results of our first experiment are consistent with hers. Further, we examine how the economy is affected if the government implements policies that target technology adoption in the presence of size-dependent distortions.

%The lesson we learn from the quantitative exercises is that if resources are devoted to smoothing the frictions in the economy instead of being used to intensify regulations, reductions of pollution in both the production and treatment stage could arise naturally as equilibrium outcomes.
% Several recent studies argued that industrial pollution in China comes from the unanticipated responses of economic agents to different policies and institutional features.\footnote{For instance, \citet{Jia:2014} studies the role of the political incentives and career concerns of provincial governors, \citet{Jiangetal:2014} emphasize differential policy treatments to firms with different ownership rights structures, and \citet{Kahnetal:2015} and \citet{Caietal:2016} estimate the difference in the reduction efforts from provincial governments on upstream and downstream counties along the rivers in response to the pollution reduction mandates imposed by the central government.} In reality, the effects of regulatory policies such as government campaigns are often ineffective. These policies suffer from the challenge in monitoring small firms operating off the radar and the quick rebound of pollution level when those campaigns end. Indeed, our results suggest that policies and institutional features could work to amplify other existing distortions in the economy and removing such distortions may well provide a different path to growth that is more sustainable. At the same time, the elimination of those distortions could also reduce the unanticipated environmental consequences of these policies and institutional features. That said, it could be that reducing economic distortions costs more, meaning that strengthening regulations pays off to a larger degree in the short run.\footnote{Another concern is that if small firms have higher growth rates, the removal of size-dependent distortions could hamper economic growth in the long run. However, the recent study of India's manufacturing sector by \citet{Martinetal:2017} suggests that as in the United States, larger and younger factories grow more quickly and create more jobs than do smaller, older factories. If the pattern is the same with China, then it posits less of a concern to the long-run growth prospect of the Chinese economy.} However, the discussion of the implementation of specific policies is beyond the scope of the current study.

%\textit{Relationship to Trade Models.---}The implication of our model on resource allocation along the intensive margin distinguishes us from many studies in international trade that use the \citet{Melitz:2003} model to study the environmental consequences of trade liberalization. A major difference between the distortions in our model and the trade costs in a standard Melitz model is that exporting is an endogenous choice of firms which brings extra revenues. A productive firm can choose not to pay the trade costs if it decides to stay domestic. As a result, though trade costs are a specific type of distortion that only affect exporting (productive) firms, they are different from the distortions in our model which are imperative.\footnote{The difference is not caused by our use of a \citet{Lucas:1978b} model instead of a \citet{Melitz:2003} model. Rather, it is because distortions in our model are a pure cost, while in the \citet{Melitz:2003} model, firms pay trade costs to access to foreign markets, which bring additional revenues. If instead we model in the \citet{Melitz:2003} model similar ``deadweight-loss'' distortions, we would get similar results. In fact, trade costs that lead to higher average factor product are indeed captured as distortions by our model.}
%
%Removing the distortions in our model is the same as trade liberalization in these models. However, intensified regulations yield different implications on resource allocation in our model comparing to those that arise from the standard Melitz model. The reason is that stronger regulation works as an increase in the entry costs, which foreign markets serve as a shelter to exporting firms. Hence, less productive firms are affected more. This illustrates how domestic distortions are conceptually different from trade costs. We believe that such difference is important because the scenario where regulations are strengthened while keeping the distortions is closely related to environmental policies in practice. In this sense, our findings compliment those in the trade literature by pointing out a different way that domestic distortions could affect the economy.\footnote{As an example, we compare our results with those from \citet{Bustos:2011}. The results of our first experiment are consistent with both the theoretical and empirical results of trade liberalization on technology adoption in her paper. We are further able to examine how the economy is affected if the government implements policies that target technology adoption in the presence of size-dependent distortions.}

%\textit{Partial Equilibrium Interpretations.}--Some readers may be concerned that in reality, reduction of labor demand in one industry does not necessarily cause a decline of wages across the whole economy, thereby challenging the main mechanism of our model. We argue that the general logic carries well onto other model environments. First, what eventually matters in our model is not the wage {per se}, but the profits which are revenue less factor costs. In a partial equilibrium version or a \citet{Melitz:2003}-style version of our model where the wage and demand are both fixed, $\tau_z$ will push up the equilibrium price that the firms could charge. This is because with size-dependent $\tau_z$, the supply from the large productive firms is suppressed, where competition among consumers bids up the equilibrium price. Much of the quantitative results are still maintained in these situations. We choose to work with a general equilibrium Lucas model because it enables us to discuss the changes in output level. In fact, the results we present in the general equilibrium setting are stronger than those in the partial equilibrium counterparts. A decrease in the {average intensity} could well be accompanied by an increase in {aggregate pollution} if the output increases by too much. Second, if the labor markets in different industries are segmented, then a decrease in labor demand will in fact trigger a wage drop. Third, the nature of our comparison is to contrast economies {with} and {without} frictions as a whole, not any particular industry. Therefore, as long as the product market frictions exist at the aggregate level, wages will be {lower} in cases with frictions.

\subsection{Effects of the Progressiveness of the Distortions}
\label{sec:progressiveness}

In Experiment (i), by setting $\tau_z = 0$, we change both the progressiveness and and the average level of firm-level distortions. To isolate the importance of the progressiveness of $\tau_z$, we solve a version of the model where the progressive taxes are replaced with uniform taxes by setting $\phi_1 =$ 0, and change $\phi_0$ to keep the total implicit tax revenue constant.\footnote{This strategy has a long tradition in the public finance and development economics literature. See for example, \citet{Ventura:1999}, \citet{ConesaKrueger:2006}, \citet{Bhattacharyaetal:2013} and \citet{BentoRestuccia:2016}, among others.} %To this end, the experiments resemble those in the taxation literature, where progressive taxes are compared with flat taxes. If we interpret $\tau_z$ literally as a tax, it would be unwise to suggest eliminating the whole tax system on manufacturing sector. This exercise could be interpreted as a counterfactual on policies that remove the size-dependent feature of current taxation policies in China.
Not surprisingly, the implied uniform tax rate $\tau_z =$ 18\% is higher than the average tax rate of 13\% under the progressive tax system. Column (i') in Table \ref{tab:uniform} summarizes the results of this experiment and compares them to those from Experiment (i). A comparison of Column Benchmark and Column (i') reveals the effects of a change in progressiveness of $\tau_z$, while a comparison of Column (i') and (i) highlights the effects of a uniform change in tax rates across firms.

%To further isolate the effect of the firm size distribution, we solve a version of the model where uniform distortions are imposed on all firms in the economy. More specifically, in this exercise, we set values of $\tau_z$ such that the total amount of taxes collected is the same as in the benchmark case with size-dependent $\tau_z$. This experiment reveals the quantitative effect of the size dependency of $\tau_z$.\footnote{To this end, the experiment resembles those in the taxation literature, where progressive taxes are compared with flat taxes. See, for example, \citet{Ventura:1999} and \citet{ConesaKrueger:2006}.} %To this end, the experiments resemble those in the taxation literature, where progressive taxes are compared with flat taxes. If we interpret $\tau_z$ literally as a tax, it would be unwise to suggest eliminating the whole tax system on manufacturing sector. This exercise could be interpreted as a counterfactual on policies that remove the size-dependent feature of current taxation policies in China.
%The implied tax rate $\tau_z =$ 18\% is higher than the average tax rate in the size-dependent case, 13\%. Table \ref{tab:uniform} summarizes the results of the experiment. Columns benchmark and (i) present the results of the benchmark calibration and those from setting $\tau_z = 0$ from Table \ref{tab:aggregates}. We label the uniform tax case as (i'). Comparing the benchmark with (i') allows us to assess the effect of the size dependency of $\tau_z$. Similarly, a comparison of (i') with (i) reveals the effect of levying a flat tax. Note that because a uniform $\tau_z$ imposed on all firms does not change the extensive margin compared with the zero $\tau_z$ situation, the measures of average talents, number of firms, and mean/median firm size are unaffected. This means that the effects of the uniform tax all work through the intensive margin.
\begin{table}[t]
\footnotesize
\centering
\begin{threeparttable}
\caption{The Effects of the Progressiveness of Distortions}
\begin{tabular}{p{90pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt}>{\centering\arraybackslash}p{40pt} >{\centering\arraybackslash}p{5pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{40pt}}
    \hline \hline
                         & \multicolumn{3}{c}{Polluting}    & & \multicolumn{3}{c}{Non-polluting} \\
    \cmidrule{2-4} \cmidrule{6-8}
    Statistics           & Benchmark & (i')      & (i)      & & Benchmark     & (i')    & (i) \\
    \midrule
    Output               & 100.00    & 108.15   & 131.16    & & 100.00        & 106.87 & 129.62 \\
    Capital              & 100.00    & 110.91   & 163.06    & & 100.00        & 109.63 & 161.26 \\
    Consumption\tnote{a} & 100.00    & 106.57   & 123.63    & & 100.00        & 106.57 & 123.63 \\
    Wage\tnote{a}        & 100.00    & 108.76   & 160.00    & & 100.00        & 108.76 & 160.00 \\
    Output per Worker    & 100.00    & 105.74   & 128.25    & & 100.00        & 106.03 & 128.60 \\
%    Output per Firm      & 100.00    & 245.39   & 297.62    & & 100.00        & 257.12 & 311.85 \\
   \midrule
      \# of Firms          & 100.00    & 44.07    &44.07     & & 100.00        & 41.56  & 41.56 \\
    Mean Size            & 59.98     & 139.19   & 139.19     & & 52.18         & 126.53 &  126.53 \\
    \midrule
    Pollution            & 100       & 70.04    & 76.67     \\
    Intensity            & 100       & 64.78    & 58.46     \\
%    Pollution            & 100       & 73.27    & 81.66     \\
%    Intensity            & 100       & 67.78    & 62.27     \\
    Clean Share          & 57.77     & 73.33    & 85.61     \\
    \bottomrule
\end{tabular}
\begin{tablenotes}
     \footnotesize{
     \item[$\dagger$]
     Note: All of the values are percentages except for mean firm size, which are the numbers of workers.
     \item[a]
     Consumption and wage are not sector specific.}
\end{tablenotes}
\label{tab:uniform}
\end{threeparttable}
\end{table}

As we can see from Table \ref{tab:uniform}, removing the progressiveness of $\tau_z$ increases aggregate output, capital, consumption and output per worker. However, the 6\% to 11\% increases here are rather moderate compared to the 20\% to 60\% increases when all the distortions are eliminated. These differences are mainly caused by the relatively high level of the uniform tax rate of 18\% in Experiment (i'), which discourages capital accumulation and results in much lower capital stock compared to the case of $\tau_z =$ 0.\footnote{Our results on aggregate output also appear rather moderate compared to those in \citet{RestucciaRogerson:2008}, who control for the capital accumulation effects in their quantitative exercises. This is mainly because that the distortions in their paper cause large rank reversals between firm TFP and size, while the distortions in our paper preserve ranks. Please see the related discussions in \citet{Hopenhayn:2014}.}

However, removing the progressiveness of $\tau_z$ causes a large change in firm size distribution, and therefore large declines in aggregate pollution and pollution intensity. In fact, the resulting number of firms and mean firm size are the same between Experiment (i) and (i'), since the only difference between them is a uniform increase in tax rates across firms. The average pollution intensity decreases by 35\% when the progressiveness of $\tau_z$ is removed, compared to a 42\% decline when all the correlated distortions are eliminated. Again, the decline in average pollution intensity is caused by both an improved resource allocation and improved adoption of the clean technology among firms. However, the second channel is not as strong as in Experiment (i) (16 versus 28 percentage points), since the increase in the slope of the firm's profit functions is not as large as in Experiment (i). Moreover, since the output of the polluting sector only increases by 8\% in this experiment compared to 30\% in Experiment (i), the 30\% decline in aggregate pollution here is larger than the 23\% in Experiment (i).

In sum, our findings imply that it is the progressiveness of distortions, rather than the average level of firm-level distortions, that plays a dominate role in amplifying aggregate pollution and pollution intensity.

%The aggregate output, capital, consumption, output per worker, and output per firm in case (i') all increase compared with the benchmark case, which is consistent with the findings on size-dependent distortions in the literature. What is different here is that since the uniform tax rate needed to generate the same tax revenue is relatively high (the tax rate for the largest firm in the size-dependent case is 25\%), the tax itself still results in considerable output loss. The source of this output loss in our model works mainly through decrease in capital accumulation. However, for average pollution intensity, much of the reduction is achieved through the elimination of the size dependency of $\tau_z$. The adoption rate of the clean technology increases by 16 percentage points in case (i'), which is 56\% of the total increase resulting from the complete elimination of $\tau_z$. Moreover, 85\% of the total decline in average pollution intensity in the zero $\tau_z$ case is achieved by simply removing the size-dependent feature of $\tau_z$. Here, the non-polluting sector again behaves in a similar way, and hence the discussion is omitted.

%\textit{Discussion.}---The finding that the size dependency of $\tau_z$ affects economic efficiency moderately is consistent with the results of \citet{Hopenhayn:2014}. In particular, he shows that the size dependency of the policy does not necessarily imply large distortions. What matters for the size of the distortion is the total amount of resources that are affected, not to which firms these resources belong. Studies such as \citet{RestucciaRogerson:2008} and \citet{Guneretal:2008} find that size-dependent policies affect economic efficiency more than flat taxes because such policies happen to affect a large amount of resources. The fact that we impose fairly large $\tau_z$ in the flat tax case explains why removing the size dependency of $\tau_z$ hardly improves the efficiency of the economy.

%However, the size dependency of $\tau_z$ does matter for pollution, since it affects the {size distributions}. The reduction in pollution comes from two sources: the mechanical decrease in the production stage and the increase in the adoption rate of the clean technology since the intensity of regulation $\xi\pi^0(z)$ is increasing in $z$. More specifically, because the elimination of the size dependency of $\tau_z$ improves the resource allocation on the intensive margin, which increases the profits of large firms, the punishment from violating environmental regulations actually increases, thereby giving firms higher incentives to install clean technology.
%Since the size distribution of the firm is purely affected by the {size-dependency} of $\tau_z$ and {not} by the size of the distortions (in the definition of \citet{Hopenhayn:2014}), although size-dependency does {not} necessarily imply significant efficiency loss, it {necessarily} results in aggravation of pollution.

%\subsection{Environmental Regulations and Size Distribution}
%
%In our benchmark calibration, $k_E$ is about 2.5 times the equilibrium wage of a typical worker. This relatively small value limits the extent to which environmental policy can affect the real economy. In this section, we show that when $k_E$ is increased to 10 times the value used in the benchmark calibration, environmental policy influences the resource allocation markedly. We consider two scenarios. In the first case, we again solve the model with all the parameters remaining at the benchmark calibration level, but increase $k_E$ to 41.5. With no changes in regulation intensity, the adoption rate of the clean technology drops. In the second case, we increase $\xi$ such that the adoption rate in the benchmark case (57\%) is restored. We label these two experiments (iii) and (iv).
%\begin{table}[t]
%\footnotesize
%\centering
%\begin{threeparttable}
%\caption{The Effect of Regulations: Higher $k_E$}
%\begin{tabular}{p{125pt} >{\centering\arraybackslash}p{75pt} >{\centering\arraybackslash}p{75pt} >{\centering\arraybackslash}p{75pt}}
%    \hline \hline
%    Statistics                      & Benchmark             & (iii)                 & (iv)    \\
%    \midrule
%    Aggregate Output                & 100.00                & 100.00                & 100.41  \\
%    Capital                         & 100.00                & 100.28                & 101.45  \\
%    Consumption                     & 100.00                & 99.94                 & 100.21  \\
%    Output per Worker (Wage)        & 100.00                & 100.00                & 99.63   \\
%    Output per Firm                 & 100.00                & 100.00                & 187.70  \\
%    Average Talent                  & 100.00                & 100.00                & 184.14  \\
%%    TFP                             & 100.00                & 100.00                & 100.81  \\
%    \midrule
%    Number of Firms                 & 100.00                & 100.00                & 53.50   \\
%    Mean Size                       & 59.27                 & 59.27                 & 111.65  \\
%    Median Size                     & 22.95                 & 22.95                 & 63.24   \\
%    \midrule
%    Aggregate Pollution             & 100.00                & 125.64                & 87.56 \\
%    Average Intensity               & 100.00                & 125.64                & 87.20 \\
%    Clean Share                     & 57.40                 & 8.89                  & 57.06  \\
%    Monitoring                      & 20.50                 & 20.50                 & 73.50  \\
%    \bottomrule
%\end{tabular}
%\begin{tablenotes}
%     \footnotesize{
%     \item[$\dagger$]
%     Note: All values reported are in percentage points except mean and median size, which are the absolute numbers of workers.}
%\end{tablenotes}
%\label{tab:aggregates_ke}
%\end{threeparttable}
%\end{table}
%
%Table \ref{tab:aggregates_ke} presents the results of experiments (iii) and (iv). For case (iii), we first note that the size distributions and efficiency of the economy remain virtually the same as in the benchmark case. However, the clean technology adoption rate falls to 9\% since a large number of firms now find it unprofitable to use the clean technology and higher pollution thus follows. Because the benchmark case can be understood as a decrease in the price of the clean technology from case (iii), this result suggests that subsidizing clean technologies can serve as a substitute for stronger regulations.
%
%If we increase the regulations to the level such that the original technology adoption rate is restored, as opposed to the results of case (ii) where the size distributions and resource allocation are only mildly affected, the effects of the regulation in experiment (iv) are much larger. Figure \ref{fig:envpolicy} shows the size and employment distributions for case (iv).
%\begin{figure}[t]
%    \begin{center}
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/PrelimResults/sizedist_bench_highcost.eps}
%    \includegraphics[width=0.45\textwidth]{D:/PaperDrafts/Pollution/PrelimResults/eshare_bench_highcost.eps}
%    \caption{Environmental Policy and Size Distribution}
%    \label{fig:envpolicy}
%    \end{center}
%\end{figure}
%In contrast to the size and employment distributions in the benchmark case, all the firms in the first group are driven out. This fact is also reflected in column (iv) of Table \ref{tab:aggregates_ke} as increases in average talent and in the mean/median size of the firms. Therefore, environmental policy {improves} the resource allocation on the {extensive} margin by forcing small unproductive firms to exit the market. However, the gains from these improvements are limited because the allocation of resources on the {intensive} margin worsens (see Table \ref{tab:talentdist_ke}).
%\begin{table}[t]
%\footnotesize
%\centering
%\begin{threeparttable}
%\caption{Resource Allocation on the Intensive Margin: High $k_E$}
%\begin{tabular}{p{65pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt} >{\centering\arraybackslash}p{45pt}}
%    \hline \hline
%    Economy     & $\text{QU}_1$ & $\text{QU}_2$ & $\text{QU}_3$ & $\text{QU}_4$ & $\text{QU}_5$ \\
%    \midrule
%    Benchmark   & 2.69          & 4.19          & 7.29          & 16.55         & 69.28 \\
%    Case (iv)   & 4.44          & 6.63          & 10.98         & 21.86         & 56.10 \\
%    Case (i)    & 1.50          & 2.83          & 6.34          & 18.06         & 71.27 \\
%    Case (ii)   & 2.93          & 4.47          & 7.77          & 17.37         & 67.46 \\
%    \bottomrule
%\end{tabular}
%\begin{tablenotes}
%     \footnotesize{
%     \item[$\dagger$]
%     Note: $\text{QU}_1$ to $\text{QU}_5$ represent the first to fifth quintiles, respectively.}
%\end{tablenotes}
%\label{tab:talentdist_ke}
%\end{threeparttable}
%\end{table}
%An efficient allocation on the intensive margin is achieved in the no-size-dependent $\tau_z$ case [case (i)]. Instead of an allocation of production toward larger firms as in the efficient case, here the production of firms in the bottom four quantiles expands significantly at the expense of the production share of the most productive firms. As a result, much of the gains in economic efficiency are offset by the worsening of the resource allocations on the intensive margin. Qualitatively, this result is the same as in case (ii). However, by comparing cases (ii) and (iv), we see that the larger fixed cost $k_E$, the higher is the degree of resource misallocation on the intensive margin . Although the level of industrial pollution lowers, the reduction could be much larger if the allocation on the intensive margin also improves, as in the case of the elimination of $\tau_z$.

%\textit{Alternative Calibrations.}---Describe and analyze the results using the U.S calibration.
%
%\textit{Capital Market Frictions.}---Describe and analyze the results with capital market frictions.

\section{Conclusions}

This paper proposes a novel explanation for the industrial water pollution problem in China. In our theory, large productive firms are more likely to use clean technology and have lower pollution intensity, largely due to the fixed adoption cost of clean technology and imperfect environmental regulations. Distortions that increase with firm TFP reallocate production factors away from large productive firms, therefore reduce aggregate output while increasing aggregate pollution intensity. We support our theory by firm-level data on production, emissions, and pollution treatment technologies from China. Quantitatively, we find that distortions correlated with firm TFP reduce aggregate output by about 30\% and increase aggregate pollution by about 20\%.

Our findings have novel implications for both resource misallocation and environmental pollution. They imply that the welfare losses from misallocation include not only lower economic output, which is the focus of previous studies, and but also more pollution. This could have important welfare consequences given the severity of pollution in China and other developing countries. Our findings also challenge a popular view that economic growth necessarily comes at the cost of more pollution in developing countries. They suggest that a growth path with both higher output and lower pollution is attainable, if policy and institutional distortions can be removed in a cost effective way.

% Despite our novel findings, an important policy-related question remains: are there empirically observable size-dependent distortions that can generate quantitatively large effects on both firm size and pollution? Identifying observable policy distortions would also make it easier to collect more information on the costs of particular policy instruments, thereby allowing us to conduct a more complete cost-benefit analysis. Along this direction, our own work \citet{AGpaper} studies how migration and land policies affect the pollution from the overuse of agricultural chemicals in China.

The analytical framework in this paper can be extended to other pollutants, sectors, and countries. For other pollutants, China also has a severe air pollution problem, and casual observations reveal that small and medium-sized firms in high-polluting industries such as steel contribute dispropotionately to air pollution in China. For other sectors, \citet{AGpaper} find that misallocation caused by migration and land policies is an important cause of agricultural pollution in China. For other countries, \citet{HsiehKlenow:2014} and \citet{BentoRestuccia:2016} find evidence of distortions that are highly correlated with firm TFP in a number of developing countries, while \citet{Dasguptaetal:1998} find that small firms have higher air pollution intensity in Brazil and Mexico. It would be interesting to assess the impact of distortions on pollution in these countries.

%Another limitation of this study is that we restricted our attention to examining industrial water pollution in China. Researchers could consider extending our analysis to other pollutants, sectors, and countries. For instance, China also has a severe air pollution problem, and it has overtaken the United States as the world's largest emitter of greenhouse gases. Casual observations reveal that small and medium-sized firms in high-polluting industries such as steel are an important source of sulfur dioxide, hazardous haze, and greenhouse gas emissions in China. For other developing countries, \citet{HsiehKlenow:2014} and \citet{BentoRestuccia:2016} find that size-dependent distortions negatively affect average firm size, while \citet{Dasguptaetal:1998} find that small firms have higher air pollution intensity in Brazil and Mexico. It would thus be interesting for future research to assess whether size-dependent distortions amplify pollution in these countries as well.

%Finally, studies of urban economics and macroeconomics that adopt models that have rich macro and spatial implications are rising (\citet{Combesetal:2012}, \citet{DesmetRossihansberg:2013b}, \citet{DesmetRossihansberg:2014}). Industrial pollution naturally calls for interdisciplinary studies that analyze the environmental consequences of economic activities from a spatial perspective. Similar to the case presented herein, the agglomeration of pollution sources would be likely to change the costs and benefits structure of current studies where pollution is off the screen. We leave all these extensions to future studies.

%\newpage
%\begin{appendices}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%% Adjust the Layout of the Appendix %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\setcounter{equation}{0}
%\setcounter{lemma}{0}
%\setcounter{corollary}{0}
%\setcounter{proposition}{0}
%\setcounter{figure}{0}
%\setcounter{table}{0}
%\renewcommand{\thesection}{\textnormal{\Alph{section}}}
%\renewcommand{\thesubsubsection}{\textnormal{\thesubsection\arabic{subsubsection}}}
%\titleformat*{\subsubsection}{\it \center}
%\renewcommand{\theequation}{\thesubsection\arabic{equation}}
%\renewcommand{\thefigure}{\thesubsection\arabic{figure}}
%\renewcommand{\thetable}{\thesubsection\arabic{table}}

%%%%%%%%%%%%%%%%%%%%%%%%%%% Beginning of the Appendix %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{center}
%\textsc{Appendix}
%\end{center}
%
%\subsection{Accounting Exercises}
%
%This appendix provides robustness calculations of the accounting exercises.
%
%\textit{Estimation Strategies.}---Ideally, we would like to have the pollution intensity over firm size. However, such data do not exist since the NGSPS only reports total value of production and total amount of pollution. Therefore, we would need to construct pollution intensity over the number of employees. We use the CNEC for this purpose. In particular, we estimate the corresponding bins for production for each employment bin from the U.S data.
%
%SUSB reports firm size in 22 bins. For the U.S size bins, we construct the corresponding production bins to be used in NGSPS. CNES is used to bridge the employment bins (SUSB) to the production bins (NGSPS).
%
%\begin{enumerate}
%  \item
%  \textit{Non-parametric}:
%    \begin{itemize}
%    \item
%    For each U.S employment bin, we compute the 1st quartile and 3rd quartile production levels for Chinese firms within that employment bin. The two quartiles are used as the lower and upper bounds for the production bins in NGSPS.
%    \item
%    We then use the median pollution intensity of firms within the newly defined production bins as the average pollution intensity for those bins.
%    \item
%    Lastly, we calculate the aggregate pollution by assigning to each bin the corresponding share of production. The NGSPS production bins are used for China and the employment bins are used for U.S.
%    \end{itemize}
%  \item
%  \textit{Piecewise Linear}:
%  \begin{itemize}
%    \item
%    For each U.S employment bin, we regress log-product on log-employment using the subset of Chinese firms within that employment bin. The lower and upper bounds for the production bins in this case are calculated as the predicted value of the above regression.
%    \item
%    We then run piecewise log-linear regression of pollution intensity on production within each new production bin. The average pollution intensity is chosen to be the predicted intensity at the midpoint of the new log-production bin.
%    \item
%    Lastly, the average intensity is applied to the production share distributions. The U.S distribution does not change, however, a new distribution for China is calculated since the endpoints of the production bins are different.
%    \end{itemize}
%  \item
%  \textit{Parametric}:
%  \begin{itemize}
%    \item
%    Using CNEC, we regress log-production on log-number of workers, which yields a {parametric} relationship between the number of workers and production.
%    \item
%    Using NGSPS, we regress log-intensity on log-production, which yields a {parametric} relationship between intensity and production. From these two relationships, we can subsequently construct a new \emph{parametric} relationship between intensity and number of employees. The average intensity is chosen to be the midpoint of each U.S employment bin. Notice that in this case we have a direct functional form for employment and intensity).
%    \item
%    Lastly, the average intensity is applied to the production share distributions. The U.S distribution does not change, but a new distribution for China is calculated since the endpoints of the production bins are different. Notice that this distribution for China is the one we use in Section \rom{1}.C.
%\end{itemize}
%\end{enumerate}
%
%The estimation results are shown in Table \ref{tab:accounting_appen}. Each of these three methods has its own advantages and disadvantages. The two non-parametric methods capture more of the variation at the local level, which could be washed out in a parametric estimation across the whole state space. However, this local nature also introduces a lot of instability on the estimations. Further, there are situations when there are gaps not covered by adjacent production bins and situations when these production bins overlap with each other. Under these conditions, some information will be lost while other is used for multiple times. Nevertheless, the results are robust across different estimation strategies.
%\begin{table}[t]
%\footnotesize
%\centering
%\begin{threeparttable}
%\caption{Size Distribution on Pollution}
%\begin{tabular}{lccccccc}
%    \hline \hline
%    Methods             & Paper  & Agricultural Food   & Textile    & Chemistry    & Beverage        & Average      \\
%    \midrule
%    Non-parametric      & 39.8\% & 60.7\% & 81.6\% & 102.5\% & 103.7\%      & 63.5\%  \\
%    Piecewise-linear    & 34.8\% & 69.4\% & 93.5\% & 180.1\% & N/A\tnote{a} & 75.4\%  \\
%    Parametric          & 43.5\% & 61.1\% & 97.5\% & 101.2\% & 89.0\%       & 67.0\%  \\
%    \bottomrule
%\end{tabular}
%\begin{tablenotes}
%     \footnotesize{
%     \item[$\dagger$]
%     Note: Please see notes of Table \ref{tab:top5summary} for acronyms of industries. For individual industries, the numbers reported are the aggregate pollution from the artificial U.S production structure as percentage from that of China. We use the 1st and 3rd quartile in the non-parametric calculation. Column 6 (Average) calculates the weighted average of these ratios using the percentage contribution in row one of Table \ref{tab:top5summary} as weights.}
%     \item[a]
%     Since the beverage industry has a lot fewer firms than the others, there are employment size bins with no corresponding firms in China, which invalidates the method. We set the ratio to 100\% in the calculation of the last two averages.
%\end{tablenotes}
%\label{tab:accounting_appen}
%\end{threeparttable}
%\end{table}
%
%\subsection{Proofs}
%
%In this section we provide formal proofs to the results in Section \rom{2}.C. For convenience, we state those results here again.
%
%\noindent PROOF OF LEMMA \ref{lemma:nofrictionpi_appen}:
%\begin{lemma}
%\label{lemma:nofrictionpi_appen}
%In an economy with no product market frictions, $\pi_0(z)$ and $\pi_1(z)$ are both increasing and linear with respect to $z$. In addition, the slope of $\pi_1(z)$ is steeper than that of $\pi_0(z)$:
%\begin{equation}
%    \frac{\partial \pi_0 (z)}{\partial z} = (1 - p\xi)\frac{\partial \pi_1 (z)}{\partial z}, \qquad \forall z \in Z
%\label{eq:lemma1ap}
%\end{equation}
%\end{lemma}
%\begin{proof}
%Since $k_E$ is sunk-cost, it does not affect firms' decision once is paid. The factor demand decisions for the two types of firms are therefore the same. The first order conditions for capital and labor are respectively
%\begin{align}
%    \label{eq:fock}
%    & \frac{\partial \pi_i(z)}{\partial k}: \qquad \alpha \gamma z^{1-\gamma} k^{\alpha \gamma -1} n^{(1-\alpha)\gamma} = R \\
%    \label{eq:focn}
%    & \frac{\partial \pi_i(z)}{\partial n}: \qquad (1-\alpha) \gamma z^{1-\gamma} k^{\alpha \gamma} n^{(1-\alpha)\gamma-1} = W, \qquad i = 0,1
%\end{align}
%Dividing (\ref{eq:fock}) with (\ref{eq:focn}) yields constant capital to labor ratio $h$
%\begin{equation}
%    h = \frac{k}{n} = \frac{\alpha W}{(1-\alpha)R}
%\end{equation}
%which says more capital is demanded when technology is capital intensive (higher $\alpha$) or when capital rental price $R$ low. Notice that the system of equations (\ref{eq:fock}) with (\ref{eq:focn}) is log-linear and thus has closed-form solution. With some algebra, the solutions are characterized by
%\begin{align}
%    \label{eq:n}
%    & n(z) = \Phi_1 R^{\frac{\alpha \gamma}{\gamma -1}} W^{\frac{1-\alpha \gamma}{\gamma-1}} \cdot z, \qquad \qquad \Phi_1 = \left[ \frac{(1-\alpha)^{\alpha \gamma}}{(1-\alpha)\gamma \alpha^{\alpha \gamma}}\right]^{\frac{1}{\gamma -1}} \\
%    \label{eq:k}
%    & k(z) = \Phi_2 R^{\frac{1+\gamma(\alpha -1)}{\gamma -1}} W^{\frac{\gamma(1-\alpha)}{\gamma-1}} \cdot z, \qquad \Phi_2 = \frac{\alpha}{1-\alpha}\Phi_1
%\end{align}
%Substitute the optimal solutions (\ref{eq:n}) and (\ref{eq:k}) back to the definition of profits functions (\ref{eq:pi1z}) and (\ref{eq:pi0z}), we have
%\begin{align*}
%    & \pi_0(z) = (1-p\xi)\left(\Omega - \frac{1}{1-\alpha}\Phi_1 \right) \kappa z, \qquad \Omega = \left(\frac{\alpha}{1-\alpha}\right)^{\alpha \gamma} \Phi_1^{\gamma} \text{ and } \kappa = W^{\frac{\gamma(1-\alpha)}{\gamma-1}} R^{\frac{\alpha\gamma}{\gamma-1}} \\
%    & \pi_1(z) = \left(\Omega - \frac{1}{1-\alpha}\Phi_1 \right) \kappa z - R k_E
%\end{align*}
%where it is clear that both functions are increasing and linear in $z$ and (\ref{eq:lemma1ap}) is true.
%\end{proof}
%
%\noindent PROOF OF COROLLARY \ref{lemma:frictionpi_appen}:
%\begin{corollary}
%\label{lemma:frictionpi_appen}
%Suppose the product market frictions are specified as $\max\left\{ 0, 1 - z^{\phi_1} \right\}$ with $1 - \gamma + \phi_1 > 0$, then $\pi_0(z)$ and $\pi_1(z)$ are both increasing and concave with respect to $z$. In addition, the slope of $\pi_1(z)$ is steeper than that of $\pi_0(z)$:
%\begin{equation*}
%    \frac{\partial \pi_0 (z)}{\partial z} = (1 - p\xi)\frac{\partial \pi_1 (z)}{\partial z}, \qquad \forall z \in Z
%\end{equation*}
%\end{corollary}
%\begin{proof}
%The proof is straightforward given Lemma \ref{lemma:nofrictionpi_appen}. Substituting in the tax function, $\pi_0(z)$ and $\pi_1(z)$ now becomes
%\begin{align*}
%    & \pi_0(z) = (1-p\xi)\left(\Omega - \frac{1}{1-\alpha}\Phi_1 \right) \kappa z^{\frac{1-\gamma+\phi_1}{1-\gamma}} \\
%    & \pi_1(z) = \left(\Omega - \frac{1}{1-\alpha}\Phi_1 \right) \kappa z^{\frac{1-\gamma+\phi_1}{1-\gamma}} - R k_E
%\end{align*}
%where $\Omega, \Phi_1$ and $\kappa$ are defined as in Lemma \ref{lemma:nofrictionpi_appen}.
%
%Assumption $1-\gamma+\phi_1 > 0$ guarantees the monotonicity of the profits functions. Concavity is easily verified by taking second order derivatives.
%\end{proof}
%
%\noindent PROOF OF PROPOSITION \ref{prop:threshold_appen}:
%\begin{proposition}
%\label{prop:threshold_appen}
%There exists a unique threshold $\hat{z}$ such that all household members with $z \leq \hat{z}$ choose to be workers and those with $z \geq \hat{z}$ become entrepreneurs. Further, $\hat{z}$ is the solution of $W = \pi(\hat{z})$
%\end{proposition}
%\begin{proof}
%Since the overall profit function $\pi(z)$ is the upper envelope of $\pi_0(z)$ and $\pi_1(z)$, from Lemma \ref{lemma:nofrictionpi_appen} and \ref{lemma:frictionpi_appen} we know $\pi(z)$ is monotonic increasing. It is easy to verify that $\pi(0)=0$. Therefore, as long as $0<W<\pi(\overline{z})$, we can find a unique $\hat{z}$ such that $\pi(\hat{z})=W$, where uniqueness follows from monotonicity. The condition $0<W<\pi(\overline{z})$ is guaranteed in the general equilibrium version of our model by Inada condition on the production function.
%\end{proof}
%
%\noindent PROOF OF PROPOSITION \ref{prop:friction_appen}:
%\begin{proposition}
%\label{prop:friction_appen}
%Given $k_E, W, \tau_z$ and $R$, there exist unique thresholds $\tilde{z}_n$ and $\tilde{z}_f$ such that:
%\begin{enumerate}[(i)]
%    \item
%    In the economy with no product market frictions, entrepreneurs with $z \leq \tilde{z}_n$ produce using dirty technology while those with $z > \tilde{z}_n$ produce using clean technology.
%    \item
%    In the economy with product market frictions, entrepreneurs with $z \leq \tilde{z}_f$ produce using dirty technology while those with $z > \tilde{z}_f$ produce using clean technology.
%    \item
%    $\tilde{z}_n < \tilde{z}_f$, that is, product market frictions impede the technology upgrade.
%\end{enumerate}
%\end{proposition}
%\begin{proof}
%Uniqueness follows from
%\begin{equation*}
%    \frac{\partial \pi_0 (z)}{\partial z} = (1 - p\xi)\frac{\partial \pi_1 (z)}{\partial z}, \qquad \forall z \in Z
%\end{equation*}
%and monotonicity under both the case with and without frictions.
%
%We can solve for analytical expression for $z_n$:
%\begin{equation}
%    z_n = \frac{R k_E}{p \xi \left[ \Omega \kappa^{\gamma} - \frac{1}{1-\alpha} \kappa \right]}
%\label{eq:zn}
%\end{equation}
%where $z_n$ is simply the ``distance'' ($R k_E$) over ``speed'' ($p \phi \left[ \Omega \kappa^{\gamma} - \frac{1}{1-\alpha} \kappa \right]$). The ``distance'' in both cases are the same, so eventually whether $z_f$ lies left or right to $z_n$ depends on the ``speed'' of convergence.
%
%Using expressions of the profits functions with frictions, we can show that
%\begin{equation}
%\label{eq:zf}
%    z_n = \frac{R k_E}{p \xi \left[ \Omega - \frac{1}{1-\alpha} \right] \kappa} < \left(\frac{R k_E}{p \xi \left[ \Omega - \frac{1}{1-\alpha} \right]\kappa}\right)^{\frac{1-\gamma}{1-\gamma+\phi_1}} = z_f^*
%\end{equation}
%which proves the proposition.
%
%One caveat is that the second inequality holds only if the number in the parentheses is greater than 1. We verify this in our quantitative analysis but restrain ourselves from discussing extreme cases where the condition is not hold.
%\end{proof}
%\end{appendices}
%
%\newpage
\bibliography{D:/Dissertation/Literature/Dissertation1}
%\bibliography{./Dissertation1}
\bibliographystyle{aea}
\bibliographystyle{apacite}
%\bibliographystyle{apalike}
%
%\newpage
%\begin{appendices}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%% Adjust the Layout of the Appendix %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\setcounter{equation}{0}
%\setcounter{lemma}{0}
%\setcounter{proposition}{0}
%\setcounter{table}{0}
%\setcounter{page}{1}
%\renewcommand{\thesection}{\textnormal{\Alph{section}}}
%\renewcommand{\thesubsubsection}{\textnormal{\thesubsection\arabic{subsubsection}}}
%\titleformat*{\subsubsection}{\it \center}
%\renewcommand{\theequation}{\thesubsection\arabic{equation}}
%\renewcommand{\thetable}{\thesubsection\arabic{table}}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Beginning of the Appendix %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{center}
%\textsc{Online Appendix: Not for Publication}
%\end{center}
%[TO BE COMPLETED.]
%
%The section is a backup of the contents to be contained in Online Appendix only and will be removed from the paper upon the completion of the first draft.
%
%\subsection{Detailed Descriptions of Data Sources}
%
%The source of data that we draw upon in Sections \rom{1}.B and \rom{1}.C of the main text is the \textit{National General Survey of Pollution Sources} (henceforth, NGSPS). Since to the best of our knowledge, the data were not used before in economic studies, we provide a detailed description of the dataset in this section. The content in this section is combined from:
%\begin{itemize}
%    \item
%    \textit{Census Program of the National General Survey of Pollution Sources} {\begin{CJK}{GBK}{kai}(第一次全国污染源普查方案)\end{CJK}};
%    \item
%    \textit{Regulation on National General Survey of Pollution Sources [Decree of the State Council of the People's Republic of China (No.508)]} {\begin{CJK}{GBK}{kai}(全国污染源普查条例[中华人民共和国国务院令第508号])\end{CJK}}.\footnote{An official English translation is provided at the end of the current document.}
%    \item
%    \textit{Technical Specifications Requirements of the First National General Survey of Pollution Sources} {\begin{CJK}{GBK}{kai}(第一次全国污染源普查技术规定)\end{CJK}}.
%\end{itemize}
%
%\subsubsection{General Introduction}
%
%The purpose of the NGSPS is to understand the number of pollution sources and their distribution in different industries and regions; to understand the generation, discharge and treatment of major pollutants; to establish records for key pollution sources; to build a pollution source information database and an environmental statistics platform; and to provide basis for formulating policies and plans for economic and social development and environmental protection.
%
%The term ``pollution source'' here refers to premises, facilities and equipment which discharge pollutants to environment in the process of production, living or other activities or have adverse impact on environment, as well as other sources that result in pollution.
%
%The NGSPS surveys all entities and self-employed households which have pollution sources within the borders of the People's Republic of China. The scope of the survey includes industrial pollution sources, agricultural pollution sources, domestic pollution sources, facilities for centralized treatment of pollution, and other facilities generating or discharging pollutants. In this paper, we focus on the industrial sources. The industrial sources include all production entities that belong to any of the 39 manufacturing industries according to the \textit{Industrial Classification for National Economic Activities} (GB/T4547-2002). A complete list of the industries is provided in Table .
%
%\textit{Key Sources.---}The observations in the industrial sources are divided into two groups, \textit{key sources} and \textit{regular sources}. Key firms are required to file questionnaires that are more detailed than those filed by the regular firms. A firm is categorized as a key source, if one of the following condition is satisfied:
%\begin{enumerate}
%    \item
%    All production entities that discharge pollutants that contain heavy metal, hazardous waste, and radioactive substance.
%    \item
%    All production entities that belong to the 11 heavily polluting industries, which include: Paper and Paper Products; Food Processing; Raw Chemical Materials and Chemical Products; Textile; Ferrous Metal Smelting and Rolling Processing; Food Manufacturing; Production and Supply of Electric and Heating Power; Manufacturing of Leather, Fur, and Feather; Processing of Petroleum, Coking, and Nuclear Fuel; Manufacturing of Non-metallic Mineral Goods; Ferrous Metal Smelting and Rolling Processing.
%    \item
%    Firms with more than CNY 5 million in revenue that belong to the 16 key industries, which include: Beverage Manufacturing; Medicine Manufacturing; Chemical Fibers Manufacturing; Transportation Equipment Manufacturing; Coal Mining and Washing; Non-ferrous Metal Mining; Processing of Timber, and Manufacture of Wood, Bamboo, Rattan, Palm and Straw Products; Petroleum and Natural Gas Exploitation; General Purpose Machinery Manufacturing; Ferrous Metal Mining; Non-metal Mining; Apparel, Footwear and Caps Manufacturing; Water Production and Supply; Metal Products Manufacturing; Special Purpose Machinery Manufacturing; Communication Equipment, Computers and Other Electronic Equipment.
%\end{enumerate}
%A firm is categorized as a regular firm, if none of the above conditions holds.
%
%\textit{Variables.---}The industrial sources contain information of the following variables:
%\begin{enumerate}
%    \item
%    Firm's basic registration information, geographic latitude and longitude, receiving water of firm's discharged wastewater, etc;
%    \item
%    The consumption of raw and intermediate inputs, including: water consumption, energy consumption (coal, petroleum, gas, electricity, etc), sulfur content of fuel, the consumption of hazardous intermediate input, etc.
%    \item
%    The quantity of each of the products produced by the firm.
%    \item
%    The type, size and number of the pollutants treatment equipment that the firms own.
%    \item
%    The generation, control, discharge and comprehensive utilization of various kinds of pollutants, and the operation of various kinds of pollution prevention and control facilities.
%    \item
%    The monitoring of pollutants emission, including: the date and frequency of the monitoring; the type, quantity, and concentration of the pollutant.
%\end{enumerate}
%
%\textit{Pollutants Included.---}The pollutants included for the industrial sources are those that have general implications on pollution control. In particular, they are
%\begin{enumerate}
%    \item
%    Wastewater: Chemical Oxygen Demand (COD), Ammonian, Petrochemicals, Volatile Phenols, Mercury, Cadmium, Plumbum, Arsenium, Hexavalent Chromium, Cyanidium. For Paper and Paper Product, Food Processing, Food Manufacturing and Beverage Manufacturing Industries, Five-days Biochemical Oxygen Demand ($\text{BOD}_5$) is added. For Urban Sewage Treatment Plants, Total Phosphorus, Total Nitrogen, and $\text{BOD}_5$ are added.
%    \item
%    Exhaust: Soot, Industrial Dust, Sulfur Dioxide. For Electrolytic Aluminium, Cement, Ceramic, Frosted Glass industries, Fluoride. For Vehicle Exhaust Census, Carbon Monoxide and Hydrocarbon are added.
%    \item
%    Industrial Solid Waste: Hazardous Waste (according to the \textit{National Catalogue of Hazardous Wastes}), Smelting Waste, Fly Ash, Slag, Coal Refuse, Gangue, Radioactive Slag.
%    \item
%    Plaster discharged by Desulfurization Facilities, Sludge generated by waste water treatment facilities, and remaining from burning hazardous waste.
%    \item
%    Radioactive pollution sources from utilization of Concomitant Radioactive Mineral and from Civil Nuclear Power Generation.
%\end{enumerate}
%
%\subsubsection{Methods of Measurement}
%
%There are three methods that the data of pollutants discharge are collected, which in the priority of adoption when data from multiple sources are available, are \textit{Monitoring, Method of Emission Coefficient}, and \textit{Material Balance Method}. In this section, we go over these three methods and the principles that the three methods are used in practice.
%
%\textit{Monitoring.---}Monitored data are collected by measuring the actual quantity and pollutants concentration of wastewater and exhaust to infer the annual quantity of generation and discharge of certain types of pollutants. In addition, there are three sources that monitored data are obtained, which in the priority of adoption when data from multiple sources are available are: \textit{NGSPS monitoring, historical monitoring}, and \textit{online monitoring}. The NGSPS monitoring data are measured by the field staff of the survey in the survey year. The historical monitoring data are those measured and recorded in past three years, with the latest data having the highest priority.\footnote{There are in addition four different sources of historical data, in the order of adoption when more than one source are available, these four sources are: the historical data monitored by the local environmental authorities, the data monitored upon the completion of a newly constructed project, the data monitored by a third-party agency, and those that are self-reported by the firms.} The online monitoring data are those that are automatically uploaded by computerized and Internet-connected pollutants treatment equipment. Before historical and online monitoring data are accepted, the field staff of the survey examine to ensure that the production and pollutants treatment technology have not undergone substantial changes.
%
%A pollution source is subject to reporting monitored data, if any of the following four criteria is satisfied:
%\begin{enumerate}
%    \item
%    National Monitored Key Pollution Sources {\begin{CJK}{GBK}{kai}(国控重点污染源)\end{CJK}}: all firms listed in the \textit{List of National Intensively Monitored Firms} {\begin{CJK}{GBK}{kai}(国家重点监控企业名单,环办函[2007]93号)\end{CJK}}.
%    \item
%    Facilities for Centralized Treatment of Pollution.
%    \item
%    Provincial Monitored Key Pollution Sources {\begin{CJK}{GBK}{kai}(省控重点污染源)\end{CJK}} that in sum account for 65\% of total provincial emission as is recorded in the \textit{2005 Environmental Statistic Yearbook}. That is for each of the major pollutants, starting from the most polluting firm and adding up the quantity of emission, all firms until the summation reaches the 65\% of the total provincial emission are recorded. The union of the list of firms for all major pollutants are thus the firms that belong to this category.
%    \item
%    Newly established projects since 2005 whose pollutants discharge is more than the least polluting firms in the Provincial Monitored Key Pollution Sources.
%\end{enumerate}
%These firms are later referred to as the \textit{monitored firms} in this document.
%
%Starting from 2007Q1, for all National Monitored Key Pollution Sources, Centralized Treatment Facilities, and Provincial Monitored Key Pollution Sources, waste water sources have to be monitored at least once per quarter, and exhaust sources have to be monitored at least once per six months.
%
%The actual monitoring practice is subject to the following guidelines:
%\begin{enumerate}
%    \item
%    The monitoring of Mercury, Cadmium, Hexavalent Chromium, Plumbum, and Arsenium must be conducted at the discharge outlets of the individual factory workshops, with each outlets monitored separately.
%    \item
%    Other water-based pollutants are monitored at the outlets of the factory.
%    \item
%    The flow rates of wastewater and exhaust are monitored at the same time with the monitoring of pollutant discharge.
%    \item
%    The monitoring of pollution sources are conducted at the representative polluting time periods.
%    \item
%    All monitoring technical standards of wastewater pollutants follow \textit{Technical Specifications Requirements for the Monitoring of Surface Water and Wastewater} {\begin{CJK}{GBK}{kai}(地表水和污水监测技术规范, HJ/T91-2002)\end{CJK}} and \textit{Technical Specifications Requirements for the Monitoring of Wastewater Pollutants Emission} {\begin{CJK}{GBK}{kai}(水污染物排放总量监测技术规范, HJ/T92-2002)\end{CJK}}.
%    \item
%    For unstable wastewater sources, the samples are collected using \textit{Water Ratio Automatic Sampler}.
%    \item
%    All monitoring technical standards of exhaust pollutants follow \textit{The Determination of Particulates and Sampling Methods of Gaseous Pollutants from Exhaust Gas of Stationary Sources} {\begin{CJK}{GBK}{kai}(固定污染源排气中颗粒物测定和气态污染物采样方法, GB/T16157-1996)\end{CJK}}.
%\end{enumerate}
%
%\textit{Method of Emission Coefficient.---}The generation and discharge of pollutants are calculated according to the \textit{Handbook of Emission Coefficient}, edited by the \textit{Chinese Academy of Sciences}. The benchmark coefficients are estimated based on firm's production technology, production scale, etc. The actual coefficients used are modified from the benchmark coefficients according to firm's use of intermediate inputs, managerial practice, and pollution treatment equipment installed. The actual pollutants generation and discharge are then calculated according to the actual production scale of the firm in year 2007.
%
%\textit{Material Balance Method.---}The Material Balance Method is an application of conservation of mass to the analysis of physical systems. By accounting for material entering and leaving a system, mass flows can be identified which might have been unknown, or difficult to measure without this technique. The exact conservation law used in the analysis of the system depends on the context of the problem, but all revolve around mass conservation, i.e. that matter cannot disappear or be created spontaneously. When applied, a firm's usage of intermediate inputs, energy consumption, water consumption, and production technology are considered simultaneously, in order that the calculated pollutants generation and discharge can realistically reflect the actual production and emission of the firm in reality.
%
%\textit{Principles of Data Source Selection.---}The adoption of the methods of measurement follow the following guidelines:
%\begin{enumerate}
%    \item
%    The data of the Key Pollution Sources are obtained mainly through monitoring and the method of emission coefficient, while the material balance method is only used when the other two methods are not feasible. More specifically, the data of all firms that are identified as the monitored firms are obtained by monitoring. The data of other firms in the key sources are obtained by the method of emission coefficient.
%    \item
%    The data of the Regular Pollution Sources are obtained mainly through the method of emission coefficient, while the material balance method is only used when the method of emission coefficient is not feasible.
%    \item
%    Before accepted, the monitored data are compared with those calculated from the method of emission coefficient. If the discrepancy is less than 20\% of the monitored value, the monitored value is used. If the discrepancy is larger than 20\%, the production technology and operating status of the firm is examined. If the operating status is in compliance with that specified in the technical regulations of the monitoring practice (for example, the firm has to reach at least 75\% of its production capacity when the emission is monitored), the monitored data are accepted. Otherwise, the emission is calculated using the method of emission coefficient.
%\end{enumerate}
%
%\subsubsection{Comparisons of Statistical Features}
%
%In this section, we present detailed evidence to show that the samples covered in the CNEC and those in the NGSPS are broadly comparable.
%
%\subsection{Modeling Strategies}
%\label{sec:model_appen}
%In the paper, we model the costs of adopting abatement technology as a fixed costs $k_E$, regardless of the production scale of firms. The readers may concern that this could be a mis-specification. Two alternatives arise naturally: fixed costs plus operating costs and scale-related fixed costs. In this section, we show that each alternative contradicts some of the empirical evidence we document from the data.
%
%\textit{Operating Costs.}---We start with the case of operating costs. Trimming the top 1\% observations for outliers, the distribution of the ratio of operating costs over total value of production is shown in Table \ref{tab:oprcost_appen}.
%\begin{table}[h]
%    \footnotesize
%    \centering
%    \begin{threeparttable}
%    \caption{Operating Costs as a Fraction of Output}
%    \begin{tabular}{cccccc}
%    \hline \hline
%    Minimum & 25\%      & 50\%      & 75\%      & Maximum   & Mean \\
%    \midrule
%    0       & 0.0013    & 0.0049    & 0.0143    & 0.2084    & 0.0150 \\
%    \bottomrule
%    \end{tabular}
%    \label{tab:oprcost_appen}
%    \end{threeparttable}
%\end{table}
%On average, operating costs of abatement equipment takes about 1.5\% of a firm's annual value of production. In addition, the median of the ratio is less than 0.5\%, suggesting that operating costs are negligible for more than 50\% of the firms. For conciseness consideration, we choose to omit them from the model.
%
%\textit{Fixed Costs Proportional to Production Level.}---We do observe in the data that firms with larger production scale tend to make larger investment in abatement equipments. For instance, the correlation between the log value of production and abatement equipment investment is equal to 0.64. However, the relationship significantly weakens if we remove the percentage interpretation implied by taking logarithmic. In fact, if we calculate the ratio of abatement equipment investment over total production of firms below certain quantiles, the ratio decreases gradually as we include more large firms into the calculation. Table \ref{tab:key_appen} reports the ratio of \textit{aggregate} treatment equipments installation costs as a fraction of \textit{aggregate} value of output when firms with size below certain quantiles are included.
%\begin{table}[h]
%    \footnotesize
%    \centering
%    \begin{threeparttable}
%    \caption{Installation Costs as a Fraction of Output}
%    \begin{tabular}{cccccc}
%    \hline \hline
%    25\%      & 40\%      & 60\%      & 80\%      & 95\% \\
%    \midrule
%    0.6195    & 0.2475    & 0.1151    & 0.0668    & 0.0433 \\
%    \bottomrule
%    \end{tabular}
%    \label{tab:key_appen}
%    \end{threeparttable}
%\end{table}
%Recall that we have documented that the unit-cost per processing capacity is also decreasing with the equipment capacity, together these facts suggest the existence of returns to scale of treatment equipments.
%
%\textit{Relative Prices of Physical versus Biological Equipment.}---We assume in the main text that only the installation of biological equipments requires a fixed costs. In this paragraph, we provide evidence in support of this choice. The distributions of prices for physical and biological equipments in absolute term (CNY 10,000 in year 2007) are listed in Table \ref{tab:fixcosts_appen}.
%\begin{table}[t]
%    \footnotesize
%    \centering
%    \begin{threeparttable}
%    \caption{Relative Prices of Treatment Equipments}
%    \begin{tabular}{lcccccc}
%    \hline \hline
%    Technologies    & Minimum & 25\%      & 50\%      & 75\%      & Maximum   & Mean \\
%    \midrule
%    Physical        & 0.002   & 2.000     & 6.000     & 25.000    & 800.000   & 36.140 \\
%    Biological      & 0.04    & 30.00     & 85.00     & 260.00    & 4000.00   & 249.90 \\
%    \bottomrule
%    \end{tabular}
%    \label{tab:fixcosts_appen}
%    \end{threeparttable}
%\end{table}
%As is shown in Table \ref{tab:fixcosts_appen}, the costs of biological equipments is 7 to 15 times of those of the physical equipments. We further calculate the ratio of \textit{aggregate} treatment equipments installation costs as a fraction of \textit{aggregate} value of output for different technologies. We find that the ratio is 3.07\% for physical technology and 12.83\% for biological technology. In addition, the total investment of physical equipment over biological equipment is 0.087, meaning that the installation costs of physical equipments when aggregated across the economy are only 9\% of those of the biological equipments.
%
%These evidence suggest that comparing to biological equipment, the costs of installing physical treatment equipments are nearly negligible. Therefore in our later calculation we assume that only the adoption of biological equipment is costly.
%
%In summary, the evidence suggests that firms' investment in abatement equipment is increasing with production scale but to a much lesser extent in absolute term. Again for simplicity and without loss of generality, we choose the fixed costs modeling strategy.
%
%\subsection{Intensity-Size Correlation}
%
%\subsection{Size-Distribution}
%
%\subsection{Accounting Exercise}
%
%\subsection{Calibration}
%
%\subsection{The Partial Equilibrium Model}
%\label{sec:partial_appen}
%
%\end{appendices}

\end{document}